### 2023-01-31 code/analysis_code/exploratory.qmd
Working with copies/mL dataset

---
title: "Exploratory"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
```

Make mini dataframes for each site
```{r}
copies2<- copies %>%
  select(date,copies_mL,site) %>%
drop_na()

bo<- copies2 %>% filter(site %in% "BO")
c2 <- copies2 %>% filter(site %in% "C2")
rd<- copies2 %>% filter(site %in% "RD")
summary(copies2)
```

```{r}
rd %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```

```{r}
c2 %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```

```{r}
bo %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL)
  )
```
```{r}
copies2 %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site)
```

### 2023-01-31 code/processing_code/copies_calculator.qmd
Working with copies/mL dataset

---
title: "Copies/mL"
format:
  html:
    theme: default
---
Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
data <-read_csv(here("data", "raw_data", "copies_master.csv"))
```

Select and Filter for Daily Time Series 
```{r}
data2<- data %>%
select(Type,Sample, Cq,filter_amt, dilution) %>%
  filter(Type %in% "Daily")
```

Go from CQ --> Copies/mL
```{r}
data3<- data2 %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eulted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

  select(Sample,copies_mL)
```

Create Site Column from Sample Name 
```{r}
data4<- data3 %>%
  mutate(site = substr(data3$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Take the average of qPCR triplicates 
```{r}
data5<- data4%>% group_by(Sample,site) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() 
```

Save as RDS and CSV
```{r}
saveRDS(data5, file= here("data","processed_data","clean_data.rds"))
write.csv(data5,file= here("data","processed_data","clean_data.csv")) 
```

### 2023-01-31 code/processing_code/processingcode.R

@@ -1,139 +1,0 @@
-###############################
-# processing script
-#
-#this script loads the raw data, processes and cleans it 
-#and saves it as Rds file in the processed_data folder
-#
-# Note the ## ---- name ---- notation
-# This is done so one can pull in the chunks of code into the Quarto document
-# see here: https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html
-
-
-## ---- packages --------
-#load needed packages. make sure they are installed.
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
-
-## ---- loaddata --------
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw_data","exampledata.xlsx")
-
-#load data. 
-#note that for functions that come from specific packages (instead of base R)
-# I often specify both package and function like so
-#package::function() that's not required one could just call the function
-#specifying the package makes it clearer where the function "lives",
-#but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
-
-## ---- exploredata --------
-#take a look at the data
-dplyr::glimpse(rawdata)
-
-#another way to look at the data
-summary(rawdata)
-
-#yet another way to get an idea of the data
-head(rawdata)
-
-#this is a nice way to look at data
-skimr::skim(rawdata)
-
-# looks like we have the following data
-# height (seems like it's in centimeters) 
-# weight (seems to be in kilogram)
-# Sex
-
-
-
-## ---- cleandata1 --------
-# Inspecting the data, we find some problems that need addressing:
-# First, there is an entry for height which says "sixty" instead of a number. 
-# Does that mean it should be a numeric 60? It somehow doesn't make
-# sense since the weight is 60kg, which can't happen for a 60cm person (a baby)
-# Since we don't know how to fix this, we might decide to remove the person.
-# This "sixty" entry also turned all Height entries into characters instead of numeric.
-# That conversion to character also means that our summary function isn't very meaningful.
-# So let's fix that first.
-
-d1 <- rawdata %>% dplyr::filter( Height != "sixty" ) %>% 
-                  dplyr::mutate(Height = as.numeric(Height))
-
-
-# look at partially fixed data again
-skimr::skim(d1)
-hist(d1$Height)
-
-
-## ---- cleandata2 --------
[…]

### 2023-01-31 code/processing_code/copies_calculator.qmd
Total Vibrio Figures

@@
     copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
-    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eulted DNA from ZYMO) / (filter amount from vacuum filtration)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
     copies_mL = copies_uL_extract*concentration_factor) %>%
@@
   summarize_if(is.numeric, mean) %>%
-  ungroup() 
+  ungroup() #Ungroup to spit everything back out
 ```

### 2023-01-31 code/analysis_code/exploratory_daily.qmd
Daily and Monthly Copies_mL work

---
title: "Exploratory"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
```

Make mini dataframes for each site
```{r}
#Daily All Sites
daily<- copies %>%
  select(date,copies_mL,site,Type) %>%
  filter(Type %in% "Daily") %>%
drop_na() #Get rid of NAs

#Daily by Site
bo_daily <- daily %>% filter(site %in% "BO") #Blind Oso Bay
c2_daily <- daily %>% filter(site %in% "C2") #Canals
rd_daily <- daily %>% filter(site %in% "RD") #Gulf Site

```

Total Daily Vibrio at Gulf
```{r}
rd_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkblue") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Gulf Site") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Daily Vibrio at Canals
```{r}
c2_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkorange") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Canals") +
  theme(plot.title = element_text(hjust = 0.5))
```
Total Daily Vibrio at Blind Oso
```{r}
bo_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkgreen") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Daily Vibrio
```{r}
daily %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites") +
  theme(plot.title = element_text(hjust = 0.5)) + #center title
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #Tilt x axis labels 
  theme(legend.position = "none") #Removes legend 
```

### 2023-01-31 code/analysis_code/exploratory_monthly.qmd
Daily and Monthly Copies_mL work

---
title: "Exploratory_monthly"
format:
  html:
    theme: default
---

Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
```

Make mini dataframes for each site
```{r}

#Monthly All sites
monthly<- copies %>%
  select(date,copies_mL,site,Type) %>%
  filter(Type %in% "Monthly")

#Monthly by Site
bo_monthly <- monthly %>% filter(site %in% "BO") #Blind Oso Bay
c2_monthly <- monthly %>% filter(site %in% "C2") #Canals
rd_monthly <- monthly %>% filter(site %in% "RD") #Gulf Site

```

Total Monthly Vibrio at Gulf
```{r}
rd_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkblue") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Gulf Site") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio at Canals
```{r}
c2_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkorange") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Canals") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio at Blind Oso
```{r}
bo_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkgreen") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio
```{r}
monthly %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at all Sites") +
  theme(plot.title = element_text(hjust = 0.5)) + #center title
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #Tilt x axis labels 
  theme(legend.position = "none") #Removes legend 
```

### 2023-01-31 code/processing_code/copies_calculator.qmd
Daily and Monthly Copies_mL work

@@
 
-Select and Filter for Daily Time Series 
+Select and Filter for Daily and Monthly Time Series 
 ```{r}
-data2<- data %>%
+daily1<- data %>%
 select(Type,Sample, Cq,filter_amt, dilution) %>%
   filter(Type %in% "Daily")
+
+monthly1<- data %>%
+select(Type,Sample, Cq,filter_amt, dilution) %>%
+  filter(Type %in% "Monthly")
 ```
@@
 ```{r}
-data3<- data2 %>%
+daily2<- daily1 %>%
     mutate(
@@
 
-  select(Sample,copies_mL)
+  select(Sample,copies_mL,Type)
+
+monthly2<- monthly1 %>%
+    mutate(
+    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
+    sq = 10^log_sq, #Unlog 
+    copies_rxn = sq*10, #SQ*Reaction Volume
+    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
+    copies_mL = copies_uL_extract*concentration_factor) %>%
+
+  select(Sample,copies_mL,Type)
+
 ```
@@
 ```{r}
-data4<- data3 %>%
-  mutate(site = substr(data3$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+daily3<- daily2 %>%
+  mutate(site = substr(daily2$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+
+monthly3<- monthly2 %>%
+  mutate(site = substr(monthly2$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
 ```
@@
 ```{r}
-data5<- data4%>% group_by(Sample,site) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+daily_final<- daily3 %>% group_by(Sample,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+  summarize_if(is.numeric, mean) %>%
+  ungroup() #Ungroup to spit everything back out
+
+monthly_final<- monthly3 %>% group_by(Sample,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
   summarize_if(is.numeric, mean) %>%
@@
 
+```{r}
+copies_final<- full_join(monthly_final,daily_final)
+```
+
 Save as RDS and CSV
 ```{r}
-saveRDS(data5, file= here("data","processed_data","clean_data.rds"))
-write.csv(data5,file= here("data","processed_data","clean_data.csv")) 
+saveRDS(copies_final, file= here("data","processed_data","clean_data.rds"))
+write.csv(copies_final,file= here("data","processed_data","clean_data.csv"))
+
 ```

### 2023-02-02 code/processing_code/copies_calculator.qmd
Final edits to project_README.md

@@
 ---
+*This script takes Cq value outputs from the CFX and computes it to copies of gene targets of Vibrio*
+
 Libraries
@@
 copies_final<- full_join(monthly_final,daily_final)
+summary(copies_final)
+str(copies_final)
 ```

### 2023-02-03 code/processing_code/dust_conc_processing.qmd
Transfer Part 1 from readme to Manuscript.qmd

---
title: "Dust Concentration"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
nrl<- read_csv(here("data","raw_data","nrl_conc.csv"))
```

Make new dataframes 
```{r}
dust_tot<- nrl %>%  #Dust data for all hours
  select(dust,date,hr)

dust_0<- dust_tot %>% #Dust data at t = 0
  filter(hr %in% "0")

dust_6<-dust_tot %>% #Dust data at t = 6
  filter(hr %in% "6")

dust_12<- dust_tot %>% #Dust data at t =12
  filter(hr %in% "12")

dust_18<- dust_tot %>% #Dust data at t = 18
  filter(hr %in% "18")

dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>%
  rename("dust" = "x")
#dust_sum<- dust_sum %>%
 #left_join(dust_sum)
```

```{r}
dust_sum %>% ggplot() +geom_line(
  aes(x = date, 
      y = x)
)
```

### 2023-02-03 products/manuscript/Manuscript.qmd
Transfer Part 1 from readme to Manuscript.qmd

---
title: "Nathan Greenslit Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information

Each year, plumes of Saharan dust travel across the Atlantic via easterly trade winds. These plumes are then deposited in the surface waters of the mid-Atlantic, Caribbean, and Gulf of Mexico. These dust aerosols harbor a wide range of fungi, bacteria, virus-like-particles, minerals, and nutrients (NO3, Fe, PO4). Addition of these nutrients to otherwise oligotrophic settings can result in large and rapid blooms of potentially harmful microbes, presenting a danger to both marine and human health.

Bacteria belonging to the genus Vibrio are marine opportunistic heterotrophs that are ubiquitous in nature and are among the first to respond to the influx of nutrients. Species of Vibrio are known to cause a variety of diseases in humans ranging from  cholera (V. cholerae), to ear and wound infections (V. alginolyticus), and in some extreme cases necrotizing fasciitis and septicemia (V. vulnificus). Vibrio have also been shown to impact marine health, causing coral bleaching (V. coralliiyticus, V. shiloi, V. alginolyticus), and shellfish and fish mortality (V. harveyi, V. parahaemolyticus, V. vulnificus). Vibrio population dynamics are primarily driven by temperature and salinity, with prime conditions between 25-30°C and salinities ranging from 20-35. While there is evidence of these blooms occurring in low-nutrient settings like the Florida Keys, less is known regarding microbial response to dust input in areas with *higher baseline nutrients*.

## Description of data and data source

Data was collected at 3 sites in Corpus Christi, TX before, during, and after a Saharan dust event (starting on July 7th and ending on July 19th, 2022). Sites represent a gradient in background nutrient levels (Blind Oso and Canals are high, the Gulf is low). Water samples were collected for inorganic nutrients, dissolved and particulate organic matter, and microbial analysis. Vertical profiles of salinity, temperature, pH, and dissolved oxygen were obtained using a YSI ProPlus sonde.

Variables will include `Temperature (°C)`, `Salinity`, `Nitrate`, `Phosphate`, `Copies/mL`  produced using quantitative PCR. Provides gene copies of Vibrio bacteria (I am hoping to also get data about specific species), and `Dust Concentration (AOT)` from the naval research laboratory satellite data.

## Questions/Hypotheses to be addressed

How do Saharan dust events influence Vibrio populations in high nutrient coastal waters?

*What I am looking for* 
I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels. If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response. Nutrients, dust input, salinity, and temperature will be the primary factors of interest.

*How I will analyze it* 
- Look at Vibrio growth over time series (estimated from qPCR) 
- Examine relationships between dust input and Vibrio growth 
- Examine influence of site-specific environmental parameters (Temperature, Salinity, Nutrients) on growth response 
- Run linear models on factors like: dust x growth and nutrients x growth 
- NMDS to see which parameters have the strongest influence on growth

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by sex."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References

### 2023-02-03 code/analysis_code/exploratory_daily.qmd
Added Exploratory dust qmd

@@
 ```{r}
-copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
+copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
 ```

### 2023-02-03 code/analysis_code/exploratory_dust.qmd
Added Exploratory dust qmd

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
```

Load Data
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds"))
```

Daily Dataframe
```{r}
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

Plot Daily Dust Conc
```{r}
daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust)
)
```

Plot Total Dust Conc
```{r}
dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust)
)
```

### 2023-02-03 code/analysis_code/exploratory_monthly.qmd
Added Exploratory dust qmd

@@
 ```{r}
-copies<- read_csv(here("data", "processed_data", "clean_data_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
+copies<- read_csv(here("data", "processed_data", "clean_data_copies.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
 ```

### 2023-02-03 code/processing_code/copies_calculator.qmd
Added Exploratory dust qmd

@@
 ```{r}
-saveRDS(copies_final, file= here("data","processed_data","clean_data.rds"))
-write.csv(copies_final,file= here("data","processed_data","clean_data.csv"))
+saveRDS(copies_final, file= here("data","processed_data","clean_copies.rds"))
+write.csv(copies_final,file= here("data","processed_data","clean_copies.csv"))
 

### 2023-02-09 code/analysis_code/exploratory_daily.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

@@
 ```{r}
-daily %>% ggplot()+geom_line(
+tot_vib<- daily %>% ggplot()+geom_line(
   aes(x = date,
@@
   theme(legend.position = "none") #Removes legend
+
+tot_vib
 ```
+Save as PNG
+```{r}
+png(file = here("results", "plots", "tot_vib_allsites_daily.png"))
+tot_vib
+dev.off()
+```

### 2023-02-09 code/analysis_code/exploratory_dust.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

@@
 
-Load Data
+Load Data that has just DUST CONC
 ```{r}
-dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds"))
+dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
+dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
+dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
+dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
+dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))
+
 ```
 
-Daily Dataframe
+Daily Dataframe with Sum dust
 ```{r}
 daily_dust_sum<- dust_sum %>%
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
 ```
@@
 ```{r}
-daily_dust_sum %>% ggplot() +geom_line(
+dust_sum_daily<- daily_dust_sum %>% ggplot() +geom_line(
   aes(x = date,
-      y = dust)
-)
+      y = dust))
+
+dust_sum_daily
 ```
@@
 ```{r}
-dust_sum %>% ggplot() +geom_line(
+dust_sum_2022<- dust_sum %>% ggplot() +geom_line(
   aes(x = date,
-      y = dust)
-)
+      y = dust))
+
+dust_sum_2022
 ```
+
+Save as PNG
+```{r}
+png(file = here("results","plots", "dust_sum_daily.png"))
+dust_sum_daily
+dev.off()
+
+png(file = here("results","plots", "dust_sum_2022.png"))
+dust_sum_2022
+dev.off()
+```

### 2023-02-09 code/analysis_code/exploratory_monthly.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

--- 2023-01-31 code/analysis_code/exploratory_monthly.qmd
@@
 ```{r}
-monthly %>% ggplot()+geom_line(
+tot_vib_allsites_monthly <- monthly %>% ggplot()+geom_line(
   aes(x = date,
@@
   theme(legend.position = "none") #Removes legend
+
+tot_vib_allsites_monthly
 ```
+Save as PNG
+```{r}
+png(file = here("results","plots", "tot_vib_allsites_monthly.png"))
+tot_vib_allsites_monthly
+dev.off()
+```

### 2023-02-09 code/analysis_code/exploratory_vp.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

---
title: "Vp Explore"
format:
  html:
    theme: default
---

Libraries
```{r}
library(tidyverse)
library(here)
```

Data
```{r}
vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies_date.csv")) #This file had dates manually added
sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
```
Standard Curve
```{r}
std<- sq %>%
  filter(
    Fluor %in% "HEX",
    Content %in% "Std")

std %>%
  ggplot() + geom_point(
    aes(x = log10(SQ),
        y = Cq))

```

Make Site df
```{r}
rd6<- vp %>%
  filter(site %in% "RD")

c2<- vp %>%
  filter(site %in% "C2")
```

Plot Vp at C2
```{r}
c2_vp<- c2 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "purple") +
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "purple")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Canals") +
  theme(plot.title = element_text(hjust = 0.5))

c2_vp
```

Plot Vp at RD
```{r}
rd_vp<- rd6 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "red")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "red")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vp at Gulf") +
  theme(plot.title = element_text(hjust = 0.5))

rd_vp
```

Save Vp C2/RD as PNG
```{r}
png(file = here("results", "plots", "vp_c2.png")) #Saving c2
c2_vp
dev.off()

png(file = here("results", "plots","vp_rd6.png")) #Saving RD
rd_vp
dev.off()
```

### 2023-02-09 code/analysis_code/exploratory_vv.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

---
title: "Vv explore"
format:
  html:
    theme: default
---

Libraries
```{r}
library(tidyverse)
library(here)
library(ggpubr)
```

Data
```{r}
vv<- read_csv(file = here("data", "processed_data", "vv_clean_copies_date.csv")) #This file had dates manually added
sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv"))
```

Vv Standard Curve
```{r}
std<- sq %>%
  filter(Content %in% "Std",
         Fluor %in% "HEX") #%>%
  #ggplot() + geom_point(
  #  aes(x = log(SQ),
      #  y = Cq))
#std
```

Make Site df
```{r}
rd6<- vv %>%
  filter(site %in% "RD")

c2<- vv %>%
  filter(site %in% "C2")

bo<- vv %>%
  filter(site %in% "BO")
```

Plot Vv at C2
```{r}
c2_vv<- c2 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkred") +
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkred")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Canals") +
  theme(plot.title = element_text(hjust = 0.5))

c2_vv
```

Plot Vv at RD
```{r}
rd_vv<- rd6 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkblue")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkblue")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Gulf") +
  theme(plot.title = element_text(hjust = 0.5))

rd_vv
```

Plot Vv at BO
```{r}
bo_vv<- bo %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkgreen")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkgreen")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))

bo_vv
```
Stack
```{r}
figure<- ggarrange(c2_vv, bo_vv, rd_vv + font("x.text", size = 10),
                    ncol = 1, nrow = 3) #Final plot is 1 column and 2 rows

figure
```

All sites
```{r}
all<- vv %>% 
  mutate(site = recode(site, "BO" = "Blind Oso",
                              "C2" = "Canals",
                              "RD" = "Gulf")) %>%
  ggplot() + geom_line(
    aes(x = date,
        y = copies_mL,
        color = site)) + 
  scale_color_manual( values = c("Blind Oso" = "darkgreen",
                                "Canals" = "darkred",
                                "Gulf" = "darkblue"))+
  theme_bw()+
  facet_wrap(~site)+
  labs(
    x = "Date", 
    y = "Copies/mL",
    title = "Vibrio vulnificus Levels Across Dust Time Series")+
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5))

all
```

Save Vv at all sites as PNG
```{r}
png(file = here("results", "plots", "vv_all_sites.png")) #Saving c2
all
dev.off()

```

### 2023-02-09 code/analysis_code/statiscial_analysis.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

---
title: "Statistical analysis"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data that has COPIES and DUST CONC
```{r}
dust_sum<- readRDS(here("data", "processed_data", "dust_sum_copies.rds"))
dust_0 <- readRDS(here("data", "processed_data", "dust_0_copies.rds"))
dust_6 <- readRDS(here("data", "processed_data", "dust_6_copies.rds"))
dust_12 <- readRDS(here("data", "processed_data", "dust_12_copies.rds"))
dust_18 <- readRDS(here("data", "processed_data", "dust_18_copies.rds"))
```

Make df for Daily time series with dust sum 
```{r}
daily<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
```

df with daily for t0-t18
```{r}
daily0<- dust_0 %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

daily6<- dust_6 %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

daily12<- dust_12 %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

daily18<- dust_18 %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
```

Running Stats
```{r}
tsum<- lm(dust~copies_mL, data = daily)
summary(tsum)

t0<- lm(dust~copies_mL, data = daily0)
summary(t0)

t6<- lm(dust~copies_mL, data = daily6)
summary(t6)

t12<- lm(dust~copies_mL, data = daily12)
summary(t12)
```

### 2023-02-09 code/processing_code/vp_copies_calculator.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

---
title: "Vp qPCR"
format:
  html:
    theme: default
---
*This dataset is Daily Vp qPCR*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
```
Filter for HEX dye (samples). Cy5 is for IAC
```{r}
sq2<- sq %>% 
  select(filter_amt,SQ, Sample, Fluor) %>%
  filter(Fluor %in% "HEX")

```

SQ --> Copies/mL
```{r}
copies<- sq2 %>%
    mutate(
    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    #sq = 10^log_sq, #Unlog 
    copies_rxn = SQ*15, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *1, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor)%>%

  select(Sample,copies_mL)
```

Create Site Column from Sample Name 
```{r}
copies2<- copies %>%
  mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Drop NAs
```{r}
copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
  select(copies_mL,site, Sample)
```
*Dropping NAs got rid of BO completely*

Take the average of qPCR triplicates 
```{r}
copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

```

Save as RDS and manually add date in excel
```{r}
write.csv(copies4, file= here("data","processed_data","vp_clean_copies.csv"))
```

### 2023-02-09 code/processing_code/vv_copies_calclulator.qmd
Commented out Dr. Handel's code for manuscript
Now it can properly render Manuscript.qmd

---
title: "Vv caluclator"
format:
  html:
    theme: default
---
*This dataset is Daily Vv qPCR*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
```

Filter for HEX dye (samples). Cy5 is for IAC
```{r}
sq2<- sq %>% 
  select(filter_amt,SQ, Sample, Fluor, dilution) %>%
  filter(Fluor %in% "HEX")

```

SQ --> Copies/mL
```{r}
copies<- sq2 %>%
    mutate(
    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    #sq = 10^log_sq, #Unlog 
    copies_rxn = SQ*15, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor)%>%

  select(Sample,copies_mL)
```

Create Site Column from Sample Name 
```{r}
copies2<- copies %>%
  mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Drop NAs
```{r}
copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
  select(copies_mL,site, Sample)
```

Take the average of qPCR triplicates 
```{r}
copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

```

Save as RDS and manually add date in excel
```{r}
write.csv(copies4, file= here("data","processed_data","vv_clean_copies.csv"))
```

### 2023-02-09 code/processing_code/vv_copies_calclulator.qmd
Vv data

@@
 sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
+redo<- read_csv(here("data", "raw_data", "vv_qPCR_daily_0_redo.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
 ```
@@
 sq2<- sq %>% 
-  select(filter_amt,SQ, Sample, Fluor, dilution) %>%
+  select(filter_amt,SQ,Cq, Sample, Fluor, dilution) %>%
+  filter(Fluor %in% "HEX")
+
+redo2<- redo %>% 
+  select(filter_amt,Cq, Sample, Fluor, dilution) %>%
   filter(Fluor %in% "HEX")
@@
     mutate(
-    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
-    #sq = 10^log_sq, #Unlog 
-    copies_rxn = SQ*15, #SQ*Reaction Volume
+    log_sq = ((Cq-34.934)/-3.1444), #Equation of the line to get log(SQ)
+    sq = 10^log_sq, #Unlog 
+    copies_rxn = sq*15, #SQ*Reaction Volume
+    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
+    copies_mL = copies_uL_extract*concentration_factor)%>%
+
+  select(Sample,copies_mL)
+
+copies_redo<- redo2 %>%
+    mutate(
+    log_sq = ((Cq-34.934)/-3.1444), #Equation of the line to get log(SQ)
+    sq = 10^log_sq, #Unlog 
+    copies_rxn = sq*15, #SQ*Reaction Volume
     copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
@@
   mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+
+copies2_redo<- copies_redo %>%
+  mutate(site = substr(copies_redo$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+
 ```
@@
 ```{r}
-copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
+copies3<- copies2 %>% 
+  filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
+  select(copies_mL,site, Sample)
+
+copies3_redo<- copies2_redo %>%
+   filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
   select(copies_mL,site, Sample)
@@
 ```{r}
-copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+copies4<- copies3 %>% 
+  group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
   summarize_if(is.numeric, mean) %>%
@@
 
+copies4_redo<- copies3_redo %>% 
+  group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+  summarize_if(is.numeric, mean) %>%
+  ungroup() #Ungroup to spit everything back out 
 ```
@@
 write.csv(copies4, file= here("data","processed_data","vv_clean_copies.csv"))
+write.csv(copies4_redo, file= here("data","processed_data","vv_clean_copies_redo.csv"))
 ```

### 2023-02-09 code/analysis_code/exploratory_vp.qmd
Corrections to Vv and Vp calcs

@@
 vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies_date.csv")) #This file had dates manually added
-sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
+sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) #Manually replaced NaN with 0 for no amplifcation 
 ```
@@
 ```
+All Sites
+```{r}
+all<- vp %>% 
+    mutate(site = recode(site, "BO" = "Blind Oso",
+                              "C2" = "Canals",
+                              "RD" = "Gulf")) %>%
+  ggplot() + geom_line(
+  aes(x = date,
+      y = copies_mL,
+      color = site)) +
+  facet_wrap(~site)+
+  geom_point(
+    aes(x = date,
+      y = copies_mL,
+      color = site))+
+  theme_bw() +
+  labs(x = "Date",
+       y = "Copies/mL",
+       title = "Vp at All Sites") +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)) +
+
+ scale_color_manual( values = c("Blind Oso" = "darkgreen",
+                                "Canals" = "darkred",
+                                "Gulf" = "darkblue"))
+
+all
+```
 
@@
 ```{r}
-png(file = here("results", "plots", "vp_c2.png")) #Saving c2
-c2_vp
+png(file = here("results", "plots", "vp_all_sites.png")) #Saving c2
+all
 dev.off()
 
-png(file = here("results", "plots","vp_rd6.png")) #Saving RD
-rd_vp
-dev.off()
 ```

### 2023-02-09 code/analysis_code/exploratory_vv.qmd
Corrections to Vv and Vp calcs

@@
 vv<- read_csv(file = here("data", "processed_data", "vv_clean_copies_date.csv")) #This file had dates manually added
-sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv"))
+#sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv"))
+redo<- read_csv(here("data", "processed_data", "vv_clean_copies_redo_date.csv"))
 ```
@@
 ```{r}
-std<- sq %>%
-  filter(Content %in% "Std",
-         Fluor %in% "HEX") #%>%
+#std<- sq %>%
+ # filter(Content %in% "Std",
+        # Fluor %in% "HEX") #%>%
   #ggplot() + geom_point(
@@
     y = "Copies/mL",
-    title = "Vibrio vulnificus Levels Across Dust Time Series")+
+    title = "Vibrio vulnificus Levels Across Dust Time Series: Trial 1")+
   theme(
@@
 all
+```
+All Redo
+```{r}
+all_redo<- redo %>% 
+  mutate(site = recode(site, "BO" = "Blind Oso",
+                              "C2" = "Canals",
+                              "RD" = "Gulf")) %>%
+  ggplot() + geom_line(
+    aes(x = date,
+        y = copies_mL,
+        color = site)) + 
+  scale_color_manual( values = c("Blind Oso" = "darkgreen",
+                                "Canals" = "darkred",
+                                "Gulf" = "darkblue"))+
+  theme_bw()+
+  facet_wrap(~site)+
+  labs(
+    x = "Date", 
+    y = "Copies/mL",
+    title = "Vibrio vulnificus Levels Across Dust Time Series: Trial 2")+
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5))
+
+all_redo
 ```
@@
 
+png(file = here("results", "plots", "vv_all_sites_redo.png")) #Saving c2
+all_redo
+dev.off()
+
 ```

### 2023-02-09 code/processing_code/vp_copies_calculator.qmd
Corrections to Vv and Vp calcs

@@
 ```{r}
-sq <-read_csv(here("data", "raw_data", "vp_qPCR_0.csv")) #Manually replaced NaN with 0 for no amplifcation 
+sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) 
 ```
@@
 
-Drop NAs
+Change NA to 0 and drop STD CRV NAs
 ```{r}
-copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
-  select(copies_mL,site, Sample)
+copies3<- copies2 %>% 
+ mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
+filter(!is.na(Sample))  #Dropping NAs from STD Curve
 ```

### 2023-02-09 code/processing_code/vv_copies_calclulator.qmd
Corrections to Vv and Vp calcs

@@
 ```{r}
-sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
-redo<- read_csv(here("data", "raw_data", "vv_qPCR_daily_0_redo.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
+sq <-read_csv(here("data", "raw_data", "vv_run_daily_TX_dust_2_7_23 -  Quantification Summary.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
+redo<- read_csv(here("data", "raw_data", "vv_run_daily_TX_dust_2_9_23 -  Quantification Summary.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
 ```
@@
 redo2<- redo %>% 
-  select(filter_amt,Cq, Sample, Fluor, dilution) %>%
+  select(filter_amt,Cq, Sample, Fluor, dilution, Well) %>%
+  filter(Fluor %in% "HEX",
+         !Well %in% "K19")
+
+SQ_test<- sq %>% 
+  select(filter_amt,SQ,Cq, Sample, Fluor, dilution) %>%
   filter(Fluor %in% "HEX")
@@
 
-SQ --> Copies/mL
+Cq --> Copies/mL
 ```{r}
@@
    mutate(
-    log_sq = ((Cq-34.934)/-3.1444), #Equation of the line to get log(SQ)
+   log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
     sq = 10^log_sq, #Unlog 
@@
     concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
-    copies_mL = copies_uL_extract*concentration_factor)%>%
+    copies_mL = copies_uL_extract*concentration_factor) %>%
 
@@
     mutate(
-    log_sq = ((Cq-34.934)/-3.1444), #Equation of the line to get log(SQ)
+    log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
     sq = 10^log_sq, #Unlog 
@@
     concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
-    copies_mL = copies_uL_extract*concentration_factor)%>%
+    copies_mL = copies_uL_extract*concentration_factor) %>%
+
+select(Sample,copies_mL)
+
+SQ_test1<- sq %>%
+    mutate(
+   # log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+   # sq = 10^log_sq, #Unlog 
+    copies_rxn = SQ*15, #SQ*Reaction Volume
+    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
+    copies_mL = copies_uL_extract*concentration_factor) %>%
 
@@
 
-Drop NAs
+Replace NaN with 0 for copies/mL
 ```{r}
 copies3<- copies2 %>% 
-  filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
-  select(copies_mL,site, Sample)
+ mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
+filter(!is.na(Sample))  #Dropping NAs from STD Curve
 
 copies3_redo<- copies2_redo %>%
-   filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
-  select(copies_mL,site, Sample)
+ mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
+filter(!is.na(Sample))  #Dropping NAs from STD Curve
 ```

### 2023-02-10 code/analysis_code/exploratory_vv.qmd
No longer need to edit raw data

@@
 ```{r}
-vv<- read_csv(file = here("data", "processed_data", "vv_clean_copies_date.csv")) #This file had dates manually added
+vv<- read_csv(file = here("data", "processed_data", "vv_clean_copies.csv")) 
 #sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv"))
-redo<- read_csv(here("data", "processed_data", "vv_clean_copies_redo_date.csv"))
+redo<- read_csv(here("data", "processed_data", "vv_clean_copies_redo.csv"))
 ```

### 2023-02-10 code/processing_code/vv_copies_calclulator.qmd
No longer need to edit raw data

--- 2023-02-09 code/processing_code/vv_copies_calclulator.qmd
@@
 library(here)
+library(ggpubr)
 ```
@@
 ```{r}
-sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv")) #Manually replaced NaN with 0 for no amplifcation, added filter_amt and dilution
+sq <-read_csv(here("data", "raw_data", "vv_run_daily_TX_dust_2_7_23 -  Quantification Summary.csv")) 
+redo<- read_csv(here("data", "raw_data", "vv_run_daily_TX_dust_2_9_23 -  Quantification Summary.csv")) 
 ```
@@
 sq2<- sq %>% 
-  select(filter_amt,SQ, Sample, Fluor, dilution) %>%
+  select(filter_amt,SQ,Cq, Sample, Fluor, dilution) %>%
   filter(Fluor %in% "HEX")
+
+redo2<- redo %>% 
+  select(Cq, Sample, Fluor, Well) %>%
+  filter(Fluor %in% "HEX",
+         !Well %in% "K19")
 
@@
 
-SQ --> Copies/mL
+Create Site Column from Sample Name 
 ```{r}
-copies<- sq2 %>%
+site<- sq %>%
+  mutate(site = substr(sq$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+
+site_redo<- redo2 %>%
+  mutate(site = substr(redo2$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+
+```
+
+Add Dilution Column and Filter column
+```{r}
+dilution<- site%>%
+  mutate(DIL = case_when(site == "BO"~"10",
+                              site  == "C2"~ "1",
+                              site == "RD"~"1"), 
+        dilution = as.numeric(dilution))
+
+dilution_redo<- site_redo%>%
+  mutate(dilution = case_when(site == "BO"~"10",
+                              site  == "C2"~ "1",
+                              site == "RD"~"1"),
+         dilution = as.numeric(dilution))
+
+filter<- dilution %>%
+   mutate(FIL = case_when(Sample == "BO_7"~"100",
+                          Sample == "BO_8"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_11"~"100",
+                          Sample == "BO_12"~"50",
+                          Sample == "BO_13"~"50",
+                          Sample == "BO_14"~"50",
+                          Sample == "BO_15"~"50",
+                          Sample == "BO_16"~"30",
+                          Sample == "BO_17"~"30",
+                          Sample == "BO_18"~"30",
+                          Sample == "BO_19"~"30",
+                              site  == "C2"~ "100",
+                              site == "RD"~"100"),
+        filter_amt = as.numeric(filter_amt))
+
+filter_redo<- dilution_redo %>%
+   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
+                          Sample == "BO_8"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_11"~"100",
+                          Sample == "BO_12"~"50",
+                          Sample == "BO_13"~"50",
+                          Sample == "BO_14"~"50",
+                          Sample == "BO_15"~"50",
+                          Sample == "BO_16"~"30",
+                          Sample == "BO_17"~"30",
+                          Sample == "BO_18"~"30",
+                          Sample == "BO_19"~"30",
+                              site  == "C2"~ "100",
+                              site == "RD"~"100"),
+          filter_amt = as.numeric(filter_amt)) 
+```
+
+Cq --> Copies/mL
+```{r}
+copies<- filter %>%
    mutate(
-    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
-    #sq = 10^log_sq, #Unlog 
+   log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+    sq = 10^log_sq, #Unlog 
+    copies_rxn = sq*15, #SQ*Reaction Volume
+    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
+    copies_mL = copies_uL_extract*concentration_factor) %>%
+
+select(Sample,copies_mL,site,Fluor) %>%
+  filter(Fluor%in%"HEX")
+
+copies_redo<- filter_redo %>%
+    mutate(
+    log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+    sq = 10^log_sq, #Unlog 
+    copies_rxn = sq*15, #SQ*Reaction Volume
+    copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
+    copies_mL = copies_uL_extract*concentration_factor) %>%
+
+select(Sample,copies_mL, site)
+
+```
+
+Replace NaN with 0 for copies/mL
+```{r}
+replace<- copies %>% 
+ mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
+filter(!is.na(Sample))   #Dropping NAs from STD Curve
+
+replace_redo<- copies_redo %>%
+ mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
+filter(!is.na(Sample))   
+ #Dropping NAs from STD Curve
+```
+
+Date Column
+```{r}
+date<- replace %>%
+  mutate(day = substr(replace$Sample,4,5)) %>% #Creates new column `date` 
+  mutate(date = case_when(day == "7"~"2022-07-07",
+                         day ==  "8"~"2022-07-08",
+                         day ==  "9"~"2022-07-09",
+                         day ==  "11"~"2022-07-11",
+                         day == "12"~"2022-07-12",
+                         day ==  "13"~"2022-07-13",
+                         day ==  "14"~"2022-07-14",
+                         day ==  "15"~"2022-07-15",
+                         day ==  "16"~"2022-07-16",
+                         day == "17"~"2022-07-17",
+                         day ==  "18"~"2022-07-18",
+                         day == "19"~"2022-07-19"))
+
+date_redo<- replace_redo %>%
+  mutate(day = substr(replace_redo$Sample,4,5)) %>% #Creates new column `date`
+  mutate(date = case_when(day == "7"~"2022-07-07",
+                         day ==  "8"~"2022-07-08",
+                         day ==  "9"~"2022-07-09",
+                         day ==  "11"~"2022-07-11",
+                         day == "12"~"2022-07-12",
+                         day ==  "13"~"2022-07-13",
+                         day ==  "14"~"2022-07-14",
+                         day ==  "15"~"2022-07-15",
+                         day ==  "16"~"2022-07-16",
+                         day == "17"~"2022-07-17",
+                         day ==  "18"~"2022-07-18",
+                         day == "19"~"2022-07-19"))
+
+```
+
+Take the average of qPCR triplicates 
+```{r}
+final<- date %>% 
+  group_by(site, Sample,date) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+  summarize_if(is.numeric, mean) %>%
+  ungroup() %>% #Ungroup to spit everything back out 
+  select(date,copies_mL,site)
+
+final_redo<- date_redo %>% 
+  group_by(site, Sample,date) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+  summarize_if(is.numeric, mean) %>%
+  ungroup() %>% #Ungroup to spit everything back out 
+  select(date,copies_mL,site)
+```
+
+Save as RDS and manually add date in excel
+```{r}
+write.csv(final, file= here("data","processed_data","vv_clean_copies.csv"))
+write.csv(final_redo, file= here("data","processed_data","vv_clean_copies_redo.csv"))
+```
+
+_________________
+Compare test
+```{r}
+SQ<- sq %>% 
+  select(filter_amt,SQ,Cq, Sample, Fluor, dilution) %>%
+  filter(Fluor %in% "HEX")
+
+SQ_copies<- SQ %>%
+    mutate(
+   # log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+   # sq = 10^log_sq, #Unlog 
     copies_rxn = SQ*15, #SQ*Reaction Volume
@@
     concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
-    copies_mL = copies_uL_extract*concentration_factor)%>%
+    copies_mL = copies_uL_extract*concentration_factor) %>%
+  select(Sample,copies_mL)
 
-  select(Sample,copies_mL)
+comp<-copies %>% filter(Fluor %in% "HEX")
+
+  bind<- SQ_copies %>% cbind(x = comp$copies_mL) %>%
+  select(x,copies_mL, Sample) %>%
+  rename("Curve Copies" = "x",
+         "Equation Copies" = "copies_mL")
 ```
 
-Create Site Column from Sample Name 
 ```{r}
-copies2<- copies %>%
-  mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
-```
+fig1<- bind %>% ggplot() +geom_boxplot(
+  aes(
+    x = Sample,
+    y = `Equation Copies`),
+  color = "darkblue") +
+  theme_bw()+
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
-Drop NAs
-```{r}
-copies3<- copies2 %>% filter(!is.na(copies_mL)) %>% #Dropping NAs from STD Curve
-  select(copies_mL,site, Sample)
-```
+fig2<- bind %>% ggplot() +geom_boxplot(
+  aes(
+    x = Sample,
+    y = `Curve Copies`),
+  color = "darkgreen") +
+  theme_bw()+
+  theme(axis.text.x = element_text(angle = 45, 
+                                   hjust = 1))
 
-Take the average of qPCR triplicates 
-```{r}
-copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
-  summarize_if(is.numeric, mean) %>%
-  ungroup() #Ungroup to spit everything back out
+ggarrange(fig1, fig2,
+          ncol = 1,
+          nrow= 2)
 
 ```
-
-Save as RDS and manually add date in excel
-```{r}
-write.csv(copies4, file= here("data","processed_data","vv_clean_copies.csv"))
-```

### 2023-02-10 code/processing_code/vp_copies_calculator.qmd
Made edits to Vp as well
Now i no longer need to alter any raw data

@@
 
+*This dataset is Daily Vv qPCR*
+
 Libraries
@@
 ```{r}
-sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) 
+sq <-read_csv(here("data", "raw_data", "vp_run_TX_Dust_2_3_23 -  Quantification Summary.csv")) 
 ```
+
 Filter for HEX dye (samples). Cy5 is for IAC
@@
 sq2<- sq %>% 
-  select(filter_amt,SQ, Sample, Fluor) %>%
+  select(SQ,Cq, Sample, Fluor) %>%
   filter(Fluor %in% "HEX")
@@
 
-SQ --> Copies/mL
+Create Site Column from Sample Name 
 ```{r}
-copies<- sq2 %>%
+site<- sq2 %>%
+  mutate(site = substr(sq2$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+```
+
+Add Dilution Column and Filter column
+```{r}
+dilution<- site%>%
+  mutate(dilution = case_when(site == "BO"~"10",
+                              site  == "C2"~ "1",
+                              site == "RD"~"1"), 
+        dilution = as.numeric(dilution))
+
+filter<- dilution %>%
+   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
+                          Sample == "BO_8"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_9"~"100",
+                          Sample == "BO_11"~"100",
+                          Sample == "BO_12"~"50",
+                          Sample == "BO_13"~"50",
+                          Sample == "BO_14"~"50",
+                          Sample == "BO_15"~"50",
+                          Sample == "BO_16"~"30",
+                          Sample == "BO_17"~"30",
+                          Sample == "BO_18"~"30",
+                          Sample == "BO_19"~"30",
+                              site  == "C2"~ "100",
+                              site == "RD"~"100"),
+        filter_amt = as.numeric(filter_amt))
+
+```
+
+Cq --> Copies/mL
+```{r}
+copies<- filter %>%
    mutate(
-    #log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
-    #sq = 10^log_sq, #Unlog 
+   #log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+   # sq = 10^log_sq, #Unlog 
     copies_rxn = SQ*15, #SQ*Reaction Volume
     copies_uL_1_10 = copies_rxn/1.2, #Copies/RXN / DNA Template Volume
-    copies_uL_extract = copies_uL_1_10 *1, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
+    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
     concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
-    copies_mL = copies_uL_extract*concentration_factor)%>%
+    copies_mL = copies_uL_extract*concentration_factor) %>%
 
-  select(Sample,copies_mL)
+select(Sample,copies_mL,site)
+
 ```
 
-Create Site Column from Sample Name 
+Replace NaN with 0 for copies/mL
 ```{r}
-copies2<- copies %>%
-  mutate(site = substr(copies$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
-```
-
-Change NA to 0 and drop STD CRV NAs
-```{r}
-copies3<- copies2 %>% 
+replace<- copies %>% 
  mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
@@
 ```
-*Dropping NAs got rid of BO completely*
+
+Date Column
+```{r}
+date<- replace %>%
+  mutate(day = substr(replace$Sample,4,5)) %>% #Creates new column `date` 
+  mutate(date = case_when(day == "7"~"2022-07-07",
+                         day ==  "8"~"2022-07-08",
+                         day ==  "9"~"2022-07-09",
+                         day ==  "11"~"2022-07-11",
+                         day == "12"~"2022-07-12",
+                         day ==  "13"~"2022-07-13",
+                         day ==  "14"~"2022-07-14",
+                         day ==  "15"~"2022-07-15",
+                         day ==  "16"~"2022-07-16",
+                         day == "17"~"2022-07-17",
+                         day ==  "18"~"2022-07-18",
+                         day == "19"~"2022-07-19"))
+
+```
 
@@
 ```{r}
-copies4<- copies3 %>% group_by(site, Sample) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
+final<- date %>% 
+  group_by(site, Sample,date) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
   summarize_if(is.numeric, mean) %>%
-  ungroup() #Ungroup to spit everything back out
-
+  ungroup() %>% #Ungroup to spit everything back out 
+  select(date,copies_mL,site)
 ```
@@
 ```{r}
-write.csv(copies4, file= here("data","processed_data","vp_clean_copies.csv"))
+#stop() #Write error so it wont overwrite
+write.csv(final, file= here("data","processed_data","vp_clean_copies.csv"))
 ```

### 2023-02-10 code/processing_code/vv_copies_calclulator.qmd
Made edits to Vp as well
Now i no longer need to alter any raw data

@@
 ```{r}
-site<- sq %>%
-  mutate(site = substr(sq$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
+site<- sq2 %>%
+  mutate(site = substr(sq2$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
 
@@
 dilution<- site%>%
-  mutate(DIL = case_when(site == "BO"~"10",
+  mutate(dilution = case_when(site == "BO"~"10",
                               site  == "C2"~ "1",
@@
 filter<- dilution %>%
-   mutate(FIL = case_when(Sample == "BO_7"~"100",
+   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
                           Sample == "BO_8"~"100",
@@
 ```{r}
+#stop() #Write error so it wont overwrite
 write.csv(final, file= here("data","processed_data","vv_clean_copies.csv"))

### 2023-02-12 code/analysis_code/exploratory_monthly.qmd
Total vib
Do not need to manually add in raw data

@@
 ```{r}
-copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
+copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
 ```
@@
   select(date,copies_mL,site,Type) %>%
-  filter(Type %in% "Monthly")
+  filter(Type %in% "Monthly") #Need to add July date here somehow
 
@@
 ```{r}
-tot_vib_allsites_monthly <- monthly %>% ggplot()+geom_line(
+tot_vib_allsites_monthly <- monthly %>%
+  ggplot()+geom_line(
   aes(x = date,
@@
       color = site))+
+  geom_point(
+    aes(x = date,
+        y = copies_mL,
+        color = site ))+
   facet_wrap(~site) +
@@
     y = "Copies/mL",
-    title = "Total Monthly Vibrio at all Sites") +
+    title = "Total Monthly Vibrio at all Sites",
+    caption = "This includes 3 May events and does not include July") +
   theme(plot.title = element_text(hjust = 0.5)) + #center title

### 2023-02-12 code/processing_code/total_copies_calculator.qmd
Total vib
Do not need to manually add in raw data

---
title: "Copies/mL"
format:
  html:
    theme: default
---
*This script takes Cq value outputs from the CFX and computes it to copies of gene targets of Vibrio*

Libraries
```{r}
library(tidyverse)
library(here)
```

Load data
```{r}
data <-read_csv(here("data", "raw_data", "copies_master.csv"))
```

Select and Filter for Daily and Monthly Time Series 
```{r}
daily<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Daily")

monthly<- data %>%
select(Type,Sample, Cq) %>%
  filter(Type %in% "Monthly")
```

Create Site Column from Sample Name 
```{r}
site_daily<- daily %>%
  mutate(site = substr(daily$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)

site_monthly<- monthly %>%
  mutate(site = substr(monthly$Sample,1,2)) #Creates new column `Site` from the first two letters in `Sample` (BO16 becomes BO)
```

Add Dilution Column and Filter column
```{r}
dilution_daily<- site_daily %>%
  mutate(dilution = case_when(
                          Sample == "BO7"~"10",
                          Sample == "BO8"~"10",
                          Sample == "BO9"~"10",
                          Sample == "BO11"~"10",
                          Sample == "BO12"~"10",
                          Sample == "BO13"~"10",
                          Sample == "BO14"~"10",
                          Sample == "BO15"~"10",
                          Sample == "BO16REDO"~"1",
                          Sample == "BO17"~"10",
                          Sample == "BO18"~"10",
                          Sample == "BO19"~"10",
                              site  == "C2"~ "1",
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

dilution_monthly<- site_monthly %>%
  mutate(dilution = case_when(
                              site == "BO"~ "1",
                              site  == "C2"~ "1",
                              site == "RD"~"1"), 
        dilution = as.numeric(dilution))

filter_daily<- dilution_daily %>%
   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
                          Sample == "BO7"~"100",
                          Sample == "BO8"~"100",
                          Sample == "BO9"~"100",
                          Sample == "BO11"~"100",
                          Sample == "BO12"~"50",
                          Sample == "BO13"~"50",
                          Sample == "BO14"~"50",
                          Sample == "BO15"~"50",
                          Sample == "BO16REDO"~"30",
                          Sample == "BO17"~"30",
                          Sample == "BO18"~"30",
                          Sample == "BO19"~"30",
                              site  == "C2"~ "100",
                              site == "RD"~"100"),
        filter_amt = as.numeric(filter_amt))

filter_monthly<- dilution_monthly %>%
   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
                          Sample == "BO_FEB"~"100",
                          Sample == "BO_MAR"~"100",
                          Sample == "BO_APR"~"100",
                          Sample == "BO_MAY_11"~"100",
                          Sample == "BO_MAY_20"~"100",
                          Sample == "BO_MAY_24"~"100",
                          Sample == "BO_JUN"~"100",
                          Sample == "BO_AUG"~"30",
                          Sample == "BO_SEP"~"45",
                          Sample == "BO_OCT"~"60",
                          Sample == "BO_NOV"~"30",
                          Sample == "BO_DEC"~"60",
                              site  == "C2"~ "100",
                              site == "RD"~"100"),
        filter_amt = as.numeric(filter_amt)) 
```

Go from CQ --> Copies/mL
```{r}
copies_daily<- filter_daily %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

  select(Sample,copies_mL,site,Type)

copies_monthly<- filter_monthly %>%
    mutate(
    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
    sq = 10^log_sq, #Unlog 
    copies_rxn = sq*10, #SQ*Reaction Volume
    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
    copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
    concentration_factor = 100/filter_amt, #(Concentration factor = amount of eluted DNA from ZYMO) / (filter amount from vacuum filtration)
    copies_mL = copies_uL_extract*concentration_factor) %>%

 select(Sample,copies_mL,Type, site)

```

Replace NaN with 0 for copies/mL
```{r}
replace_daily<- copies_daily %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve

replace_monthly<- copies_monthly %>% 
 mutate(copies_mL = ifelse(is.na(copies_mL), 0, copies_mL)) %>%
filter(!is.na(Sample))   #Dropping NAs from STD Curve 
```

Date Column
```{r}
date_daily<- replace_daily %>%
  mutate(day = substr(replace_daily$Sample,3,4)) %>% #Creates new column `day` 
  mutate(date = case_when(day == "7"~"2022-07-07", #Makes date based off of day
                          day == "7R"~"2022-07-07",
                         day ==  "8"~"2022-07-08",
                         day ==  "9"~"2022-07-09",
                         day ==  "11"~"2022-07-11",
                         day ==  "12"~"2022-07-12",
                         day ==  "13"~"2022-07-13",
                         day ==  "14"~"2022-07-14",
                         day ==  "15"~"2022-07-15",
                         day ==  "16"~"2022-07-16",
                         day ==  "17"~"2022-07-17",
                         day ==  "18"~"2022-07-18",
                         day ==  "19"~"2022-07-19")) %>%
  select(copies_mL,site,date,Type)

date_monthly<- replace_monthly %>%
  mutate(month = substr(replace_monthly$Sample,4,9)) %>% #Creates new column `month` 
  mutate(date = case_when(month == "FEB"~"2022-02-16", #Makes date based off of day
                         month ==  "MAR"~"2022-03-09",
                         month ==  "APR"~"2022-04-20",
                         month ==  "MAY_11"~"2022-05-11",
                         month ==  "MAY_20"~"2022-05-20",
                         month ==  "MAY_24"~"2022-05-24",
                         month ==  "JUN"~"2022-06-14",
                         month ==  "AUG"~"2022-08-09",
                         month ==  "SEP"~"2022-09-13",
                         month ==  "OCT"~"2022-10-18",
                         month ==  "NOV"~"2022-11-15",
                         month ==  "DEC"~"2022-12-14")) %>%
  select(copies_mL,site,date,Type)

```

Take the average of qPCR triplicates 
```{r}
final_daily<- date_daily %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
  summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out

final_monthly<- date_monthly %>% group_by(date,site,Type) %>% #Group all chr together. All that should be left out is the numeric vector we want to work with (in this case, copies_mL)
 summarize_if(is.numeric, mean) %>%
  ungroup() #Ungroup to spit everything back out
```

Combine Daily and Monthly 
```{r}
copies_final<- full_join(final_monthly,final_daily)
summary(copies_final)
str(copies_final)
```

Save as RDS and CSV
```{r}
saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
write.csv(copies_final,file= here("data","processed_data","total_clean_copies.csv"))

```

### 2023-02-12 code/processing_code/vp_copies_calculator.qmd
Total vib
Do not need to manually add in raw data

@@
 *This dataset is Daily Vp qPCR*
-
-*This dataset is Daily Vv qPCR*
 
@@
    mutate(filter_amt = case_when(Sample == "BO_7"~"100",
+                          Sample == "BO_7"~"100",
                           Sample == "BO_8"~"100",
-                          Sample == "BO_9"~"100",
                           Sample == "BO_9"~"100",
@@
 date<- replace %>%
-  mutate(day = substr(replace$Sample,4,5)) %>% #Creates new column `date` 
+  mutate(day = substr(replace$Sample,4,5)) %>% #Creates new column `day` 
   mutate(date = case_when(day == "7"~"2022-07-07",

### 2023-02-12 code/analysis_code/exploratory_vv.qmd
Minor edits

@@
     axis.text.x = element_text(angle = 45, hjust = 1),
-    plot.title = element_text(hjust = 0.5))
+    plot.title = element_text(hjust = 0.5),
+    axis.text.y = element_text(size = 15))
 
@@
     axis.text.x = element_text(angle = 45, hjust = 1),
-    plot.title = element_text(hjust = 0.5))
+    plot.title = element_text(hjust = 0.5),
+    axis.text.y = element_text(size = 15))
 
 all_redo
+```
+
+Compare Trial 1 and 2
+```{r, fig.height=7, fig.width=7}
+figure2<- ggarrange(all, all_redo,
+                    ncol = 1, nrow = 2)
+figure2
 ```

### 2023-02-16 code/analysis_code/dust.qmd
Plots for symposium

---
title: "Exploratory Dust"
format:
  html:
    theme: default
---
Library
```{r}
library(here)
library(tidyverse)
library(lubridate)
```

Load Data that has just DUST CONC
```{r}
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))

```

New Dataframes
```{r}
#Daily Dataframe with Sum dust
daily_dust_sum<- dust_sum %>%
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

#Dataframe with just month dates
dust_month <- dust_sum %>%
  filter(date %in% as.Date(c("2022-02-16", "2022-03-09", "2022-04-20", "2022-05-11", "2022-06-14", "2022-07-07", "2022-08-09", "2022-09-13", "2022-10-18", "2022-11-15", "2022-12-13")))
```

Plot Daily Dust Conc
```{r}
dust_sum_daily<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust))

dust_sum_daily
```

Plot Total Dust Conc
```{r}
dust_sum_2022<- dust_sum %>% ggplot() +geom_line(
 aes(x = date,
      y = dust),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_vline(aes(xintercept =as.Date("2022-07-07")), linetype = "dotdash", color = "4c848c")+
  geom_vline(aes(xintercept =as.Date("2022-07-19")), linetype = "dotdash", color = "4c848c")+
  scale_x_date(date_breaks = "months" , date_labels = "%b") +

#Plot Month Sample Dates over
geom_point(
  data=dust_month,
  aes(x = date,
      y = dust),
  color = "4c848c",
  size = 3,
  shape = 4,
  stroke = 1)

dust_sum_2022
```

Save as PNG
```{r}
png(file = here("results","plots", "dust_sum_daily.png"))
dust_sum_daily
dev.off()

png(file = here("results","plots", "dust_sum_2022.png"))
dust_sum_2022
dev.off()
```

### 2023-02-16 code/analysis_code/dust_and_copies.qmd
Plots for symposium

---
title: "Dust_Copies_Plot"
format:
  html:
    theme: default
---

Library
```{r}
library(here)
library(tidyverse)
library(ggpubr)
```

Load Data that has just DUST CONC
```{r}
#Summed Dust Concentration per day
dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) 
#Dust Conc at t = 0
#dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
#Dust Conc at t = 6
#dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
#Dust Conc at t = 12
#dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
#Dust Conc at t = 18
#dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))
#Copies 
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
```

Daily Dataframe with Summed dust by day
```{r}
daily_dust_sum<- dust_sum %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10

copies_d <- copies %>%
  filter(Type %in% "Daily") %>%
  mutate(site = recode(site, "BO" = "Blind Oso",
                             "C2" = "Canals",
                            "RD" = "Gulf"))

```

Plot Daily Dust Conc
```{r}
dust_fig<- daily_dust_sum %>% ggplot() +geom_line(
  aes(x = date,
      y = dust),
  color = "#935d2e",
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Dust Aerosol Optical Density")+
  theme(
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10)
  )

dust_fig
```

Total Daily Vibrio
```{r}
tot_vib<- copies_d %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites",
    color = "  Site") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),  #center title
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    legend.position = c(0.1, 0.8),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
  scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

tot_vib
```

Combine Dust and Copies
```{r, fig.height = 7, fig.width= 7}
fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

fig
```

### 2023-02-16 code/analysis_code/monthly.qmd
Plots for symposium

---
title: "Exploratory_monthly"
format:
  html:
    theme: default
---

Library
```{r}
library(tidyverse)
library(here)
library(lubridate)
```

Load Data
```{r}
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
```

Make mini dataframes for each site
```{r}
copies<- copies %>%
  mutate(site= recode(site, "BO" = "Blind Oso",
                       "C2" = "Canals",
                       "RD" = "Gulf"))

#Monthly All sites: Includes 3 Sample events in May 
monthly_may<- copies %>%
  mutate(monthly = case_when(date == "2022-07-07" ~ "Monthly", #This adds a July date to be included in monthly analysis
                             Type == "Monthly" ~ "Monthly")) %>%
  select(date,copies_mL,site,monthly) %>%
  filter(monthly %in% "Monthly")

#Monthly All sites: Includes ONE sample from May 
monthly<- monthly_may %>%
  filter(!date %in% as.Date(c("2022-05-20", "2022-05-24")))

#Monthly by Site
bo_monthly <- monthly %>% filter(site %in% "Blind Oso") #Blind Oso Bay
c2_monthly <- monthly %>% filter(site %in% "Canals") #Canals
rd_monthly <- monthly %>% filter(site %in% "Gulf") #Gulf Site

```

Total Monthly Vibrio at Gulf
```{r}
rd_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "#4c848c",
    size = 1) +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Gulf Site") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio at Canals
```{r}
c2_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "#ba4716",
    size = 1) +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Canals") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio at Blind Oso
```{r}
bo_monthly %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "#809C52",
    size =1) +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))
```

Total Monthly Vibrio
```{r}
tot_vib_allsites_monthly <- monthly %>%
  ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site),
  size = 1)+
  facet_wrap(~site) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Monthly Vibrio at all Sites") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 15),  #center title
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    legend.position = "none")+ #Removes legend 
scale_color_manual(values = c("Blind Oso" = "#809C52",
                              "Canals" = "#ba4716",
                              "Gulf" = "4c848c"))
tot_vib_allsites_monthly
```
Save as PNG
```{r}
png(file = here("results","plots", "tot_vib_allsites_monthly.png"))
tot_vib_allsites_monthly
dev.off()
```

### 2023-02-16 code/analysis_code/vp.qmd
Plots for symposium

--- 2023-02-09 code/analysis_code/exploratory_vp.qmd
@@
 ```{r}
-vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies_date.csv")) #This file had dates manually added
-sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) #Manually replaced NaN with 0 for no amplifcation 
+vp<- read_csv(file = here("data", "processed_data", "vp_clean_copies.csv")) #This file had dates manually added
+#sq <-read_csv(here("data", "raw_data", "vp_qPCR.csv")) #Manually replaced NaN with 0 for no amplifcation 
 ```
@@
 ```{r}
-std<- sq %>%
-  filter(
-    Fluor %in% "HEX",
-    Content %in% "Std")
-
-std %>%
-  ggplot() + geom_point(
-    aes(x = log10(SQ),
-        y = Cq))
-
+#std<- sq %>%
+  #filter(
+    #Fluor %in% "HEX",
+    #Content %in% "Std")
 ```

### 2023-02-16 code/analysis_code/vv.qmd
Plots for symposium

---
title: "Vv explore"
format:
  html:
    theme: default
---

Libraries
```{r}
library(tidyverse)
library(here)
library(ggpubr)
```

Data
```{r}
vv<- read_csv(file = here("data", "processed_data", "vv_clean_copies.csv")) 
#sq <-read_csv(here("data", "raw_data", "vv_qPCR_daily_0.csv"))
redo<- read_csv(here("data", "processed_data", "vv_clean_copies_redo.csv"))
t3<-  read_csv(here("data", "processed_data", "vv_clean_copies_trial3.csv"))
```

Vv Standard Curve
```{r}
#std<- sq %>%
 # filter(Content %in% "Std",
        # Fluor %in% "HEX") #%>%
  #ggplot() + geom_point(
  #  aes(x = log(SQ),
      #  y = Cq))
#std
```

Make Site df
```{r}
rd6<- t3 %>%
  filter(site %in% "RD")

c2<- t3 %>%
  filter(site %in% "C2")

bo<- t3 %>%
  filter(site %in% "BO")
```

Plot Vv at C2
```{r}
c2_vv<- c2 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkred") +
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkred")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Canals") +
  theme(plot.title = element_text(hjust = 0.5))

c2_vv
```

Plot Vv at RD
```{r}
rd_vv<- rd6 %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkblue")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkblue")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Gulf") +
  theme(plot.title = element_text(hjust = 0.5))

rd_vv
```

Plot Vv at BO
```{r}
bo_vv<- bo %>% ggplot() + geom_line(
  aes(x = date,
      y = copies_mL),
  color = "darkgreen")+
  geom_point(
    aes(x = date,
      y = copies_mL),
  color = "darkgreen")+
  theme_bw() +
  labs(x = "Date",
       y = "Copies/mL",
       title = "Vv at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))

bo_vv
```
Stack
```{r}
figure<- ggarrange(c2_vv, bo_vv, rd_vv + font("x.text", size = 10),
                    ncol = 1, nrow = 3) #Final plot is 1 column and 2 rows

figure
```

All sites Trial 1
```{r}
all<- vv %>% 
  mutate(site = recode(site, "BO" = "Blind Oso",
                              "C2" = "Canals",
                              "RD" = "Gulf")) %>%
  ggplot() + geom_line(
    aes(x = date,
        y = copies_mL,
        color = site)) + 
  scale_color_manual( values = c("Blind Oso" = "darkgreen",
                                "Canals" = "darkred",
                                "Gulf" = "darkblue"))+
  theme_bw()+
  facet_wrap(~site)+
  labs(
    x = "Date", 
    y = "Copies/mL",
    title = "Vibrio vulnificus Levels Across Dust Time Series: Trial 1")+
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 15))

all
```

All Site Trial 2
```{r}
all_redo<- redo %>% 
  mutate(site = recode(site, "BO" = "Blind Oso",
                              "C2" = "Canals",
                              "RD" = "Gulf")) %>%
  ggplot() + geom_line(
    aes(x = date,
        y = copies_mL,
        color = site)) + 
  scale_color_manual( values = c("Blind Oso" = "darkgreen",
                                "Canals" = "darkred",
                                "Gulf" = "darkblue"))+
  theme_bw()+
  facet_wrap(~site)+
  labs(
    x = "Date", 
    y = "Copies/mL",
    title = "Vibrio vulnificus Levels Across Dust Time Series: Trial 2")+
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 15))

all_redo
```

All Sites Trial 3
```{r}
all_t3<- t3 %>% 
  mutate(site = recode(site, "BO" = "Blind Oso",
                              "C2" = "Canals",
                              "RD" = "Gulf")) %>%
  ggplot() + geom_line(
    aes(x = date,
        y = copies_mL,
        color = site)) + 
  scale_color_manual( values = c("Blind Oso" = "darkgreen",
                                "Canals" = "darkred",
                                "Gulf" = "darkblue"))+
  theme_bw()+
  facet_wrap(~site)+
  labs(
    x = "Date", 
    y = "Copies/mL",
    title = "Vibrio vulnificus Levels Across Dust Time Series: Trial 3")+
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 15))

all_t3
```

Compare Trial 1 and 2
```{r, fig.height=7, fig.width=7}
figure2<- ggarrange(all, all_redo, all_t3,
                    ncol = 1, nrow = 3)
figure2
```

Save Vv at all sites as PNG
```{r}
png(file = here("results", "plots", "vv_all_sites.png")) #Saving c2
all
dev.off()

png(file = here("results", "plots", "vv_all_sites_redo.png")) #Saving c2
all_redo
dev.off()

```

### 2023-02-16 code/wrangling_code/dust_conc_processing.qmd
Plots for symposium

---
title: "Dust Concentration"
format:
  html:
    theme: default
---
Library
```{r}
library(tidyverse)
library(here)
```

Load Data
```{r}
nrl<- read_csv(here("data","raw_data","nrl_conc.csv"))
copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv"))
```

Make new dataframes  for Date and Dust for Exploratory 
```{r}
dust_tot<- nrl %>%  #Dust data for all hours
  select(dust,date,hr)

dust_0<- dust_tot %>% #Dust data at t = 0
  filter(hr %in% "0")

dust_6<-dust_tot %>% #Dust data at t = 6
  filter(hr %in% "6")

dust_12<- dust_tot %>% #Dust data at t =12
  filter(hr %in% "12")

dust_18<- dust_tot %>% #Dust data at t = 18
  filter(hr %in% "18")

dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>% #could also use groupby
  rename("dust" = "x")
#dust_sum<- dust_sum %>%
 #left_join(dust_sum)
```

Make dataframe with Dates, Copies, and Concentration For Stats anaylsis
```{r}
copies2 <- copies %>% select(
  date, copies_mL, site)

dust_sum_copies<- dust_sum %>% left_join(copies2) #Takes df on right of pipe and joins to the df on the left of pipe

dust_t0 <- dust_0 %>% right_join(copies2)

dust_t6 <- dust_6 %>% right_join(copies2)

dust_t12<- dust_12 %>% right_join(copies2)

dust_t18<- dust_18 %>% right_join(copies2)

```

Save Dataframes as RDS for STATS (Copies and Dust)
```{r}
saveRDS(dust_sum_copies, file = here("data", "processed_data", "dust_sum_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t0, file = here("data", "processed_data", "dust_0_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t6, file = here("data", "processed_data", "dust_6_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t12, file = here("data", "processed_data", "dust_12_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_t18, file = here("data", "processed_data", "dust_18_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

```

Save Dataframes as RDS for EXPLORE (Dust)
```{r}
saveRDS(dust_sum, file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with dust concentration

saveRDS(dust_0, file = here("data", "processed_data", "dust_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_6, file = here("data", "processed_data", "dust_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_12, file = here("data", "processed_data", "dust_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)

saveRDS(dust_18, file = here("data", "processed_data", "dust_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
```

### 2023-02-16 code/wrangling_code/vv_copies_calclulator.qmd
Plots for symposium

--- 2023-02-10 code/processing_code/vv_copies_calclulator.qmd
@@
 sq2<- sq %>% 
-  select(filter_amt,SQ,Cq, Sample, Fluor, dilution) %>%
+  select(SQ,Cq, Sample, Fluor) %>%
   filter(Fluor %in% "HEX")
@@
   filter(Fluor %in% "HEX",
-         !Well %in% "K19")
+        !Well %in% "K19") #This triplicate amplified
 
@@
    mutate(
-   log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+   log_sq = ((Cq-37.606)/-3.409), #Equation of the line to get log(SQ). Equation is from Oyster STD Curve.
     sq = 10^log_sq, #Unlog 
@@
     mutate(
-    log_sq = ((Cq-36.497)/-3.166), #Equation of the line to get log(SQ)
+    log_sq = ((Cq-37.606)/-3.409), #Equation of the line to get log(SQ). Equation is from Oyster STD CUrve.
     sq = 10^log_sq, #Unlog 
@@
 ```
+
+Cimpare Cq between OG and Redo
+```{r}
+#Deop NAs from +/- and lok at HEX
+sq0<- sq %>%
+  filter(
+    !is.na(Sample),
+    Fluor %in% "HEX")
+
+redo0<- redo %>%
+  filter(!is.na(Sample),
+         Fluor %in% "HEX")
+
+#copies0<- copies %>%
+ # filter(!is.na(Sample))
+
+#copies_redo0<- copies_redo %>%
+ # filter(!is.na(Sample))
+
+#Merge df
+cq_comp<- 
+  cbind(redo0, Cq_og= sq0$Cq) %>%
+  #cbind(copies_redo0, copies_redo1 = sq0$copies_mL) %>%
+ # cbind(copies0, copies_og1 = copies0$copies_mL) %>%
+  select(Sample, Cq, Cq_og) %>%
+  rename("Cq_redo" = "Cq")
+
+```

### 2023-02-24 code/analysis_code/analysis_README.Rmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

This folder contains code for analyzing cleaned data sets on Tempertature, Salinity, Dust, and Vibrio counts. `daily.qmd` assesses Vibrio counts over the daily time series, `dust.qmd` pertains to the dust concentrations sourced from NRL, `dust_and_copies.qmd` look at the former two data sets in tandem, and `env_par.qmd` looks at temperature and salinity.

### 2023-02-24 code/analysis_code/daily.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

--- 2023-02-09 code/analysis_code/exploratory_daily.qmd
@@
----
+zx---
 title: "Exploratory"
@@
 ```{r}
-copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv")) #This is the clean_data.csv from copies_calculator but with date manually added in,
+copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
 ```

### 2023-02-24 code/analysis_code/dust.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

@@
 
-Load Data that has just DUST CONC
+Load Data that has just dust concentrations
 ```{r}
-dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
-dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
-dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
-dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
-dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))
+dust_sum<-readRDS(file = here("data", "processed_data", "dust_conc","dust_sum.rds")) #Dataframe with copies_mL and dust concentration for 2022
+#We will primary use the dust_sum data for exploratory analysis
+
+dust_0 <- readRDS(here("data", "processed_data", "dust_conc", "dust_0.rds"))
+dust_6 <- readRDS(here("data", "processed_data", "dust_conc", "dust_6.rds"))
+dust_12 <- readRDS(here("data", "processed_data", "dust_conc", "dust_12.rds"))
+dust_18 <- readRDS(here("data", "processed_data", "dust_conc", "dust_18.rds"))
 
@@
 
-New Dataframes
+New Data frame for Daily time series 
 ```{r}
@@
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
-
-#Dataframe with just month dates
-dust_month <- dust_sum %>%
-  filter(date %in% as.Date(c("2022-02-16", "2022-03-09", "2022-04-20", "2022-05-11", "2022-06-14", "2022-07-07", "2022-08-09", "2022-09-13", "2022-10-18", "2022-11-15", "2022-12-13")))
 ```
 
-Plot Daily Dust Conc
+Plot Daily Dust Concentration over Daily Time Series 
 ```{r}
@@
 
-Plot Total Dust Conc
-```{r}
-dust_sum_2022<- dust_sum %>% ggplot() +geom_line(
- aes(x = date,
-      y = dust),
-  color = "#935d2e",
-  size = 1)+
-  theme_bw()+
-  labs(
-    x = "Date",
-    y = "Dust Aerosol Optical Density")+
-  theme(
-    axis.title.y = element_text(size = 15),
-    axis.text.y = element_text(size = 10),
-    axis.text.x = element_text(angle = 45, hjust = 1))+
-  geom_vline(aes(xintercept =as.Date("2022-07-07")), linetype = "dotdash", color = "4c848c")+
-  geom_vline(aes(xintercept =as.Date("2022-07-19")), linetype = "dotdash", color = "4c848c")+
-  scale_x_date(date_breaks = "months" , date_labels = "%b") +
-
-#Plot Month Sample Dates over
-geom_point(
-  data=dust_month,
-  aes(x = date,
-      y = dust),
-  color = "4c848c",
-  size = 3,
-  shape = 4,
-  stroke = 1)
-
-dust_sum_2022
-```
-
 Save as PNG
@@
 dev.off()
-
-png(file = here("results","plots", "dust_sum_2022.png"))
-dust_sum_2022
-dev.off()
 ```

### 2023-02-24 code/analysis_code/dust_and_copies.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

@@
 #Summed Dust Concentration per day
-dust_sum<-readRDS(file = here("data", "processed_data", "dust_sum.rds")) 
-#Dust Conc at t = 0
-#dust_0 <- readRDS(here("data", "processed_data", "dust_0.rds"))
-#Dust Conc at t = 6
-#dust_6 <- readRDS(here("data", "processed_data", "dust_6.rds"))
-#Dust Conc at t = 12
-#dust_12 <- readRDS(here("data", "processed_data", "dust_12.rds"))
-#Dust Conc at t = 18
-#dust_18 <- readRDS(here("data", "processed_data", "dust_18.rds"))
-#Copies 
+dust_sum<-readRDS(file = here("data", "processed_data", "dust_copies", "dust_copies_sum.rds")) 
 copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
@@
 
-Daily Dataframe with Summed dust by day
+Daily Data frame with Summed dust by day
 ```{r}
 daily_dust_sum<- dust_sum %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10
+  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10. No sampling occured then.
 
@@
 
-Plot Daily Dust Conc
+Plot Daily Dust Concentration
 ```{r}
@@
     axis.title.y = element_text(size = 15),
-    axis.text.y = element_text(size = 10)
-  )
+    axis.text.y = element_text(size = 10),
+    axis.title.x = element_text(size = 15),
+    axis.text.x = element_text(size = 10))
 
@@
 Combine Dust and Copies
-```{r, fig.height = 7, fig.width= 7}
+```{r, fig.height = 4, fig.width= 4}
 fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")
@@
 ```
+
+Save as PNG
+```{r}
+png(file = here("results", "plots", "vib_dust_daily.png"))
+fig
+dev.off()
+```

### 2023-02-24 code/analysis_code/env_par.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

---
title: "Environmental Parameters"
format:
  html:
    theme: default
---
*This data set contains environmental parameters: Temp and Salinity. I am waiting on an updated Nutrient data set*

Libraries
```{r}
library(tidyverse)
library(here)
library(ggpubr)
```

Load Data
```{r}
d1<- read_csv(here("data", "raw_data", "ysi_.csv"))
```

Daily data frame
```{r}
d1<- d1 %>%
  mutate(site = recode(site, "BO" = "Blind Oso",
                             "C2" = "Canals",
                            "RD" = "Gulf"))
daily<- d1 %>%
  filter(type %in% "Daily")

```

Plotting Daily Temperature 
```{r}
temp<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = temp,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Temperature (°C)",
    title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    legend.position = c(0.1,0.2),
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

temp

```

Plotting Daily Salinity
```{r}
sal<- daily %>% ggplot() + geom_line(
  aes(
    x = date,
    y = salinity,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Salinity",
    title = "Salinity across Daily Time Series by Site",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))

sal
```
Combine Temperature and Salinity Plots 
```{r, fig.width= 4, fig.height=4}
fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

fig
```

Save as PNG
```{r}
png(file = here("results", "plots", "temp_sal.png"))
fig
dev.off()
```

### 2023-02-24 code/wrangling_code/dust_conc_processing.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

@@
 ---
+This data set is sourced from the Naval Research Lab. Aerosol optical thickness (AOT) was measures at 4 time points per day (t = 0,6,12,18).
+
 Library
@@
 ```{r}
-nrl<- read_csv(here("data","raw_data","nrl_conc.csv"))
-copies<- read_csv(here("data", "processed_data", "clean_copies_date.csv"))
+nrl<- read_csv(here("data","raw_data","nrl_conc.csv")) #Dust concentration data set
+copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Vibrio counts data set from qPCR
 ```
 
-Make new dataframes  for Date and Dust for Exploratory 
+Make new data frames for each time point as well as sum AOT per day
 ```{r}
@@
 
-dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>% #could also use groupby
+dust_sum<- aggregate(dust_tot$dust, by=list(date=dust_tot$date),FUN=sum) %>% #Sum of AOT by day
   rename("dust" = "x")
-#dust_sum<- dust_sum %>%
- #left_join(dust_sum)
+
 ```
 
-Make dataframe with Dates, Copies, and Concentration For Stats anaylsis
+Make dataframe with Dates, Copies/mL, and Dust Concentration For Stats analysis. Here we are combining dust data with copies/mL data. 
 ```{r}
+#Condense copies dateset
 copies2 <- copies %>% select(
@@
 
-dust_sum_copies<- dust_sum %>% left_join(copies2) #Takes df on right of pipe and joins to the df on the left of pipe
+#Takes df on right of pipe and joins to the df on the left of pipe 
+dust_copies_sum<- dust_sum %>% left_join(copies2) #Dust at all time points
+dust_copies_0 <- dust_0 %>% right_join(copies2) #T0
+dust_copies_6 <- dust_6 %>% right_join(copies2) #T6
+dust_copies_12<- dust_12 %>% right_join(copies2) #T12
+dust_copies_18<- dust_18 %>% right_join(copies2) #T18
+```
 
-dust_t0 <- dust_0 %>% right_join(copies2)
+Save **dust/copies** dataframes as RDS for Stats
+```{r}
+saveRDS(dust_copies_sum, file = here("data", "processed_data", "dust_copies", "dust_copies_sum.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-dust_t6 <- dust_6 %>% right_join(copies2)
+saveRDS(dust_copies_0, file = here("data", "processed_data",  "dust_copies","dust_copies_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-dust_t12<- dust_12 %>% right_join(copies2)
+saveRDS(dust_copies_6, file = here("data", "processed_data", "dust_copies", "dust_copies_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-dust_t18<- dust_18 %>% right_join(copies2)
+saveRDS(dust_copies_12, file = here("data", "processed_data",  "dust_copies","dust_copies_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
+
+saveRDS(dust_copies_18, file = here("data", "processed_data",  "dust_copies","dust_copies_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
@@
 
-Save Dataframes as RDS for STATS (Copies and Dust)
+Save **Dust Concentration** data frames as RDS (good for Dust plots)
 ```{r}
-saveRDS(dust_sum_copies, file = here("data", "processed_data", "dust_sum_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
+saveRDS(dust_sum, file = here("data", "processed_data", "dust_conc", "dust_sum.rds")) #Dataframe with dust concentration
 
-saveRDS(dust_t0, file = here("data", "processed_data", "dust_0_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
+saveRDS(dust_0, file = here("data", "processed_data","dust_conc", "dust_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-saveRDS(dust_t6, file = here("data", "processed_data", "dust_6_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
+saveRDS(dust_6, file = here("data", "processed_data","dust_conc", "dust_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-saveRDS(dust_t12, file = here("data", "processed_data", "dust_12_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
+saveRDS(dust_12, file = here("data", "processed_data","dust_conc", "dust_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 
-saveRDS(dust_t18, file = here("data", "processed_data", "dust_18_copies.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
-
+saveRDS(dust_18, file = here("data", "processed_data","dust_conc", "dust_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
 ```
-
-Save Dataframes as RDS for EXPLORE (Dust)
-```{r}
-saveRDS(dust_sum, file = here("data", "processed_data", "dust_sum.rds")) #Dataframe with dust concentration
-
-saveRDS(dust_0, file = here("data", "processed_data", "dust_0.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
-
-saveRDS(dust_6, file = here("data", "processed_data", "dust_6.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
-
-saveRDS(dust_12, file = here("data", "processed_data", "dust_12.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
-
-saveRDS(dust_18, file = here("data", "processed_data", "dust_18.rds")) #Dataframe with copuies_mL AND dust concentration (for stat analysis)
-```

### 2023-02-24 code/wrangling_code/total_copies_calculator.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

--- 2023-02-12 code/processing_code/total_copies_calculator.qmd
@@
 ---
-*This script takes Cq value outputs from the CFX and computes it to copies of gene targets of Vibrio*
+*This script is for cleaning qPCR results on Total Vibrio enumeration.
+It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio.
+The data set contains a daily and monthly time series.
+We will be examining daily.*
 
@@
 
-Add Dilution Column and Filter column
+Add Dilution Column and Filter column.
+The dilution column applies if samples were diluted prior to qPCR.
+For example, if I diluted a sample 1:10, my dilution column would denote "10".
+The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter prior to qPCR.
 ```{r}
@@
   mutate(dilution = case_when(
-                          Sample == "BO7"~"10",
+                          Sample == "BO7"~"10", #BO site was diluted 1:10
                           Sample == "BO8"~"10",
@@
                           Sample == "BO19"~"10",
-                              site  == "C2"~ "1",
+                              site  == "C2"~ "1", #Canals and Gulf were not diluted
                               site == "RD"~"1"), 
@@
 filter_daily<- dilution_daily %>%
-   mutate(filter_amt = case_when(Sample == "BO_7"~"100",
+   mutate(filter_amt = case_when(Sample == "BO_7"~"100", #100mL concnetrated onto filter
                           Sample == "BO7"~"100",
@@
                           Sample == "BO15"~"50",
-                          Sample == "BO16REDO"~"30",
+                          Sample == "BO16REDO"~"30", #30 mL concentrated onto filter 
                           Sample == "BO17"~"30",
@@
     mutate(
-    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ)
+    log_sq = ((Cq-36.114)/-3.3844), #Equation of the line to get log(SQ):  This comes from running a standard curve during qPCR. 
     sq = 10^log_sq, #Unlog 
-    copies_rxn = sq*10, #SQ*Reaction Volume
-    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume
+    copies_rxn = sq*10, #SQ*Reaction Volume: Each qPCR reaction was 10µL
+    copies_uL_1_10 = copies_rxn/2.5, #Copies/RXN / DNA Template Volume: Each 10µL reaction had 2.5µL of DNA added. 
     copies_uL_extract = copies_uL_1_10 *dilution, #(Copies/uL of 1:10)*(1 if there was no sample dilution // *10 if there was a 1:10 dilution etc.)
@@
 
-Replace NaN with 0 for copies/mL
+Some samples did not amplify: Replace NaN with 0 for copies/mL
 ```{r}
@@
 
-Date Column
+Date Column: This adds a date correpsonding the a number in the sample name. 
 ```{r}
@@
 
-Take the average of qPCR triplicates 
+Each sample (Date) was done in triplicate for qPCR: We need to take the average of qPCR triplicates 
 ```{r}
@@
 
-Combine Daily and Monthly 
+Combine Daily and Monthly data sets back togthether. 
 ```{r}
@@
 ```{r}
-saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
+#saveRDS(copies_final, file= here("data","processed_data","total_clean_copies.rds"))
 write.csv(copies_final,file= here("data","processed_data","total_clean_copies.csv"))

### 2023-02-24 code/wrangling_code/wrangling_README.Rmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

The folder contains various cleaning codes for data sets on environmental parameters (dust, temperature, salinity), and total Vibrio counts in sea surface waters.

### 2023-02-24 data/raw_data/rawdata_README.Rmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

This folder contains dust concentrations in AOT from the Naval Reserach lab, environmental parameters (temperature and salinity) that were collected during the summer at the three sites, and  qPCR results for Total Vibrio.

### 2023-02-24 products/manuscript/Manuscript.qmd
Proj Pt 2
- Cut out all TX Dust work- Made new Repo for that. 
- Annotated everything

@@
 subtitle: ""
-author: Andreas Handel
+author: Nathan Greenslit
 date: "`r Sys.Date()`"
@@
     highlight-style: github
-bibliography: ../dataanalysis_template_references.bib
-csl: ../apa.csl
 ---
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
 
@@
 # Summary/Abstract
-_Write a summary of your project._
 
@@
 
-## Description of data and data source
+## Questions/Hypotheses to be addressed
+
+How do Saharan dust events influence Vibrio populations in high nutrient coastal waters?
+
+*What I am looking for* I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels.
+If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response.
+Nutrients, dust input, salinity, and temperature will be the primary factors of interest.
+
+*How I will analyze it* - Look at Vibrio growth over time series (estimated from qPCR) - Examine relationships between dust input and Vibrio growth - Examine influence of site-specific environmental parameters (Temperature, Salinity, Nutrients) on growth response - Run linear models on factors like: dust x growth and nutrients x growth - NMDS to see which parameters have the strongest influence on growth
+
+{{< pagebreak >}}
+
+# Methods
+
+## Data aquisition
 
@@
 
-## Questions/Hypotheses to be addressed
+## Data import and cleaning
 
-How do Saharan dust events influence Vibrio populations in high nutrient coastal waters?
+This project works with three primary data sets.
+(1) Enumeration of Vibrio bacteria as copies/mL from quantitative PCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi\_.csv).
+Below is a brief summary of how each data set was cleaned.
+More details can be found in the supplementary files, containing the code and comments describing what each line does.
+Please refer to the `project_README.md` file located in the NathanGreenslit-MADA-Project folder for details on the repository contents and  instructions on reproducibility
 
-*What I am looking for* 
-I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels.
-If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response.
-Nutrients, dust input, salinity, and temperature will be the primary factors of interest.
+### Cleaning of Total Vibrio Enumerations Data set
 
-*How I will analyze it* 
-- Look at Vibrio growth over time series (estimated from qPCR) 
-- Examine relationships between dust input and Vibrio growth 
-- Examine influence of site-specific environmental parameters (Temperature, Salinity, Nutrients) on growth response 
-- Run linear models on factors like: dust x growth and nutrients x growth 
-- NMDS to see which parameters have the strongest influence on growth
+Quantitative PCR provides enumeration results as "cycle threshold or Cq" values.
+This depicts the cycle number in which enough of the DNA target was detectable to be amplified.
+Lower Cq values correspond to higher target concentrations (as it takes less cycles to amplify), and vice verse for higher Cq values.
+The cleaning of this data set consisted of converting these Cq values into something that can be used to quantify Vibrio (copies of target per mL of seawater).
+To do so, I needed to take into account the amount of water sample that was concentrated on a filter, the total amount of DNA that was eluted following a DNA extraction, and the amount of DNA template added to each mix for qPCR (to name a few).
+Taking these into account, we are able to calculate our way from Cq values to copies of Vibrio per mL of seawater.
+qPCR was conducted in triplicate, so the last step was to take the average of the three replicates to have a final value per sample.
+A date column was also added for each sample.
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key.
-Then you can include like this:
+### Cleaning of Dust Concentration (AOT) Data set
 
-{{< pagebreak >}}
+Dust concentration (AOT) was collected at time points 0 hour, 6 hour, 12 hour, and 18 hour.
+In this script, I made different data sets based on time points as well as a data set containing the summed dust concentration (of all time points) per day.
+For downstream statistical analysis, it will be helpful to have these different data sets to compare to time of sample collection.
 
-# Methods
+### Cleaning of Temperature and Salinity Data set
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-
-## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
-
-## Statistical analysis
-_Explain anything related to your statistical analyses._
+This data set did not require any cleaning.
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
+### Dust
 
-@tbl-summarytable shows a summary of the data.
+![](images/dust_sum_daily.png)
 
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
+This figure depicts the summed dust concentration (time points 0hr-18hr) across the daily time series.
+During the time series, two small period of higher dust AOT occurred on the 9th and 13th respectively.
+A much larger spike in AOT was observed on the 16th.
 
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/summarytable.rds")
-knitr::kable(resulttable)
-```
+### Daily Vibrio
+
+![](images/vib_dust_daily.png)
+
+The top portion of the figure depicts total enumerated Vibrio as copies per mL, with color by site.
+There is a noticeable shift in copies in the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
+The bottom portion of the figure depicts the dust concentration (AOT) across the daily time series.
+Initial shift in copies per mL at the Guld and Blind Oso occur around initial dust introduction on the 13th, and experience another shift around the time of higher dust input on the 16th.
+
+### Temperature and Salinity
+
+![](images/temp_sal.png)
+
+Temperature and salinity at the Canal site is relatively high and consistent as compared to the other sites.
+Blind Oso Bay experiences some shifts in temperatures and drastic fluctuations in salinity.
+The Gulf site had a drop in temperature (possibly due to upwelling) and relatively consistent ocean salinities
 
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by sex."
-#| echo: FALSE
-knitr::include_graphics(here("results","height_weight_stratified.png"))
-```
-
 ## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
 
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
 
 ## Conclusions
-_What are the main take-home messages?_
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.
 

### 2023-03-14 code/analysis_code/dust/daily_dusts.qmd
Cleaned up code

---
title: "New Dust and Copies"
format:
  html:
    theme: default
---
**This code is useful for looking at variations of daily dust trends. For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average. We will be working with a sum that is 24hr prior to sample collection. This utilizes a data set that has the variations of dust trends **
Libs
```{r}
library(here)
library(tidyverse)
```

Data
```{r}
dust<- read_csv(here("data", "processed_data", "dust_master.csv")) 
```
**There are a few columns that we do not need or that contain NaN. So let's remove those.**

Clean
```{r}
dust<- dust %>%
  select(date,t1,t7,t13,t19,tsum,t7sum) %>%
  filter(!is.na(dust$date))
```

Plot T7sum
```{r}
dust %>% ggplot()+ geom_line(
  aes(x = date,
      y = t7sum))+
  labs(title = "Dust Concentration over Daily Time Series",
       x = "Date",
       y = "Dust Concentration (AOD)")+
  theme_bw()
```

### 2023-03-14 code/analysis_code/dust/dust_and_copies.qmd
Cleaned up code

--- 2023-02-24 code/analysis_code/dust_and_copies.qmd
@@
 ---
-
+**This script is good for visualizing copies per mL data and dust data.**
 Library
@@
 
-Load Data that has just DUST CONC
+Load Data 
 ```{r}
-#Summed Dust Concentration per day
-dust_sum<-readRDS(file = here("data", "processed_data", "dust_copies", "dust_copies_sum.rds")) 
+dust_sum<-read_csv(here("data", "processed_data", "dust_master.csv")) 
 copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) 
@@
 
-Daily Data frame with Summed dust by day
+Daily Data frames
 ```{r}
 daily_dust_sum<- dust_sum %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) #We do not have copies_mL for 2022-07-10. No sampling occured then.
+  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
 
@@
   aes(x = date,
-      y = dust),
+      y = t7sum),
   color = "#935d2e",

### 2023-03-14 code/analysis_code/env/env_par.qmd
Cleaned up code

--- 2023-02-24 code/analysis_code/env_par.qmd
@@
 ---
-*This data set contains environmental parameters: Temp and Salinity. I am waiting on an updated Nutrient data set*
+*This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.*
 
@@
 library(ggpubr)
+library(lubridate)
 ```
@@
 ```{r}
-d1<- read_csv(here("data", "raw_data", "ysi_.csv"))
+d1<- read_csv(here("data", "processed_data", "tx_master.csv"))
 ```
@@
 ```{r}
-d1<- d1 %>%
-  mutate(site = recode(site, "BO" = "Blind Oso",
-                             "C2" = "Canals",
-                            "RD" = "Gulf"))
 daily<- d1 %>%
-  filter(type %in% "Daily")
-
+  filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
+  select(!...1)
 ```
@@
     x = date,
-    y = salinity,
+    y = sal,
     color = site),

### 2023-03-14 code/analysis_code/stats/env_and_dust_stats.qmd
Cleaned up code

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---
### Libraries
```{r}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
```

# LOADING AND CLEANING DATA

### Data
```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

### Clean and Trim
```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-07"):as.Date("2022-07-19")) %>%
  select_if(~ !any(is.na(.))) #Removes columns with NA

data<- data %>%
  filter(date %in% as.Date("2022-07-07"):as.Date("2022-07-19"))
```

### Combine into one big data set
```{r}
master<- 
  dust %>% left_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analaysis. 
```

### Make Site DF
```{r}
bo<- master %>%
  filter(site %in% "Blind Oso")

c2<- master %>%
  filter(site %in% "Canals")

rd<- master %>%
  filter(site %in% "Gulf")
```

#DUST

**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input. This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date. We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**

### Cross-correlation:
```{r}
#Blind Oso
ccf(bo$t7sum, bo$copies_mL)

#Canals
ccf(c2$t7sum, c2$copies_mL)

#Gulf
ccf(rd$t7sum, rd$copies_mL) 
```
**From exploratory analysis, we know that the Gulf had a dramatic response, so let's look at that site first. Lag is greatest at +2 and -1. We need to choose +2 (where dust predicts copies). -1 would be the other way around, where copies predict dust. This is not logical.**

## Gulf Site Dust

### Adding a 2 day lag in dust:
```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = rd2)
tidy(lm_fit)
```

### Plot RD T7 Sum:
```{r}
rd2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD)")
```
## BO Site Dust

### Adding a 2 day lag in dust:
```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = bo2)
tidy(lm_fit)
```

### Plot BO T7 Sum:
```{r}
bo2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO)")
```
## C2 Site Dust

### Adding a 2 day lag in dust:
```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = c22)
tidy(lm_fit)
```

### Plot C2 T7 Sum:
```{r}
c22 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2)")
```

### 2023-03-14 code/wrangling_code/df_master.qmd
Cleaned up code

---
title: "ENV_COP_DUST"
format:
  html:
    theme: default
---
**This code combines clean data sets (dust, copies, and env)**

Library   
```{r}
library(tidyverse)
library(here)
```

Load Clean Data
```{r}
dust<- readRDS(here("data", "processed_data", "dust_clean.rds")) #Clean dust data
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Clean Copies Data
env<- read_csv(here("data", "processed_data", "ysi_2.csv")) #Clean Environmental Data
```

Clean Data
```{r}
env<- env %>%
select_if(~ !any(is.na(.))) #Drops Columns with NA
```

Combine 3 data sets to make a large master data set
```{r}
master<- dust %>% full_join(env) %>% full_join(copies)
```

Clean Data Again
```{r}
master<- master %>%
  select(!...1)  %>% #Gets rid of number column from excel
  mutate(
    sal = as.numeric(master$sal), #Set character columns as numeric
    temp = as.numeric(master$temp),
    do_mgl = as.numeric(master$do_mgl),
    do_per = as.numeric(master$do_per),
    pH = as.numeric(master$pH),
    secchi = as.numeric(master$secchi),
    amm = as.numeric(master$amm),
    nn = as.numeric(master$nn),
    orthop = as.numeric(master$orthop),
    sil = as.numeric(master$sil),
    din_dip = as.numeric(master$din_dip),
    tdn = as.numeric(master$tdn),
    don = as.numeric(master$don),
    doc = as.numeric(master$doc),
    toc = as.numeric(master$toc),
    tn = as.numeric(master$tn)) %>%
    mutate(site = recode(site, "BO" = "Blind Oso", #Rename Site Columns
                             "C2" = "Canals",
                            "RD" = "Gulf"))
```

Save as CSV 
```{r}
write.csv(master,file= here("data","processed_data","tx_master.csv"))

```

### 2023-03-14 data/processed_data/processed_data_README.Rmd
Cleaned up code

---

This folder contains processed data that has been cleaned from code in the `wrangling` folder.\
(1) `total_clean_copies`: Vibrio enumeration (as copies per mL) across time series.\
(2) `dust_clean.rds`: Cleaned dust data
(3) `ysi_2.csv`: Environmental parameter data
(4) `tx_master.csv` which is a compilation of all relevant data. 
(5)`dust_master.csv`: Manually created dust data examining different time points, sums, avereages (helpful for statistics)

### 2023-03-14 data/raw_data/rawdata_README.Rmd
Cleaned up code

@@
-This folder contains dust concentrations in AOT from the Naval Reserach lab, environmental parameters (temperature and salinity) that were collected during the summer at the three sites, and  qPCR results for Total Vibrio.
+This folder contains three raw data sets:\
+(1) `dust.csv`:dust concentrations in AOT from the Naval Research lab. Concentrations (AOT) were collected at 4 time points (UTC): T0,T6,T12, and T18.\
+(2) `copies_master.csv`:  qPCR results for Total Vibrio enumeration.

### 2023-03-14 products/manuscript/Manuscript.qmd
Cleaned up code

@@
 
-## Data aquisition
+## Data acquisition
 
@@
 
-Variables will include `Temperature (°C)`, `Salinity`, `Nitrate`, `Phosphate`, `Copies/mL` produced using quantitative PCR.
-Provides gene copies of Vibrio bacteria (I am hoping to also get data about specific species), and `Dust Concentration (AOT)` from the naval research laboratory satellite data.
+Variables will include `Temperature (°C)`, `Salinity`, `Nitrate`, `Phosphate`, `Copies/mL` produced using quantitative PCR, and `Dust Concentration (AOT)` from the naval research laboratory satellite data.
 
@@
 This project works with three primary data sets.
-(1) Enumeration of Vibrio bacteria as copies/mL from quantitative PCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi\_.csv).
+(1) Enumeration of Vibrio bacteria as copies/mL from quantitative PCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi_2.csv).
 Below is a brief summary of how each data set was cleaned.
@@
 Quantitative PCR provides enumeration results as "cycle threshold or Cq" values.
-This depicts the cycle number in which enough of the DNA target was detectable to be amplified.
+This depicts the cycle number in which enough of the DNA target was present to be amplified and thus detected.
 Lower Cq values correspond to higher target concentrations (as it takes less cycles to amplify), and vice verse for higher Cq values.
@@
 
-![](images/dust_sum_daily.png)
+![](images/paste-0FE55FAA.png)
 
-This figure depicts the summed dust concentration (time points 0hr-18hr) across the daily time series.
-During the time series, two small period of higher dust AOT occurred on the 9th and 13th respectively.
+This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series.
+During the time series, two small periods of higher dust AOT occurred on the 9th and 13th respectively.
 A much larger spike in AOT was observed on the 16th.
@@
 
-![](images/vib_dust_daily.png)
+![](images/paste-C33B6E43.png)
 
@@
 The bottom portion of the figure depicts the dust concentration (AOT) across the daily time series.
-Initial shift in copies per mL at the Guld and Blind Oso occur around initial dust introduction on the 13th, and experience another shift around the time of higher dust input on the 16th.
+Initial shift in copies per mL at the Gulf and Blind Oso occur 24-48hr following initial dust introduction on the 13th, and experience another shift around 24-48hr of higher dust input on the 16th.
 

### 2023-03-14 code/analysis_code/Old/statistical_analysis.R

@@ -1,56 +1,0 @@
-###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
-
-#load needed packages. make sure they are installed.
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed_data","processeddata.rds")
-
-#load data. 
-mydata <- readRDS(data_location)
-
-
-######################################
-#Data fitting/statistical analysis
-######################################
-
-############################
-#### First model fit
-# fit linear model using height as outcome, weight as predictor
-
-lmfit1 <- lm(Height ~ Weight, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable1 <- broom::tidy(lmfit1)
-
-#look at fit results
-print(lmtable1)
-
-# save fit results table  
-table_file1 = here("results", "resulttable1.rds")
-saveRDS(lmtable1, file = table_file1)
-
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and sex as predictor
-
-lmfit2 <- lm(Height ~ Weight + Sex, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
-
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
-
-  

### 2023-03-16 products/manuscript/Manuscript.qmd
stats

@@
 
+### Cross-correlation Analysis
+
+Previous literature has shown that Vibrio respond 12-24 hours following dust deposition events.
+Based on this, we need to look at lags between the two times series for dust data and growth data.
+To do this, we can set up a cross-correlation analysis, which allows us to examine the lag/lead relationship between the two variables.
+
+![](images/00003f.png){width="475"}
+
+![](images/000043.png){width="473"}
+
+![](images/000047.png){width="473"}
+
+According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1.
+
+[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022.
+Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
+Therefore, these CCA were made with only this time series.
+This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
+But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days.
+Below we will run linear models that DO contain these prior dust days.
+And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
+But this ultiamtely means that the lag identified above.
+may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
+
+### Linear Regression Models
+
+![](images/00000f.png){width="382"}
+
+![](images/00000b.png){width="381"}
+
+![](images/000013.png){width="381"}
+
+Above are linear regressions that examine the relationship between dust concentration and copies_mL.
+As mentioned above, the lags identified may not correspond to the most significant relationship between the two variables.
+While a lag of -1 for Blind Oso and -2 for the Canals expalins the data well, a lag of -2 for the Gulf site was more fitting.
+This makes sense, as taking a look at the copies_mL plot above, we do see a \~2 day lag in response, where we have a spike in dust, and 2 days later, a spike in copies_mL.
+
 ## Full analysis

### 2023-03-16 code/analysis_code/stats/dust_copies_stats.qmd
stats

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---
### Libraries
```{r}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

# LOADING AND CLEANING DATA

### Data
```{r}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

### Clean and Trim
```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set
```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(copies_mL, t7sum, site, date) 
```

### Make Site DF
```{r}
bo<- master %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

c2<- master %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

rd<- master %>%
  filter(site %in% "Gulf"| date %in% c(as.Date("2022-07-05"), as.Date("2022-07-06")))
```

#DUST

**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input. This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date. We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**

### Cross-correlation:
```{r}
#Blind Oso
bo_ccf<- bo %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
title("Blind Oso DustxCopies CCA", line =1)

#Canals
c2_ccf<- c2 %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
title("Canals DustxCopies CCA", line=1)

#Gulf
rd_ccf<- rd %>%
  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
title("Gulf DustxCopies CCA", line =1)
```
**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1 **

## Gulf Site Dust

### Adding a 2 day lag in dust:
```{r}
rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = rd2)
tidy(lm_fit)
```

### Plot RD T7 Sum:
```{r}
rd2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)")
```
## BO Site Dust

### Adding a 2 day lag in dust:
```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = bo2)
tidy(lm_fit)
```

### Plot BO T7 Sum:
```{r}
bo2 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)")
```
## C2 Site Dust

### Adding a 2 day lag in dust:
```{r}
c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) #Make lag from ccf
```

### Create a linear Model: 
```{r}
lm_spec<- linear_reg()
lm_fit<- fit(lm_spec, copies_mL~lag, data = c22)
tidy(lm_fit)
```

### Plot C2 T7 Sum:
```{r}
c22 %>% ggplot() + geom_point(
  aes(x = lag, 
      y = copies_mL)) +
  geom_smooth(aes(x = lag,
                  y = copies_mL),
              method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)")
```

_______________________
All Variables

New data set
```{r}
master2<- dust0 %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>%
  select(!c(t1, t7, t13, t19, hr_cst, tsum, t7sum, ...1, SUM, dust,  Type, date )) %>%
  na.omit()
```
```{r}
bo_a<- master2 %>%
  filter(site %in% "Blind Oso") %>%
  select(!site)

c2_a<- master2 %>%
  filter(site %in% "Canals")%>%
  select(!site)

rd_a<- master2 %>%
  filter(site %in% "Gulf")%>%
  select(!site)
```

## BO: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_bo<- recipe(copies_mL ~ ., data = bo_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to train data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo_a)

#View as Tibble 
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_bo) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

## C2: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_c2<- recipe(copies_mL ~ ., data = c2_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to train data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c2_a)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_c2) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```
## RD: Using all Variables vs. Growth
```{r}
#Create Recipe for Growth and all other parameters
growth_recipe_rd<- recipe(copies_mL ~ ., data = rd_a)

#Set up linear model
lm_mod_all<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod_all) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to train data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd_a)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```
Plot
```{r}
tidy(growth_fit_rd) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```

### 2023-03-16 code/analysis_code/analysis_README.Rmd
editing for grammer and flow

@@
-This folder contains code for analyzing cleaned data sets on Tempertature, Salinity, Dust, and Vibrio counts.
-`daily.qmd` assesses Vibrio counts over the daily time series, `dust.qmd` pertains to the dust concentrations sourced from NRL, `dust_and_copies.qmd` look at the former two data sets in tandem, and `env_par.qmd` looks at temperature and salinity.
+This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
+(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
+(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
+(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
+(4) The stats folder contains `dust_copies_stats.qmd` that runs models on variables of interest. \

### 2023-03-16 code/wrangling_code/wrangling_README.Rmd
editing for grammer and flow

@@
-The folder contains various cleaning codes for data sets on environmental parameters (dust, temperature, salinity), and total Vibrio counts in sea surface waters.
+The folder contains various cleaning codes:\
+(1) `dust.qmd`: This code takes dust concentration data from NRL and cleans it.
+Steps include, switching to the correct timezone, creating concentration sums per day, as well as combining the dust data set with the vibrio enumeration data set.\
+(2) `total_copies_calculator.qmd`: This script is for cleaning qPCR results on Total Vibrio enumeration.
+It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio.
+The data set contains a daily and monthly time series.
+We will be examining daily.\
+(3) `df_master.qmd`: This script combines the three clean data sets for dust concentration, copies per mL, and environmental parameters

### 2023-03-16 data/processed_data/processed_data_README.Rmd
editing for grammer and flow

@@
 (2) `dust_clean.rds`: Cleaned dust data
-(3) `ysi_2.csv`: Environmental parameter data
+(3) `ysi_2.csv`: Environmental parameter data-- did not need cleaning
 (4) `tx_master.csv` which is a compilation of all relevant data. 
-(5)`dust_master.csv`: Manually created dust data examining different time points, sums, avereages (helpful for statistics)
+(5)`dust_master.csv`: Manually created dust data examining different dust permutations: time points, sums, avereages (helpful for statistics)

### 2023-03-17 code/analysis_code/stats/dust_copies_stats.qmd
Project Pt. 3: Final Edits

@@
 ---
+
 ### Libraries
+
 ```{r}
@@
 ### Data
+
 ```{r}
@@
 ### Clean and Trim
+
 ```{r}
@@
 ### Combine into one big data set
+
 ```{r}
@@
 ### Make Site DF
+
 ```{r}
@@
 
-#DUST
+# Uni-variate Models
 
@@
 
-### Cross-correlation:
+## Cross-correlation:
+
 ```{r}
@@
 ```
+
 **A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values.
@@
 In other words, dust occurs, and one day later, we see a response in copies_mL.
-C2 has a lag at -2, and RD at -1 **
+C2 has a lag at -2, and RD at -1**
 
-## Gulf Site Dust
+## Gulf Site:
 
-### Adding a 2 day lag in dust:
+#### Add two day lag
+
 ```{r}
 rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) #Make lag from ccf
+  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
 ```
 
-### Create a linear Model: 
+#### Create a linear model
+
 ```{r}
-lm_spec<- linear_reg()
-lm_fit<- fit(lm_spec, copies_mL~lag, data = rd2)
-tidy(lm_fit)
+#Create Recipe for Growth 
+growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)
+
+#Set up linear model
+lm_mod<- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+#Workflow that adds recipe to model
+Growth_wflow<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(growth_recipe_rd)
+
+#Use workflow to fit model to  data set
+growth_fit_rd<- Growth_wflow %>%
+  fit(data = rd2)
+
+#View as Tibble 
+growth_fit_rd %>%
+  extract_fit_parsnip() %>%
+  tidy()
 ```
 
-### Plot RD T7 Sum:
+### Use model to test data
+
 ```{r}
-rd2 %>% ggplot() + geom_point(
-  aes(x = lag, 
-      y = copies_mL)) +
-  geom_smooth(aes(x = lag,
-                  y = copies_mL),
-              method = "lm")+
+aug_test_rd <- augment(growth_fit_rd, rd2)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
+rd_metrics<- full_join(rmse, rsq)
+```
+
+### Plot
+
+```{r}
+aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
   stat_poly_eq(aes(x = lag, y = copies_mL),
@@
 ```
-## BO Site Dust
 
-### Adding a 2 day lag in dust:
+## Blind Oso Site:
+
+#### Add one day lag
+
 ```{r}
 bo2<- bo%>% mutate(
-  lag =Lag(bo$t7sum, shift  = 1)) #Make lag from ccf
+  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
 ```
 
-### Create a linear Model: 
+#### Create a linear model
+
 ```{r}
-lm_spec<- linear_reg()
-lm_fit<- fit(lm_spec, copies_mL~lag, data = bo2)
-tidy(lm_fit)
+#Create Recipe for Growth 
+growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)
+
+#Set up linear model
+lm_mod<- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+#Workflow that adds recipe to model
+Growth_wflow<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(growth_recipe_bo)
+
+#Use workflow to fit model to  data set
+growth_fit_bo<- Growth_wflow %>%
+  fit(data = bo2)
+
+#View as Tibble 
+growth_fit_bo %>%
+  extract_fit_parsnip() %>%
+  tidy()
 ```
 
-### Plot BO T7 Sum:
+### Use model to test data
+
 ```{r}
-bo2 %>% ggplot() + geom_point(
-  aes(x = lag, 
-      y = copies_mL)) +
-  geom_smooth(aes(x = lag,
-                  y = copies_mL),
-              method = "lm")+
+aug_test_bo <- augment(growth_fit_bo, bo2)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
+rd_metrics<- full_join(rmse, rsq)
+```
+
+### Plot
+
+```{r}
+aug_test_bo %>% ggplot(aes(copies_mL, .pred)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
   stat_poly_eq(aes(x = lag, y = copies_mL),
@@
 ```
-## C2 Site Dust
 
-### Adding a 2 day lag in dust:
+## Canals:
+
+#### Add two day lag
+
 ```{r}
 c22<- c2%>% mutate(
-  lag =Lag(c2$t7sum, shift  = 2)) #Make lag from ccf
+  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
 ```
 
-### Create a linear Model: 
+#### Create a linear model
+
 ```{r}
-lm_spec<- linear_reg()
-lm_fit<- fit(lm_spec, copies_mL~lag, data = c22)
-tidy(lm_fit)
+#Create Recipe for Growth 
+growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)
+
+#Set up linear model
+lm_mod<- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+#Workflow that adds recipe to model
+Growth_wflow<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(growth_recipe_c2)
+
+#Use workflow to fit model to  data set
+growth_fit_c2<- Growth_wflow %>%
+  fit(data = c22)
+
+#View as Tibble 
+growth_fit_c2 %>%
+  extract_fit_parsnip() %>%
+  tidy()
 ```
 
-### Plot C2 T7 Sum:
+### Use model to test data
+
 ```{r}
-c22 %>% ggplot() + geom_point(
-  aes(x = lag, 
-      y = copies_mL)) +
-  geom_smooth(aes(x = lag,
-                  y = copies_mL),
-              method = "lm")+
+aug_test_c2 <- augment(growth_fit_c2, c22)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
+rd_metrics<- full_join(rmse, rsq)
+```
+
+### Plot
+
+```{r}
+aug_test_c2 %>% ggplot(aes(copies_mL, .pred)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
   stat_poly_eq(aes(x = lag, y = copies_mL),
@@
 
-_______________________
-All Variables
+# Multivariate Models
 
-New data set
+[This code is still in progress:]{.ul}
+
+#### Make New Dataset
+
 ```{r}
@@
 ```
+
+#### Site-specific Data sets
+
 ```{r}
@@
 ## BO: Using all Variables vs. Growth
+
 ```{r}
@@
 
-#Use workflow to fit model to train data set
+#Use workflow to fit model to  data set
 growth_fit_bo<- Growth_wflow %>%
@@
 ```
+
 Plot
+
 ```{r}
@@
 ## C2: Using all Variables vs. Growth
+
 ```{r}
@@
 
-#Use workflow to fit model to train data set
+#Use workflow to fit model to  data set
 growth_fit_c2<- Growth_wflow %>%
@@
 ```
+
 ## RD: Using all Variables vs. Growth
+
 ```{r}
@@
 
-#Use workflow to fit model to train data set
+#Use workflow to fit model to  data set
 growth_fit_rd<- Growth_wflow %>%

### 2023-03-17 products/manuscript/Manuscript.qmd
Project Pt. 3: Final Edits

@@
 As mentioned above, the lags identified may not correspond to the most significant relationship between the two variables.
-While a lag of -1 for Blind Oso and -2 for the Canals expalins the data well, a lag of -2 for the Gulf site was more fitting.
+While a lag of -1 for Blind Oso and -2 for the Canals explains the data well, a lag of -2 for the Gulf site was more fitting.
 This makes sense, as taking a look at the copies_mL plot above, we do see a \~2 day lag in response, where we have a spike in dust, and 2 days later, a spike in copies_mL.
+
+[A few things to note:]{.ul}
+
+-   My stats code contains some chunks that create predicted values based off of the current data set.
+Since my data set is small, I am not able to portion of \~25% to use to test my model.
+Therefore, plotting the predicted vs.
+actual will give an R2 of 1.00.
+I am continuing to find a good validation method.
+
+-   I am also planning on running a multi-variate model to look at multiple predictors of interests (temp, salinity, nutrients), but am currently running into an issue with my model output, where rather than getting values, I receive NaN.
+This code can be found in my `dust_copies_stats.qmd` code.
 

### 2023-03-18 code/analysis_code/stats/dust_copies_stats.qmd
Fixed STATS!
- Ran AIC
- Residuals
- Better interpretations

@@
 
-## Gulf Site:
-
-#### Add two day lag
-
-```{r}
-rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-#### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_rd)
-
-#Use workflow to fit model to  data set
-growth_fit_rd<- Growth_wflow %>%
-  fit(data = rd2)
-
-#View as Tibble 
-growth_fit_rd %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_rd <- augment(growth_fit_rd, rd2)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### Plot
-
-```{r}
-aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (RD+2)")
-```
-
 ## Blind Oso Site:
@@
 rd_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+Approaches based on such selection criteria essentially try to guess how the model would perform if it were to be fit to new data, without actually trying to do it (in contrast to CV).
+
+```{r}
+bo_mod<- lm(copies_mL ~ lag, data = bo2)
+AIC(bo_mod)
+```
+
+### Residuals
+
+```{r}
+res_bo<- resid(bo_mod)
+plot(fitted(bo_mod), res_bo)
+abline(0,0)
 ```
@@
 
+### AIC to assess Performance
+
+```{r}
+c2_mod<- lm(copies_mL ~ lag, data = c22)
+AIC(c2_mod)
+```
+
+### Residuals
+
+```{r}
+res_c2<- resid(c2_mod)
+plot(fitted(c2_mod), res_c2)
+abline(0,0)
+```
+
 ### Plot
@@
 
+## Gulf Site:
+
+#### Add two day lag
+
+```{r}
+rd2<- rd%>% mutate(
+  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
+```
+
+#### Create a linear model
+
+```{r}
+#Create Recipe for Growth 
+growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)
+
+#Set up linear model
+lm_mod<- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+#Workflow that adds recipe to model
+Growth_wflow<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(growth_recipe_rd)
+
+#Use workflow to fit model to  data set
+growth_fit_rd<- Growth_wflow %>%
+  fit(data = rd2)
+
+#View as Tibble 
+growth_fit_rd %>%
+  extract_fit_parsnip() %>%
+  tidy()
+```
+
+### Use model to test data
+
+```{r}
+aug_test_rd <- augment(growth_fit_rd, rd2)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
+rd_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+rd_mod<- lm(copies_mL ~ lag, data = rd2)
+AIC(rd_mod)
+```
+
+### Residuals
+
+```{r}
+res_rd<- resid(rd_mod)
+plot(fitted(rd_mod), res_rd)
+abline(0,0)
+```
+
+### Plot
+
+```{r}
+aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4) +
+  labs(title = "LM DustxCopies (RD+2)")
+```
+
+##
+
 # Multivariate Models
 
-[This code is still in progress:]{.ul}
+[This code is still in progress. For now, please ignore:]{.ul}
 

### 2023-03-18 products/manuscript/Manuscript.qmd
Fixed STATS!
- Ran AIC
- Residuals
- Better interpretations

@@
 -   My stats code contains some chunks that create predicted values based off of the current data set.
-Since my data set is small, I am not able to portion of \~25% to use to test my model.
+Since my data set is small, I am not able to portion of \~25% to use to test my model as it could impact the strength of my model.
 Therefore, plotting the predicted vs.
@@
 
+    -   After reviewing module 9, it looks like Information Criteria may be a proper alternative (REF: MOD9/Model Performance and Overfitting).
+
+        -   Will need to interpret findings as exploratory and hypothesis generating, and will need to be careful about trying to draw generalizable conclusions.
+
+        -   AIC:
+
+            -   Blind Oso = 271.25
+
+            -   Canals = 243.97
+
+            -   Gulf = 292.13
+
 -   I am also planning on running a multi-variate model to look at multiple predictors of interests (temp, salinity, nutrients), but am currently running into an issue with my model output, where rather than getting values, I receive NaN.
 This code can be found in my `dust_copies_stats.qmd` code.
+
+### Residuals:
+
+Instead of (or in addition to) plotting observed versus predicted for continuous outcomes, you can plot the difference between the two.
+These differences are called the residuals.
+What you are looking for is a cloud of points with no discernible pattern.
+If there is a pattern (e.g., an overall skew, or more points above the 0 y-axes than below), it again suggests that there is still some pattern/signal in the data that the model didn't capture.
+
+![](images/00000b-01.png){width="406"}
+
+![](images/00000f-01.png){width="404"}
+
+![](images/000009%20(1)-01.png){width="404"}
 

### 2023-03-20 code/analysis_code/stats/dust_copies_stats.qmd
stats

@@
   select(copies_mL, t7sum, site, date) 
+```
+
+# Check Distributions
+
+### Copies per mL
+
+```{r}
+
+#Normal
+master %>% 
+  ggplot() + geom_histogram(
+    aes(
+      x = copies_mL)) +
+  labs(title = "Non-Log")+ 
+  facet_wrap(~site)
+
+#Log-Transformed
+master %>% 
+  ggplot() + geom_histogram(
+    aes(
+      x = log(copies_mL))) +
+  labs(title = "LOG-TRANS")+ 
+  facet_wrap(~site)
+```
+
+**We will use the log-transformed data as it has a \~ normal distribution.**
+
+## Log-transform Copies
+
+```{r}
+#master <- 
+ # master %>%
+ # mutate(copies_mL = log(master$copies_mL))
 ```
@@
 
-##
-
 # Multivariate Models
@@
   filter(hr_cst %in% "01") %>%
-  select(!c(t1, t7, t13, t19, hr_cst, tsum, t7sum, ...1, SUM, dust,  Type, date )) %>%
+  select(site, sal, temp, amm, nn,orthop, copies_mL, t7sum) %>%
   na.omit()
@@
 
-#View as Tibble 
+#Pull model fit using parsnip()
 growth_fit_bo %>%
@@
 ```
+### Use model to test data
 
+```{r}
+aug_test_bo_a <- augment(growth_fit_bo, bo_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_bo_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_bo_a %>% rsq(truth = copies_mL, .pred)
+bo_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+bo_a_mod<- lm(copies_mL ~ ., data = bo_a)
+AIC(bo_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_bo_a<- resid(bo_a_mod)
+plot(fitted(bo_a_mod), res_bo_a)
+abline(0,0)
+```
 ## C2: Using all Variables vs. Growth
@@
 ```
+### Use model to test data
 
+```{r}
+aug_test_c2_a <- augment(growth_fit_c2, c2_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_c2_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_c2_a %>% rsq(truth = copies_mL, .pred)
+c2_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+c2_a_mod<- lm(copies_mL ~ ., data = c2_a)
+AIC(c2_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_c2_a<- resid(c2_a_mod)
+plot(fitted(c2_a_mod), res_c2_a)
+abline(0,0)
+```
 ## RD: Using all Variables vs. Growth
@@
 ```
+### Use model to test data
+
+```{r}
+aug_test_rd_a <- augment(growth_fit_rd, rd_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_rd_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_rd_a %>% rsq(truth = copies_mL, .pred)
+rd_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+rd_a_mod<- lm(copies_mL ~ ., data = rd_a)
+AIC(rd_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_rd_a<- resid(rd_a_mod)
+plot(fitted(rd_a_mod), res_rd_a)
+abline(0,0)
+```

### 2023-03-20 code/analysis_code/stats/uni_variate.qmd
stats

--- 2023-03-18 code/analysis_code/stats/dust_copies_stats.qmd
@@
 ```{r}
-master<- 
+master0<- 
   dust %>% full_join(data) %>%
@@
 
-master<- master %>%
+master<- master0 %>%
   select(copies_mL, t7sum, site, date) 
+```
+
+# Check Distributions
+
+### Copies per mL
+
+```{r}
+
+#Normal
+master %>% 
+  ggplot() + geom_histogram(
+    aes(
+      x = copies_mL)) +
+  labs(title = "Non-Log")+ 
+  facet_wrap(~site)
+
+#Log-Transformed
+master %>% 
+  ggplot() + geom_histogram(
+    aes(
+      x = log(copies_mL))) +
+  labs(title = "LOG-TRANS")+ 
+  facet_wrap(~site)
+```
+
+**We will use the log-transformed data as it has a \~ normal distribution.**
+
+## Log-transform Copies
+
+```{r}
+#master <- 
+ # master %>%
+ # mutate(copies_mL = log(master$copies_mL))
 ```
@@
 
-##
+# TEMPERATURE
 
-# Multivariate Models
+### Blind Oso
 
-[This code is still in progress. For now, please ignore:]{.ul}
-
-#### Make New Dataset
-
-```{r}
-master2<- dust0 %>% full_join(data) %>%
-  filter(hr_cst %in% "01") %>%
-  select(!c(t1, t7, t13, t19, hr_cst, tsum, t7sum, ...1, SUM, dust,  Type, date )) %>%
-  na.omit()
+```{}
 ```
-
-#### Site-specific Data sets
-
-```{r}
-bo_a<- master2 %>%
-  filter(site %in% "Blind Oso") %>%
-  select(!site)
-
-c2_a<- master2 %>%
-  filter(site %in% "Canals")%>%
-  select(!site)
-
-rd_a<- master2 %>%
-  filter(site %in% "Gulf")%>%
-  select(!site)
-```
-
-## BO: Using all Variables vs. Growth
-
-```{r}
-#Create Recipe for Growth and all other parameters
-growth_recipe_bo<- recipe(copies_mL ~ ., data = bo_a)
-
-#Set up linear model
-lm_mod_all<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod_all) %>%
-  add_recipe(growth_recipe_bo)
-
-#Use workflow to fit model to  data set
-growth_fit_bo<- Growth_wflow %>%
-  fit(data = bo_a)
-
-#View as Tibble 
-growth_fit_bo %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-Plot
-
-```{r}
-tidy(growth_fit_bo) %>% 
-  dwplot(dot_args = list(size = 2, color = "black"),
-         whisker_args = list(color = "black"),
-         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
-```
-
-## C2: Using all Variables vs. Growth
-
-```{r}
-#Create Recipe for Growth and all other parameters
-growth_recipe_c2<- recipe(copies_mL ~ ., data = c2_a)
-
-#Set up linear model
-lm_mod_all<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod_all) %>%
-  add_recipe(growth_recipe_c2)
-
-#Use workflow to fit model to  data set
-growth_fit_c2<- Growth_wflow %>%
-  fit(data = c2_a)
-
-#View as Tibble 
-growth_fit_c2 %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-Plot
-
-```{r}
-tidy(growth_fit_c2) %>% 
-  dwplot(dot_args = list(size = 2, color = "black"),
-         whisker_args = list(color = "black"),
-         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
-```
-
-## RD: Using all Variables vs. Growth
-
-```{r}
-#Create Recipe for Growth and all other parameters
-growth_recipe_rd<- recipe(copies_mL ~ ., data = rd_a)
-
-#Set up linear model
-lm_mod_all<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod_all) %>%
-  add_recipe(growth_recipe_rd)
-
-#Use workflow to fit model to  data set
-growth_fit_rd<- Growth_wflow %>%
-  fit(data = rd_a)
-
-#View as Tibble 
-growth_fit_rd %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-Plot
-
-```{r}
-tidy(growth_fit_rd) %>% 
-  dwplot(dot_args = list(size = 2, color = "black"),
-         whisker_args = list(color = "black"),
-         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
-```

### 2023-03-22 code/analysis_code/analysis_README.Rmd
Added distribution stats

@@
 (3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
-(4) The stats folder contains `dust_copies_stats.qmd` that runs models on variables of interest. \
+(4) The stats folder contains:
+- `distributiom.qmd`: Checks distribution of variables of interest
+- `multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)
+- `uni_variate.qmd` : runs univariate models on variable and predictors of interest.

### 2023-03-22 code/analysis_code/stats/uni_variate.qmd
Added distribution stats

@@
   select(copies_mL, t7sum, site, date) 
-```
-
-# Check Distributions
-
-### Copies per mL
-
-```{r}
-
-#Normal
-master %>% 
-  ggplot() + geom_histogram(
-    aes(
-      x = copies_mL)) +
-  labs(title = "Non-Log")+ 
-  facet_wrap(~site)
-
-#Log-Transformed
-master %>% 
-  ggplot() + geom_histogram(
-    aes(
-      x = log(copies_mL))) +
-  labs(title = "LOG-TRANS")+ 
-  facet_wrap(~site)
-```
-
-**We will use the log-transformed data as it has a \~ normal distribution.**
-
-## Log-transform Copies
-
-```{r}
-#master <- 
- # master %>%
- # mutate(copies_mL = log(master$copies_mL))
 ```
@@
 ### Blind Oso
-
-```{}
-```

### 2023-03-28 code/analysis_code/stats/uni_variate.qmd
edits to stats

@@
 
-```{r}
+```{r, include = FALSE}
 library(tidyverse)
@@
 
-```{r}
+```{r, include = FALSE}
 dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
@@
 
-## Blind Oso Site:
+# Blind Oso Site:
 
-#### Add one day lag
+### Add one day lag
 
@@
 
-#### Create a linear model
+### Plot
+
+```{r}
+bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4) +
+  labs(title = "LM DustxCopies (BO+1)")
+```
+
+### Create a linear model
 
@@
 
-### Plot
-
-```{r}
-aug_test_bo %>% ggplot(aes(copies_mL, .pred)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (BO+1)")
-```
-
-## Canals:
+# Canals:
 
@@
 
-#### Create a linear model
+### Plot
+
+```{r}
+c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4) +
+  labs(title = "LM DustxCopies (C2+2)")
+```
+
+### Create a linear model
 
@@
 
-### Plot
-
-```{r}
-aug_test_c2 %>% ggplot(aes(copies_mL, .pred)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (C2+2)")
-```
-
 ## Gulf Site:
 
-#### Add two day lag
+### Add two day lag
 
@@
 
-#### Create a linear model
+### Plot
+
+```{r}
+rd2 %>% ggplot(aes(lag, copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4) +
+  labs(title = "LM DustxCopies (RD+2)")
+```
+
+### Create a linear model
 
@@
 
-### Plot
-
-```{r}
-aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (RD+2)")
-```
-
 # TEMPERATURE

### 2023-03-28 code/analysis_code/dust/dust_and_copies.qmd
edits

@@
   theme(
-    axis.title.y = element_text(size = 15),
+    axis.title.y = element_text(size = 10),
     axis.text.y = element_text(size = 10),
@@
     axis.ticks.x = element_blank(),
-    axis.title.y = element_text(size = 15),
+    axis.title.y = element_text(size = 10),
     axis.text.y = element_text(size = 10),
@@
 Combine Dust and Copies
-```{r, fig.height = 4, fig.width= 4}
+```{r}
 fig<- ggarrange(tot_vib, dust_fig + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")

### 2023-03-31 code/analysis_code/stats/distribution.qmd
stats

---
title: "Distriburion"
format:
  html:
    theme: default
---

# Libs/Data

#### Libraries

```{r}
library(tidyverse)
library(here)
library(ggpubr)
```

#### Data

```{r}
d<- read_csv(here("data", "processed_data","tx_master.csv"))
dust<- read_csv(here("data", "processed_data","dust_master.csv"))
```

#### Clean Data

##### Metadata

```{r}
d_clean<- d %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"),
         hr_cst %in% "01") #Just to get rid of duplicates

master<- d_clean %>%
  select(copies_mL,site, date) %>%
  filter(!is.na(site))

#Site Specific datasets
bo<- 
  master %>%
  filter(site %in% "Blind Oso")
c2<- 
  master %>%
  filter(site %in% "Canals")
rd<- 
  master %>%
  filter(site %in% "Gulf")
```

##### Dust

```{r}
#Daily df
daily<- 
  dust %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  mutate(t7sumlog = log10(t7sum)) #Creates column that is log transformed 
```

# Check Distributions of Data

## \[Copies/mL\]

#### Visualization

**Density plot**: the **density** plot provides a visual judgment about whether the distribution is bell shaped.

```{r}
ggdensity(
  bo$copies_mL,
  main = "Blind Oso",
  xlab = "Copies per mL")

ggdensity(
  c2$copies_mL,
  main = "Canals",
  xlab = "Copies per mL")

ggdensity(
  rd$copies_mL,
  main = "Gulf",
  xlab = "Copies per mL")
```

**Q-Q plot**: **Q-Q plot** (or quantile-quantile plot) draws the correlation between a given sample and the normal distribution. A 45-degree reference line is also plotted.

```{r}
ggqqplot(bo$copies_mL)

ggqqplot(c2$copies_mL)

ggqqplot(rd$copies_mL)
```

### Normality Test

**Shapiro-Wilk's method** is widely recommended for normality test and it provides better power than K-S. It is based on the correlation between the data and the corresponding normal scores.

-   Note that, normality test is sensitive to sample size. Small samples most often pass normality tests. Therefore, it's important to combine visual inspection and significance test in order to take the right decision.

    ```{r}
    shapiro.test(bo$copies_mL)
    shapiro.test(c2$copies_mL)
    shapiro.test(rd$copies_mL)
    ```

    A p-value > 0.05 implies that the distribution of the data is **not significantly differen**t from normal distribution. In other words, **we can assume the normality.**

## \[Dust\]

**Q-Q plot**

```{r}
ggqqplot(daily$t7sum)
ggqqplot(daily$t7sumlog)
```

###

### Normality Test

```{r}
shapiro.test(daily$t7sum) #Normal data
shapiro.test(daily$t7sumlog) #Log trans data
```

### 2023-04-03 code/analysis_code/stats/cca.qmd
edits

--- 2023-03-28 code/analysis_code/stats/uni_variate.qmd
@@
 ---
-title: "env_and_copies_stats"
+title: "cca"
 format:
@@
 
-### Libraries
+# Libraries
 
@@
 library(here)
-library(Hmisc) #For lag function
-library(tidymodels) #For modeling
-library(ggpmisc) #For R2 equation on graph
-library(dotwhisker) #Fpr visualization
 ```
@@
   filter(!is.na(date),
-         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
+         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
+  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)
 
@@
   filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
+master<- master0[c(1:10,15,33)]
 
-master<- master0 %>%
-  select(copies_mL, t7sum, site, date) 
 ```
@@
 ### Make Site DF
+
+**Need to include 7/5 and 7/6 for previous dust data (will be used for lag analysis)**
 
@@
 
+\*\*Dust (t7sum) was log transformed
+
 ## Cross-correlation:
+
+**For this, we cannot have NAs so we choose the strict daily time series**
+
+\***Remember that dust is log transformed**
 
@@
 
-# Blind Oso Site:
++------+-----+
+| Site | Lag |
++======+=====+
+| BO   | 1   |
++------+-----+
+| C2   | 2   |
++------+-----+
+| RD   | 2\* |
++------+-----+
 
-### Add one day lag
-
-```{r}
-bo2<- bo%>% mutate(
-  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-### Plot
-
-```{r}
-bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (BO+1)")
-```
-
-### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_bo)
-
-#Use workflow to fit model to  data set
-growth_fit_bo<- Growth_wflow %>%
-  fit(data = bo2)
-
-#View as Tibble 
-growth_fit_bo %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_bo <- augment(growth_fit_bo, bo2)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### AIC to assess Performance
-
-Approaches based on such selection criteria essentially try to guess how the model would perform if it were to be fit to new data, without actually trying to do it (in contrast to CV).
-
-```{r}
-bo_mod<- lm(copies_mL ~ lag, data = bo2)
-AIC(bo_mod)
-```
-
-### Residuals
-
-```{r}
-res_bo<- resid(bo_mod)
-plot(fitted(bo_mod), res_bo)
-abline(0,0)
-```
-
-# Canals:
-
-#### Add two day lag
-
-```{r}
-c22<- c2%>% mutate(
-  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-### Plot
-
-```{r}
-c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (C2+2)")
-```
-
-### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_c2)
-
-#Use workflow to fit model to  data set
-growth_fit_c2<- Growth_wflow %>%
-  fit(data = c22)
-
-#View as Tibble 
-growth_fit_c2 %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_c2 <- augment(growth_fit_c2, c22)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### AIC to assess Performance
-
-```{r}
-c2_mod<- lm(copies_mL ~ lag, data = c22)
-AIC(c2_mod)
-```
-
-### Residuals
-
-```{r}
-res_c2<- resid(c2_mod)
-plot(fitted(c2_mod), res_c2)
-abline(0,0)
-```
-
-## Gulf Site:
-
-### Add two day lag
-
-```{r}
-rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-### Plot
-
-```{r}
-rd2 %>% ggplot(aes(lag, copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (RD+2)")
-```
-
-### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_rd)
-
-#Use workflow to fit model to  data set
-growth_fit_rd<- Growth_wflow %>%
-  fit(data = rd2)
-
-#View as Tibble 
-growth_fit_rd %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_rd <- augment(growth_fit_rd, rd2)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### AIC to assess Performance
-
-```{r}
-rd_mod<- lm(copies_mL ~ lag, data = rd2)
-AIC(rd_mod)
-```
-
-### Residuals
-
-```{r}
-res_rd<- resid(rd_mod)
-plot(fitted(rd_mod), res_rd)
-abline(0,0)
-```
-
-# TEMPERATURE
-
-### Blind Oso
+: \*[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022.
+Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
+Therefore, these CCA were made with only this time series.
+This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
+But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days.
+Below we will run linear models that DO contain these prior dust days.
+And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
+But this ultiamtely means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
+Using a combination of visual and CCA would be helpful.

### 2023-04-03 code/analysis_code/stats/distribution.qmd
edits

@@
 
-    A p-value > 0.05 implies that the distribution of the data is **not significantly differen**t from normal distribution. In other words, **we can assume the normality.**
+    A p-value > 0.05 implies that the distribution of the data is **not significantly differen**t from normal distribution. In other words, **we can assume the normality for the copies_mL data.**
 
@@
 ```
+
+**Based on the Shapiro test, the dust permutation that we are using (t7sum) is not normally distributed.
+After log transforming, we see that this is now normally distributed.
+For downstream analysis, we will need to use the log10(dust).**

### 2023-04-03 code/analysis_code/stats/uni_variate.qmd
edits

@@
   filter(!is.na(date),
-         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
+         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
+  mutate(t7sumlog = log10(t7sum)) #Log transform Dust conc (as it is not normal)
 
@@
 
-**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input.
-This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date.
-We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**
-
-## Cross-correlation:
-
-```{r}
-#Blind Oso
-bo_ccf<- bo %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
-title("Blind Oso DustxCopies CCA", line =1)
-
-#Canals
-c2_ccf<- c2 %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
-title("Canals DustxCopies CCA", line=1)
-
-#Gulf
-rd_ccf<- rd %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
-title("Gulf DustxCopies CCA", line =1)
-```
-
-**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values.
-A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship.
-According to the ccf plots, BO has a significant lag relationship a value of -1.
-In other words, dust occurs, and one day later, we see a response in copies_mL.
-C2 has a lag at -2, and RD at -1**
-
-# Blind Oso Site:
+# Blind Oso:
 
@@
 bo2<- bo%>% mutate(
-  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
+  lag =Lag(bo$t7sumlog, shift  = 1)) %>% #Make lag from ccf
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
@@
 c22<- c2%>% mutate(
-  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
+  lag =Lag(c2$t7sumlog, shift  = 2)) %>% #Make lag from ccf
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
@@
 rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
+  lag =Lag(rd$t7sumlog, shift  = 2)) %>% #Make lag from ccf
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
@@
 
-# TEMPERATURE
+# SALINITY
 
-### Blind Oso
+## Blind Oso
+
+#### New Dataframe
+
+```{r}
+master2<- 
+  master0 %>%
+  select(copies_mL,sal,temp,site)
+```
+
+#### Plot
+
+```{r}
+master2 %>%
+  filter(site %in% "Blind Oso") %>%
+  ggplot(aes(
+    x = sal, 
+    y = copies_mL)) +
+  geom_point() +
+  stat_smooth(method = "lm")+
+  stat_poly_eq(aes(x = sal, y = copies_mL),
+               rr.digits = 4) +
+  labs(title = "Sal x Copies BO")
+```

### 2023-04-03 code/analysis_code/analysis_README.Rmd
Added LM for Nutrient Data

--- 2023-03-16 code/analysis_code/analysis_README.Rmd
@@
 (3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
-(4) The stats folder contains `dust_copies_stats.qmd` that runs models on variables of interest. \
+(4) The stats folder contains:
+- `1_distributiom.qmd`: Checks distribution of variables of interest
+- `2_cca.qmd`: Performs cross-correlation analysis to test for lags in time series (between dust input and Vibrio growth)
+- `3_uni_variate.qmd` : runs univariate models on variable and predictors of interest.
+- `4_multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)

### 2023-04-03 code/analysis_code/stats/2_cca.qmd
Added LM for Nutrient Data

--- 2023-04-03 code/analysis_code/stats/cca.qmd
@@
 ```{r, include = FALSE}
-dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
-data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
+dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
+data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL
 ```
@@
          date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
-  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)
+  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normally distributed)
 
@@
   filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
-master<- master0[c(1:10,15,33)]
+master<- master0[c(1:9,14,32)]
 
@@
 
-### Make Site DF
+### Make data frames for the three sites
 
-**Need to include 7/5 and 7/6 for previous dust data (will be used for lag analysis)**
+Because we are working with lags. We need to include dust data from a few days [prior]{.ul} to sampling, as the dust concentrations will have an impact on Vibrio growth 1-2 days later.
 
@@
 
-# Uni-variate Models
+# Cross-correlation Analysis
 
-**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input.
+From previous literature, we know that Vibrio have responded to dust 24-48 hours after input.
 This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date.
-We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**
+We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.
 
-\*\*Dust (t7sum) was log transformed
+-   For this, we cannot have NAs so we choose the strict daily time series (7/7 - 7/19). We will use the data frames made above for creating the appropriate lag later.
 
-## Cross-correlation:
-
-**For this, we cannot have NAs so we choose the strict daily time series**
-
-\***Remember that dust is log transformed**
+-   Remember that dust is log transformed
 
@@
   filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) 
-ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
+ccf(bo_ccf$t7sum, bo_ccf$copies_mL) #CCA between dust and copies_mL
 title("Blind Oso DustxCopies CCA", line =1)
@@
 Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
-Therefore, these CCA were made with only this time series.
+We cannot have NAs when running CCA, therefore, these CCA were made with only this time series.
 This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
@@
 But this ultiamtely means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
-Using a combination of visual and CCA would be helpful.
+Using a combination of visualization and CCA would be helpful.
+For the guld site, although CCA identified a lag of 1 day being significant, we can see in our exploratory analysis, that there is a \~2 day lag.
+
+<!--# May need to update this last statement after meeting with Erin. 1 day lag may be the best. -->
+
+<!--# May need to include nutrient data for cca -->

### 2023-04-03 code/analysis_code/stats/3_uni_variate.qmd
Added LM for Nutrient Data

---
title: "env_and_copies_stats"
format:
  html:
    theme: default
---

### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

# LOADING AND CLEANING DATA

### Data

```{r, include = FALSE}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL and nutrients 
```

### Clean and Trim

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
master<- master0[c(1,8,14,32)] #Select columns of interest

#View new dataframe
master

```

### Make Site DF

**Need to include 7/5 and 7/6 for previous dust data (will be used for lag analysis)**

```{r}
bo<- master %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

c2<- master %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))

rd<- master %>%
  filter(site %in% "Gulf"| date %in% c(as.Date("2022-07-05"), as.Date("2022-07-06")))
```

# DUST

### Add lags

```{r}
bo2<- bo%>% mutate(
  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

c22<- c2%>% mutate(
  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))

rd2<- rd%>% mutate(
  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
```

### Plot

```{r}
bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (BO+1)",
       x = "log(dust)")

c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (C2+2)",
       x = "log(dust)")

rd2 %>% ggplot(aes(lag, copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = lag, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM DustxCopies (RD+2)",
       x = "log(dust)")
```

### Create a linear model

##### BLIND OSO:

```{r}
#Create Recipe for Growth 
growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_bo)

#Use workflow to fit model to  data set
growth_fit_bo<- Growth_wflow %>%
  fit(data = bo2)

#View as Tibble 
growth_fit_bo %>%
  extract_fit_parsnip() %>%
  tidy()
```

##### CANALS:

```{r}
#Create Recipe for Growth 
growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_c2)

#Use workflow to fit model to  data set
growth_fit_c2<- Growth_wflow %>%
  fit(data = c22)

#View as Tibble 
growth_fit_c2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

##### GULF:

```{r}
#Create Recipe for Growth 
growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model
Growth_wflow<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe_rd)

#Use workflow to fit model to  data set
growth_fit_rd<- Growth_wflow %>%
  fit(data = rd2)

#View as Tibble 
growth_fit_rd %>%
  extract_fit_parsnip() %>%
  tidy()
```

### RMSE and R2 to test model performance

```{r}
aug_test_bo <- augment(growth_fit_bo, bo2)
rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
bo_metrics<- full_join(rmse, rsq)
bo_metrics

aug_test_c2 <- augment(growth_fit_c2, c22)
rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
c2_metrics<- full_join(rmse, rsq)
c2_metrics

aug_test_rd <- augment(growth_fit_rd, rd2)
rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
rd_metrics<- full_join(rmse, rsq)
rd_metrics
```

### AIC to assess Performance

Approaches based on such selection criteria essentially try to guess how the model would perform if it were to be fit to new data, without actually trying to do it (in contrast to CV). For AIC, a rule of thumb is that a difference of 10 between 2 models is meaningful.

```{r}
bo_mod<- lm(copies_mL ~ lag, data = bo2)
AIC(bo_mod)

c2_mod<- lm(copies_mL ~ lag, data = c22)
AIC(c2_mod)

rd_mod<- lm(copies_mL ~ lag, data = rd2)
AIC(rd_mod)
```

### Plot Residuals

```{r}
res_bo<- resid(bo_mod)
plot(fitted(bo_mod), res_bo)
abline(0,0)

res_c2<- resid(c2_mod)
plot(fitted(c2_mod), res_c2)
abline(0,0)

res_rd<- resid(rd_mod)
plot(fitted(rd_mod), res_rd)
abline(0,0)
```

# OTHER ENVIRONMENTAL PARAMETERS

<!--# Is there a lag for Nutrient data too? -->

## Create New Master Data set

```{r}
master2<- 
  master0 %>%
  select(copies_mL,sal,temp,site, amm, orthop, doc, don, tn, do_mgl, nn, sil, din_dip, tdn, toc, pH) #Select parameters of interest

#Make Site Specific DF
b<- 
  master2 %>%
  filter(site %in% "Blind Oso")

c<- 
  master2 %>%
  filter(site %in% "Canals")

r<- 
  master2 %>%
  filter(site %in% "Gulf")
```

## Plot

#### Temperature

```{r}
b %>% ggplot(aes(x = temp, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = temp, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM TempxCopies (BO)")

c %>% ggplot(aes(x = temp, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = temp, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM TempxCopies (C2)")

r %>% ggplot(aes(x = temp, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = temp, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM TempxCopies (RD)")
```

#### Salinity

```{r}
b %>% ggplot(aes(x = sal, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sal, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM salxCopies (BO)")

c %>% ggplot(aes(x = sal, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sal, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM salxCopies (C2)")

r %>% ggplot(aes(x = sal, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sal, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM salxCopies (RD)")
```

#### Ammonia

```{r}
b %>% ggplot(aes(x = amm, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = amm, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM ammxCopies (BO)")

c %>% ggplot(aes(x = amm, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = amm, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM ammxCopies (C2)")

r %>% ggplot(aes(x = amm, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = amm, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM ammxCopies (RD)")
```

#### Orthophosphate

```{r}
b %>% ggplot(aes(x = orthop, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = orthop, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM orthopxCopies (BO)")

c %>% ggplot(aes(x = orthop, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = orthop, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM orthopxCopies (C2)")

r %>% ggplot(aes(x = orthop, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = orthop, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM orthopxCopies (RD)")
```

#### DOC

```{r}
b %>% ggplot(aes(x = doc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = doc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM docxCopies (BO)")

c %>% ggplot(aes(x = doc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = doc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM docxCopies (C2)")

r %>% ggplot(aes(x = doc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = doc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM docxCopies (RD)")
```

#### DON

```{r}
b %>% ggplot(aes(x = don, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = don, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM donxCopies (BO)")

c %>% ggplot(aes(x = don, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = don, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM donxCopies (C2)")

r %>% ggplot(aes(x = don, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = don, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM donxCopies (RD)")
```

#### Total Nitrogen

```{r}
b %>% ggplot(aes(x = tn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tnxCopies (BO)")

c %>% ggplot(aes(x = tn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tnxCopies (C2)")

r %>% ggplot(aes(x = tn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tnxCopies (RD)")
```

#### DO

```{r}
b %>% ggplot(aes(x = do_mgl, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = do_mgl, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM do_mglxCopies (BO)")

c %>% ggplot(aes(x = do_mgl, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = do_mgl, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM do_mglxCopies (C2)")

r %>% ggplot(aes(x = do_mgl, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = do_mgl, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM do_mglxCopies (RD)")
```

#### NN

```{r}
b %>% ggplot(aes(x = nn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = nn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM nnxCopies (BO)")

c %>% ggplot(aes(x = nn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = nn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM nnxCopies (C2)")

r %>% ggplot(aes(x = nn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = nn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM nnxCopies (RD)")
```

#### Silicate

```{r}
b %>% ggplot(aes(x = sil, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sil, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM silxCopies (BO)")

c %>% ggplot(aes(x = sil, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sil, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM silxCopies (C2)")

r %>% ggplot(aes(x = sil, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = sil, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM silxCopies (RD)")
```

#### DIN:DIP

```{r}
b %>% ggplot(aes(x = din_dip, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = din_dip, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM din_dipxCopies (BO)")

c %>% ggplot(aes(x = din_dip, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = din_dip, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM din_dipxCopies (C2)")

r %>% ggplot(aes(x = din_dip, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = din_dip, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM din_dipxCopies (RD)")
```

#### TDN

```{r}
b %>% ggplot(aes(x = tdn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tdn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tdnxCopies (BO)")

c %>% ggplot(aes(x = tdn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tdn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tdnxCopies (C2)")

r %>% ggplot(aes(x = tdn, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = tdn, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tdnxCopies (RD)")
```

#### TOC

```{r}
b %>% ggplot(aes(x = toc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = toc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tocxCopies (BO)")

c %>% ggplot(aes(x = toc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = toc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tocxCopies (C2)")

r %>% ggplot(aes(x = toc, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = toc, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM tocxCopies (RD)")
```

#### pH

```{r}
b %>% ggplot(aes(x = pH, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = pH, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM pHxCopies (BO)")

c %>% ggplot(aes(x = pH, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = pH, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM pHxCopies (C2)")

r %>% ggplot(aes(x = pH, y = copies_mL)) +
  geom_point() +
  stat_smooth(method = "lm")+
  stat_poly_eq(aes(x = pH, y = copies_mL),
               rr.digits = 4) +
  labs(title = "LM pHxCopies (RD)")
```

### 2023-04-03 code/analysis_code/stats/4_multi_variate.qmd
Added LM for Nutrient Data

--- 2023-03-17 code/analysis_code/stats/dust_copies_stats.qmd
@@
 ---
-title: "env_and_copies_stats"
+title: "multi-variate"
 format:
@@
 library(dotwhisker) #Fpr visualization
+library(lubridate)#Dates
 ```
@@
 ```{r}
-dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
-data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
+dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
+data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
 ```
@@
 ```{r}
-dust<- dust0 %>%
+dust<- dust %>%
   filter(!is.na(date),
@@
 
-data<- data0 %>%
+data<- data %>%
   filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
@@
 master<- master %>%
-  select(copies_mL, t7sum, site, date) 
-```
-
-### Make Site DF
-
-```{r}
-bo<- master %>%
-  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))
-
-c2<- master %>%
-  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06")))
-
-rd<- master %>%
-  filter(site %in% "Gulf"| date %in% c(as.Date("2022-07-05"), as.Date("2022-07-06")))
-```
-
-# Uni-variate Models
-
-**From previous literature, we know that Vibrio have responded to dust 24-48 hours after input.
-This can make analysis tricky, as the dust concentration of a specific day does not necessarily correlate with the Vibrio copies of that same date.
-We can use cross-correlation analysis to assess a lag in a time series and see where this lag is most significant.**
-
-## Cross-correlation:
-
-```{r}
-#Blind Oso
-bo_ccf<- bo %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(bo_ccf$t7sum, bo_ccf$copies_mL) 
-title("Blind Oso DustxCopies CCA", line =1)
-
-#Canals
-c2_ccf<- c2 %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
-title("Canals DustxCopies CCA", line=1)
-
-#Gulf
-rd_ccf<- rd %>%
-  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
-title("Gulf DustxCopies CCA", line =1)
-```
-
-**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values.
-A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship.
-According to the ccf plots, BO has a significant lag relationship a value of -1.
-In other words, dust occurs, and one day later, we see a response in copies_mL.
-C2 has a lag at -2, and RD at -1**
-
-## Gulf Site:
-
-#### Add two day lag
-
-```{r}
-rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-#### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_rd<- recipe(copies_mL ~ lag, data = rd2)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_rd)
-
-#Use workflow to fit model to  data set
-growth_fit_rd<- Growth_wflow %>%
-  fit(data = rd2)
-
-#View as Tibble 
-growth_fit_rd %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_rd <- augment(growth_fit_rd, rd2)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_rd %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_rd %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### Plot
-
-```{r}
-aug_test_rd %>% ggplot(aes(lag, copies_mL)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (RD+2)")
-```
-
-## Blind Oso Site:
-
-#### Add one day lag
-
-```{r}
-bo2<- bo%>% mutate(
-  lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-#### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_bo<- recipe(copies_mL ~ lag, data = bo2)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_bo)
-
-#Use workflow to fit model to  data set
-growth_fit_bo<- Growth_wflow %>%
-  fit(data = bo2)
-
-#View as Tibble 
-growth_fit_bo %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_bo <- augment(growth_fit_bo, bo2)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_bo %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_bo %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### Plot
-
-```{r}
-aug_test_bo %>% ggplot(aes(copies_mL, .pred)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (BO+1)")
-```
-
-## Canals:
-
-#### Add two day lag
-
-```{r}
-c22<- c2%>% mutate(
-  lag =Lag(c2$t7sum, shift  = 2)) %>% #Make lag from ccf
-filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
-```
-
-#### Create a linear model
-
-```{r}
-#Create Recipe for Growth 
-growth_recipe_c2<- recipe(copies_mL ~ lag, data = c22)
-
-#Set up linear model
-lm_mod<- linear_reg() %>%
-  set_engine("lm") %>%
-  set_mode("regression")
-
-#Workflow that adds recipe to model
-Growth_wflow<- 
-  workflow() %>%
-  add_model(lm_mod) %>%
-  add_recipe(growth_recipe_c2)
-
-#Use workflow to fit model to  data set
-growth_fit_c2<- Growth_wflow %>%
-  fit(data = c22)
-
-#View as Tibble 
-growth_fit_c2 %>%
-  extract_fit_parsnip() %>%
-  tidy()
-```
-
-### Use model to test data
-
-```{r}
-aug_test_c2 <- augment(growth_fit_c2, c22)
-```
-
-### RMSE and R2 to test model performance
-
-```{r}
-rmse <- aug_test_c2 %>% rmse(truth = copies_mL, .pred)
-rsq <- aug_test_c2 %>% rsq(truth = copies_mL, .pred)
-rd_metrics<- full_join(rmse, rsq)
-```
-
-### Plot
-
-```{r}
-aug_test_c2 %>% ggplot(aes(copies_mL, .pred)) +
-  geom_point() +
-  stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
-  labs(title = "LM DustxCopies (C2+2)")
+  select(site, sal, temp, copies_mL,t7sum, date) #%>%
+ # na.omit()
 ```
@@
 
-[This code is still in progress:]{.ul}
+[This code is still in progress. For now, please ignore:]{.ul}
 
-#### Make New Dataset
+### Site-specific Data sets and Clean df
 
 ```{r}
-master2<- dust0 %>% full_join(data) %>%
-  filter(hr_cst %in% "01") %>%
-  select(!c(t1, t7, t13, t19, hr_cst, tsum, t7sum, ...1, SUM, dust,  Type, date )) %>%
-  na.omit()
+bo_a<- master %>%
+  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
+bo_a<- bo_a %>%
+mutate(
+  lag =Lag(bo_a$t7sum, shift  = 1)) %>% #Create 1 day lag
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
+select(!c(date,t7sum,site))
+
+c2_a<- master %>%
+  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
+c2_a<- c2_a %>%
+mutate(
+  lag =Lag(c2_a$t7sum, shift  = 2)) %>% #Make 2 day lag
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
+select(!c(date,t7sum,site))
+
+rd_a<- master %>%
+  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06')) 
+rd_a<- rd_a %>%
+mutate(
+  lag =Lag(rd_a$t7sum, shift  = 2)) %>% #Make 2 day lag
+filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%
+select(!c(date,t7sum,site))
 ```
 
-#### Site-specific Data sets
-
-```{r}
-bo_a<- master2 %>%
-  filter(site %in% "Blind Oso") %>%
-  select(!site)
-
-c2_a<- master2 %>%
-  filter(site %in% "Canals")%>%
-  select(!site)
-
-rd_a<- master2 %>%
-  filter(site %in% "Gulf")%>%
-  select(!site)
-```
-
-## BO: Using all Variables vs. Growth
+## BO: Growth\~ Sal+Dust
 
@@
 #Create Recipe for Growth and all other parameters
-growth_recipe_bo<- recipe(copies_mL ~ ., data = bo_a)
+growth_recipe_bo<- recipe(copies_mL ~ sal+lag, data = bo_a)
 
@@
          vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
+```
+
+### Use model to test data
+
+```{r}
+aug_test_bo_a <- augment(growth_fit_bo, bo_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_bo_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_bo_a %>% rsq(truth = copies_mL, .pred)
+bo_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+bo_a_mod<- lm(copies_mL ~ temp+sal+lag, data = bo_a)
+AIC(bo_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_bo_a<- resid(bo_a_mod)
+plot(fitted(bo_a_mod), res_bo_a)
+abline(0,0)
 ```
@@
 
+### Use model to test data
+
+```{r}
+aug_test_c2_a <- augment(growth_fit_c2, c2_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_c2_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_c2_a %>% rsq(truth = copies_mL, .pred)
+c2_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+c2_a_mod<- lm(copies_mL ~ ., data = c2_a)
+AIC(c2_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_c2_a<- resid(c2_a_mod)
+plot(fitted(c2_a_mod), res_c2_a)
+abline(0,0)
+```
+
 ## RD: Using all Variables vs. Growth
@@
 ```
+
+### Use model to test data
+
+```{r}
+aug_test_rd_a <- augment(growth_fit_rd, rd_a)
+```
+
+### RMSE and R2 to test model performance
+
+```{r}
+rmse <- aug_test_rd_a %>% rmse(truth = copies_mL, .pred)
+rsq <- aug_test_rd_a %>% rsq(truth = copies_mL, .pred)
+rd_a_metrics<- full_join(rmse, rsq)
+```
+
+### AIC to assess Performance
+
+```{r}
+rd_a_mod<- lm(copies_mL ~ ., data = rd_a)
+AIC(rd_a_mod)
+```
+
+### Residuals
+
+```{r}
+res_rd_a<- resid(rd_a_mod)
+plot(fitted(rd_a_mod), res_rd_a)
+abline(0,0)
+```

### 2023-04-04 code/analysis_code/dust/daily_dusts.qmd
edit
Clean scripts

@@
 ---
+
 **This code is useful for looking at variations of daily dust trends.
 For example, when running statistical analysis, it may be best to look at dust concentration at a specific hour, or the sum per day, or the average.
-We will be working with a sum that is 24hr prior to sample collection.
-This utilizes a data set that has the variations of dust trends **
-Libs
+We will be working with a sum that is 24hr prior to sample collection (t7sum).
+This utilizes a data set that has the variations of dust trends.**
+
+# Getting Started
+
+#### Libraries
+
 ```{r}
@@
 
-Data
+#### Data
+
 ```{r}
@@
 ```
-**There are a few columns that we do not need or that contain NaN. So let's remove those.**
 
-Clean
+# Cleaning
+
+There are a few columns that we do not need or that contain NaN. So let's remove those.
+
+#### Remove NAs
+
 ```{r}
 dust<- dust %>%
-  select(date,t1,t7,t13,t19,tsum,t7sum) %>%
-  filter(!is.na(dust$date))
+  filter(!date %in% as.Date('2022-07-04'))
 ```
 
-Plot T7sum
+#### Select Dates for Daily Time series
+
+```{r}
+dust<-
+  dust %>%
+  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19'))
+```
+
+### Plot T7sum over Daily Time series
+
+<!--# Make this graph look nicer -->
+
 ```{r}

### 2023-04-04 code/analysis_code/dust/dust_and_copies.qmd
Clean scripts

@@
 ---
+
 **This script is good for visualizing copies per mL data and dust data.**
-Library
+
+# Getting Started
+
+#### Libraries
+
 ```{r}
@@
 
-Load Data 
+#### Load Data
+
 ```{r}
@@
 
-Daily Data frames
+# Cleaning
+
+#### Select Daily Data frames
+
 ```{r}
@@
 
-Plot Daily Dust Concentration
+# Visualization
+
+#### Plot Daily Dust Concentration
+
 ```{r}
@@
 
-Total Daily Vibrio
+#### Total Daily Vibrio
+
 ```{r}
@@
 
-Combine Dust and Copies
+#### Combine Dust and Copies
+
 ```{r}
@@
 
-Save as PNG
+#### Save as PNG
+
 ```{r}

### 2023-04-04 code/wrangling_code/1_dust.qmd
Clean scripts

---
title: "Untitled"
format:
  html:
    theme: default
---

**This code takes dust concentration data from NRL and cleans it. Steps include, switching to the correct timezone, creating concentration sums per day, as well as combining the dust data set with the vibrio enumeration data set.**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(lubridate) #To work with dates 
library(here)
library(base) #For ColSum
```

#### Data

```{r}
nrl<- read_csv(here("data","raw_data","dust.csv")) #Dust concentration data set
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Vibrio counts data set from qPCR

```

# Cleaning

Dust AOD was captured at four time points per day. We need to make a column that depicts that date AND hour of collection.

#### Make Year-Month-Day-Hour column from yyymmddtt column

```{r}
nrl1<- nrl %>% 
  mutate(
    ymdh = ymd_h(nrl$yyymmddtt))
```

This dust data set is in Coordinated Universal Time (UTC) and needs to be converted to Central Standard/Daylight Time (which is UTC-6 hours and UTC-5 hours for daylight savings.

#### Switch from UTC --> CST

```{r}
nrl2<- nrl1 %>%
  mutate(
    cst = as_datetime(nrl1$ymdh, tz= c("America/Chicago"))) #CST/CDT
```

#### Make column with just date in CST

```{r}
#Just the data from CST/CDT
nrl3<- nrl2 %>%
  mutate(
    just_date= as.Date(substr(nrl2$cst, 1,10)))

#Create a column for hr in cst
dust_ind<- nrl3 %>%
  mutate(hr_cst = substr(nrl3$cst, 12,13)) %>%
  select(hr_cst, just_date, dust) %>% #Make final DF with dust, hours, and time in CST
  rename("date" = "just_date")
```

Now `dust_ind` data set has the date and hour of collection in CST/CDT, and the concentrations (AOD) at each time point\

For future analyses, it may be helpful to have the dust sum of each day (T0 +T6 +T12 +T18). This could give us an estimate for the dust concentration in a single day.

#### Dust Sum

```{r}
dust_sum<- aggregate(dust_ind$dust, by=list(date=dust_ind$date),FUN=sum) %>% #could also use groupby
 rename("SUM" = "x")
```

This data set `dust_sum` has the dust concentration summed for each calendar day. Now let's combine the dust SUM to the dust INDIVUDUAL data frames\

#### Combine Sum and Individual Data frames

```{r}
dust_master<- dust_sum %>% right_join(dust_ind)
```

#### Save as RDS

```{r}
saveRDS(dust_master, file = here("data", "processed_data", "dust_clean.rds")) #Dust with Cst Time and Hours

```

### 2023-04-04 code/wrangling_code/2_total_copies_calculator.qmd
Clean scripts

--- 2023-02-24 code/wrangling_code/total_copies_calculator.qmd
@@
 ---
-*This script is for cleaning qPCR results on Total Vibrio enumeration.
+
+**This script is for cleaning qPCR results on Total Vibrio enumeration.
 It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio.
 The data set contains a daily and monthly time series.
-We will be examining daily.*
+We will be examining daily.**
 
-Libraries
+# Getting Started
+
+#### Libraries
+
 ```{r}
@@
 
-Load data
+#### Data
+
 ```{r}
@@
 
-Select and Filter for Daily and Monthly Time Series 
+# Cleaning
+
+#### Select and Filter for Daily and Monthly Time Series
+
 ```{r}
@@
 
-Create Site Column from Sample Name 
+#### Create Site Column from Sample Name
+
 ```{r}
@@
 
-Add Dilution Column and Filter column.
+#### Add Dilution Column and Filter column.
+
 The dilution column applies if samples were diluted prior to qPCR.
 For example, if I diluted a sample 1:10, my dilution column would denote "10".
-The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter prior to qPCR.
+The filter column pertains to the amount of water sample (mL) that was concentrated onto a water filter upon collection.
+
 ```{r}
@@
 
-Go from CQ --> Copies/mL
+#### Go from CQ --> Copies/mL
+
 ```{r}
@@
 
-Some samples did not amplify: Replace NaN with 0 for copies/mL
+#### Replace NaN with 0 for copies/mL
+
+Some samples did not amplify and have no values:
+
 ```{r}
@@
 
-Date Column: This adds a date correpsonding the a number in the sample name. 
+#### Data Column
+
+This adds a date corresponding the a number in the sample name.
+
 ```{r}
@@
 
-Each sample (Date) was done in triplicate for qPCR: We need to take the average of qPCR triplicates 
+#### qPCR Triplicated --> Average
+
+Each sample (Date) was done in triplicate for qPCR (technical replicates): We need to take the average of qPCR triplicates
+
 ```{r}
@@
 
-Combine Daily and Monthly data sets back togthether. 
+#### Combine Daily and Monthly data sets back together.
+
 ```{r}
@@
 
-Save as RDS and CSV
+#### Save as CSV
+
 ```{r}

### 2023-04-04 code/wrangling_code/3_df_master.qmd
Clean scripts

---
title: "ENV_COP_DUST"
format:
  html:
    theme: default
---

**This code combines clean data sets (dust, copies, and env)**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
```

#### Load Clean Data

```{r}
dust<- readRDS(here("data", "processed_data", "dust_clean.rds")) #Clean dust data
copies<- read_csv(here("data", "processed_data", "total_clean_copies.csv")) #Clean Copies Data
env<- read_csv(here("data", "processed_data", "ysi_2.csv")) #Clean Environmental Data
```

# Further Cleaning

#### Drop NAs from Environmental Data

```{r}
env<- env %>%
select_if(~ !any(is.na(.))) #Drops Columns with NA
```

#### Combine 3 data sets to make a large master data set

```{r}
master<- dust %>% full_join(env) %>% full_join(copies)
```

#### Clean Data Again

We will get warnings since we have dust data for the whole year, but Nutrient and Copies_mL data for just the monthly and daily time series. This is ok.

```{r}
master<- master %>%
  select(!...1)  %>% #Gets rid of number column from excel
  mutate(
    sal = as.numeric(master$sal), #Set character columns as numeric
    temp = as.numeric(master$temp),
    do_mgl = as.numeric(master$do_mgl),
    do_per = as.numeric(master$do_per),
    pH = as.numeric(master$pH),
    secchi = as.numeric(master$secchi),
    amm = as.numeric(master$amm),
    nn = as.numeric(master$nn),
    orthop = as.numeric(master$orthop),
    sil = as.numeric(master$sil),
    din_dip = as.numeric(master$din_dip),
    tdn = as.numeric(master$tdn),
    don = as.numeric(master$don),
    doc = as.numeric(master$doc),
    toc = as.numeric(master$toc),
    tn = as.numeric(master$tn)) %>%
    mutate(site = recode(site, "BO" = "Blind Oso", #Rename Site Columns
                             "C2" = "Canals",
                            "RD" = "Gulf"))
```

#### Save as CSV

```{r}
write.csv(master,file= here("data","processed_data","tx_master.csv"))

```

### 2023-04-04 products/manuscript/Manuscript.qmd
stats
Added LM for Nutrient Data
Clean scripts

@@
 ```
+
+# Questions for Dr. Handel:
+
+1.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
+
+    -   Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.
+
+    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into,
 
@@
 
+<!--# Let's add Nutrient of interest Data -->
+
 ## Basic statistical analysis
 
-### Cross-correlation Analysis
+### Check Distributions:
+
+<!--# Add distribution statement -->
+
+### Cross-correlation Analysis:
 
@@
 ### Linear Regression Models
+
+<!--# Need to update these plots -->
 

### 2023-04-04 code/analysis_code/env/env_par.qmd
cleaning up

@@
 ---
-*This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.*
 
-Libraries
+**This data set contains environmental parameters: Temp and Salinity, among other environmental parameters.**
+
+# Getting Started
+
+#### Libraries
+
 ```{r}
@@
 
-Load Data
+#### Load Data
+
 ```{r}
@@
 
-Daily data frame
+# Cleaning
+
+#### Daily data frame
+
 ```{r}
@@
 
-Plotting Daily Temperature 
+# Visualization
+
+#### Plotting Daily Temperature
+
 ```{r}
@@
 
-Plotting Daily Salinity
+#### Plotting Daily Salinity
+
 ```{r}
@@
 ```
-Combine Temperature and Salinity Plots 
+
+#### Combine Temperature and Salinity Plots
+
 ```{r, fig.width= 4, fig.height=4}

### 2023-04-04 code/analysis_code/qPCR/daily.qmd
cleaning up

---
title: "Exploratory"
format:
  html:
    theme: default
---

**This code looks at Vibrio enumeration over the daily time series. This information can also be found in `dust_and_copies.qmd`**

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
```

#### Data

```{r}
data<- read_csv(here("data", "processed_data", "tx_master.csv")) 
```

# Cleaning

#### Make mini dataframes for each site

```{r}
#Daily All Sites
daily<- data %>%
  select(date,copies_mL,site,Type) %>%
  filter(Type %in% "Daily") %>%
drop_na() #Get rid of NAs

#Daily by Site
bo_daily <- daily %>% filter(site %in% "Blind Oso") #Blind Oso Bay
c2_daily <- daily %>% filter(site %in% "Canals") #Canals
rd_daily <- daily %>% filter(site %in% "Gulf") #Gulf Site

```

# Visualization

#### Total Daily Vibrio at Gulf

```{r}
rd_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkblue") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Gulf Site") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio at Canals

```{r}
c2_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkorange") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Canals") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio at Blind Oso

```{r}
bo_daily %>%
  ggplot() +geom_line(
    aes(x = date,
        y = copies_mL),
    color = "darkgreen") +
  theme_bw() +
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at Blind Oso") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Total Daily Vibrio

```{r}
tot_vib<- daily %>% ggplot()+geom_line(
  aes(x = date,
      y = copies_mL,
      color = site))+
  facet_wrap(~site) +
  theme_bw()+
  labs(
    x = "Date",
    y = "Copies/mL",
    title = "Total Daily Vibrio at all Sites") +
  theme(plot.title = element_text(hjust = 0.5)) + #center title
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #Tilt x axis labels 
  theme(legend.position = "none") #Removes legend

tot_vib
```

#### Save as PNG

```{r}
png(file = here("results", "plots", "tot_vib_allsites_daily.png"))
tot_vib
dev.off()
```

### 2023-04-04 code/analysis_code/stats/2_cca.qmd
edits
cleaning up

@@
 
-# Libraries
+**This code is for running cross-correlation analyses to assess significant lag in our time series.**
+
+# Getting Started
+
+#### Libraries
 
@@
 
-# LOADING AND CLEANING DATA
-
-### Data
+#### Data
 
@@
 ```
+
+# Cleaning
 
@@
 
-+------+-----+
 | Site | Lag |
-+======+=====+
+|------|-----|
 | BO   | 1   |
-+------+-----+
 | C2   | 2   |
-+------+-----+
-| RD   | 2\* |
-+------+-----+
+| RD   | 1   |
 
@@
 And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
-But this ultiamtely means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
+But this ultimately means that the lag identified above may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
 Using a combination of visualization and CCA would be helpful.
-For the guld site, although CCA identified a lag of 1 day being significant, we can see in our exploratory analysis, that there is a \~2 day lag.
-
-<!--# May need to update this last statement after meeting with Erin. 1 day lag may be the best. -->
 

### 2023-04-04 code/analysis_code/stats/3_uni_variate.qmd
edits
cleaning up

@@
 
-### Libraries
+**This code runs uni-variate models on parameters of interest.**
+
+# Getting Started
+
+#### Libraries
 
@@
 
-### Add lags
+#### Add Lag from `2_cca.qmd`
 
@@
 rd2<- rd%>% mutate(
-  lag =Lag(rd$t7sum, shift  = 2)) %>% #Make lag from ccf
+  lag =Lag(rd$t7sum, shift  = 1)) %>% #Make lag from ccf
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
@@
 
-### Plot
+#### Plot
 
@@
 
-### Create a linear model
+#### Create Linear Model
 

### 2023-04-04 products/manuscript/Manuscript.qmd
cleaning up

@@
 
+# Questions for Dr. Handel:
+
+1.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
+
+    -   Script Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.
+
+    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into.
+
+2.  Looking at `code/analysis_code/stats/4_multi_variate.qmd` , I am a little stuck on how to best create a multivariate model.
+
+    -   Example: We can see this Salinity at Blind Oso drop at the same time that dust is coming in. This will most likely have an impact of Vibrio growth, but Salinity is probably non-linear.
+
+    -   Taking a look at the section titled Blind Oso: Growth \~Salinity + Dust, does my overall workflow look correct?
+
+3.  Looking at my statistics/models, is there anything that you think is strikingly missing?
+
 ```{r, echo=FALSE, message=FALSE}
@@
 
-# Questions for Dr. Handel:
+# Summary/Abstract
 
-1.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
-
-    -   Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.
-
-    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into,
-
-# Summary/Abstract
+Each year, plumes of
+Saharan dust travel across the Atlantic to be deposited in the surface waters of the Caribbean and Gulf of Mexico.
+Dust aerosols serve as a significant source of nutrients that can elicit a fertilization effect on marine coastal waters, leading to rapid and potentially harmful blooms (e.g.
+red tides).
+Previous work has characterized the response of the marine bacteria *Vibrio* to dust input under the oligotrophic settings of the Florida Keys, but less is (1) quantify *Vibrio* population dynamics and composition, and (2) characterize
+potential microbial blooms and community shifts in response to dust input in coastal sites with higher ambient nutrient levels.
+Monthly and daily sampling influences on microbialpopulations and to capture before, during, and after a Saharan dust event.
+Culturable *Vibrio* were enumerated as CFU ml^-1^ via spread plating on selective media.
+qPCRwill be used to estimate total counts of *Vibrio* and of pathogenic *Vibrio* species (*V.
+cholerae, V.
+parahaemolyticus, V.
+vulnificus,* and *V.
+alginolyticus*) while *Vibrio* composition will be determined using hsp60 gene or 16S rRNA amplicon sequencing.
+Microbial community composition will be determined using 16S rRNA and 18S rRNA gene amplicon sequencing to examine shifts in prokaryotic and eukaryotic composition over time.
+This study will provide an increasedunderstanding of the conditions that can elicit potentially harmful blooms, highlighting the need for further research to understand effects of dust
 

### 2023-04-05 code/analysis_code/stats/2_cca.qmd
Project edits

@@
 <!--# May need to include nutrient data for cca -->
+
+#### Save Figures
+
+```{r}
+jpeg(file = here("results", "plots", "cca", "cca_bo.jpg"))
+bo_ccf<- bo %>%
+  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) 
+  ccf(bo_ccf$t7sum, bo_ccf$copies_mL) #CCA between dust and copies_mL
+  title("Blind Oso DustxCopies CCA", line =1)
+dev.off()
+
+jpeg(file = here("results", "plots", "cca", "cca_c2.jpg"))
+c2_ccf<- c2 %>%
+  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
+  ccf(c2_ccf$t7sum, c2_ccf$copies_mL)
+  title("Canals DustxCopies CCA", line=1)
+dev.off()
+
+jpeg(file = here("results", "plots", "cca", "cca_rd.jpg"))
+rd_ccf<- rd %>%
+  filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))
+  ccf(rd_ccf$t7sum, rd_ccf$copies_mL) 
+  title("Gulf DustxCopies CCA", line =1)
+dev.off()
+```

### 2023-04-05 code/analysis_code/stats/3_uni_variate.qmd
Project edits

@@
 ```{r}
-bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
+fig1<- bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
@@
 
-c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
+fig2<- c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
@@
 
-rd2 %>% ggplot(aes(lag, copies_mL)) +
+fig3<- rd2 %>% ggplot(aes(lag, copies_mL)) +
   geom_point() +
@@
        x = "log(dust)")
+```
+
+#### Stack and Save
+
+```{r}
+png(file = here("results", "plots", "stats", "lm.png"))
+ggarrange(fig1, fig2, fig3 + font("x.text", size = 10), ncol = 1, nrow = 3)
+dev.off()
 ```

### 2023-04-05 products/manuscript/Manuscript.qmd
Project edits

@@
 
-Each year, plumes of
-Saharan dust travel across the Atlantic to be deposited in the surface waters of the Caribbean and Gulf of Mexico.
+Each year, plumes of Saharan dust travel across the Atlantic to be deposited in the surface waters of the Caribbean and Gulf of Mexico.
 Dust aerosols serve as a significant source of nutrients that can elicit a fertilization effect on marine coastal waters, leading to rapid and potentially harmful blooms (e.g.
 red tides).
-Previous work has characterized the response of the marine bacteria *Vibrio* to dust input under the oligotrophic settings of the Florida Keys, but less is (1) quantify *Vibrio* population dynamics and composition, and (2) characterize
-potential microbial blooms and community shifts in response to dust input in coastal sites with higher ambient nutrient levels.
-Monthly and daily sampling influences on microbialpopulations and to capture before, during, and after a Saharan dust event.
-Culturable *Vibrio* were enumerated as CFU ml^-1^ via spread plating on selective media.
-qPCRwill be used to estimate total counts of *Vibrio* and of pathogenic *Vibrio* species (*V.
-cholerae, V.
-parahaemolyticus, V.
-vulnificus,* and *V.
-alginolyticus*) while *Vibrio* composition will be determined using hsp60 gene or 16S rRNA amplicon sequencing.
-Microbial community composition will be determined using 16S rRNA and 18S rRNA gene amplicon sequencing to examine shifts in prokaryotic and eukaryotic composition over time.
-This study will provide an increasedunderstanding of the conditions that can elicit potentially harmful blooms, highlighting the need for further research to understand effects of dust
+Previous work has characterized the response of the marine bacteria *Vibrio* to dust input under the oligotrophic settings of the Florida Keys, but less is known regarding settings with higher baseline nutrient levels.
+This project aims to quantify *Vibrio* population dynamics in response to dust input in coastal sites with higher ambient nutrient levels.
+Daily sampling took place in Corpus Christi, TX to capture before, during, and after a Saharan dust event and quantitative PCR (qPCR) was used to estimate total counts of *Vibrio.* This study will provide an increased understanding of the conditions that can elicit potentially harmful blooms, highlighting the need for further research to understand the effects of dust deposition in non-oligotrophic waters.
 
@@
 
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year ([d'Almeida, 1986](https://journals.ametsoc.org/view/journals/apme/25/7/1520-0450_1986_025_0903_amfsdt_2_0_co_2.xml)).
+Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig.
+1).
+These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
+Dust aerosols can harbor a wide diversity of bacteria ([Kellogg et al., 2004](https://link.springer.com/content/pdf/10.1023/B:AERO.0000032947.88335.bb.pdf)), fungi ([Ramírez Camejo et al., 2022](https://link.springer.com/article/10.1007/s10453-022-09752-9)), virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)).
+These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron ([Graham & Duce, 1967](https://www.sciencedirect.com/science/article/pii/0004698182901986), [Mills et al., 2004](https://pubmed.ncbi.nlm.nih.gov/15152251/), [Savoie et al., 1989](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/JD094iD04p05069?casa_token=PMRaet9QSYsAAAAA:nb8yBorxdKaEWwGfI5aNILJ0D4h6drNu2YY1y1y8ExbQ03z59EUTQTOHoxinzsxRtR4jzsfSA73_Hig)).
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575), [Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
+
+Previous work conducted in the West Florida Shelf has shown that increased iron availability can stimulate the growth of the nitrogen-fixing cyanobacterium *Trichodesmium*.
+The high amount of nitrogen produced via fixation can be adequate enough to stimulate toxic red tides, caused by the dinoflagellate *Karenia brevis* ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575)).
+
+While it is documented that these dust events can stimulate growth of harmful phytoplankton, emerging evidence shows that heterotrophic bacteria are responding to the episodic influx of limiting resources and substrate, resulting in blooms of bacteria that are associated with disease, which we refer to as harmful bacterial blooms.
+Among those that respond are bacterium belonging to the genus *Vibrio*, a group of ubiquitous marine opportunistic heterotrophs.
+While considered conditionally rare (comprising of \<1% of the microbial community), under the proper conditions, this genus can rapidly bloom to make up a large percentage of the community over a short period of time ([Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
+*Vibrio* species are found in brackish to marine waters with salinities ranging from 0-35 and will grow in waters with temperatures ranging from 5°C-40°C ([Sampaio et al., 2022](https://www.mdpi.com/1424-2818/14/2/97)).
 Each year, plumes of Saharan dust travel across the Atlantic via easterly trade winds.
@@
 
-Bacteria belonging to the genus Vibrio are marine opportunistic heterotrophs that are ubiquitous in nature and are among the first to respond to the influx of nutrients.
-Species of Vibrio are known to cause a variety of diseases in humans ranging from cholera (V.
-cholerae), to ear and wound infections (V.
-alginolyticus), and in some extreme cases necrotizing fasciitis and septicemia (V.
-vulnificus).
-Vibrio have also been shown to impact marine health, causing coral bleaching (V.
-coralliiyticus, V.
-shiloi, V.
-alginolyticus), and shellfish and fish mortality (V.
-harveyi, V.
+*Vibrio* species are excellent opportunists, having multiple advantages that allow them to quickly respond to any newly introduced substrate or nutrients.
+These advantages include multiple copies of rRNA genes, a rapid doubling time, efficient chemotactic motility, large genomic repertoire (consisting of two circular chromosomes), and multiple Fe-siderophore complexes that allow for rapid uptake of iron into the cell ([Jensen et al., 2009](https://academic.oup.com/femsle/article/294/2/207/512983), [Ringgaard et al., 2018](https://journals.asm.org/doi/10.1128/JB.00793-17), [Eagon, 1962](https://journals.asm.org/doi/abs/10.1128/jb.83.4.736-737.1962?casa_token=WdtRNnQWyhkAAAAA:Zre66xvwgEu3wXQykenhnSoy3OA08h1u-4Zl31k9cxxxfV6jq46dzL18Y02OmwPHdY7p7tgH0wVN-4I), [Okada et al., 2005](https://journals.asm.org/doi/10.1128/JB.187.2.752-757.2005), [Payne et al., 2015](https://journals.asm.org/doi/full/10.1128/MMBR.00046-15)).
+
+Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens.
+Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming.
+The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V.
+cholerae*), shellfish-induced gastroenteritis (*V.
 parahaemolyticus, V.
-vulnificus).
-Vibrio population dynamics are primarily driven by temperature and salinity, with prime conditions between 25-30°C and salinities ranging from 20-35.
-While there is evidence of these blooms occurring in low-nutrient settings like the Florida Keys, less is known regarding microbial response to dust input in areas with *higher baseline nutrients*.
+vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V.
+vulnificus*), and wound and ear infections  (*V.
+alginolyticus*, *V.
+vulnificus*) ([Baker-Austin et al., 2018](https://www.nature.com/articles/s41572-018-0005-8)).
+Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
+tubiashii, V.
+coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V.
+coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [Ben-Haim et al., 2003](https://pubmed.ncbi.nlm.nih.gov/12656189/)), and mortality in important aquaculture species such as penaeid shrimp (*V.
+parahaemolyticus*) and sea breams (*V.
+harveyi)* ([Tran et al., 2013](https://www.int-res.com/abstracts/dao/v105/n1/p45-55/), [Haldar et al., 2010](https://www.sciencedirect.com/science/article/pii/S0944501309001177)), among others.
+
+Previous work has characterized *Vibrio* blooms in response to Saharan dust input in the oligotrophic setting of the Caribbean and subtropical Atlantic (Barbados and Florida Keys, respectively), with surface water concentrations of total *Vibrio* increasing by five to thirty times that found during non-dust conditions and returning to baseline levels within 24-48 hours (Fig.
+2).
+*Vibrio* composition within the larger microbial community also shifted following dust deposition, with initial levels of \<1.4% to a peak of 19.8% of the bacterial community ([Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113)).
+A similar phenomenon was also observed in the surface waters of the tropical and open ocean mid-Atlantic with *Vibrio* populations increasing 1.5-fold in the mid-Atlantic following deposition ([Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
+In a follow up study in the Florida Keys, episodic dust events during a daily time series promoted a succession of bacterial responses, with declines in *Prochlorococcus* coinciding with initial increases in bacteria belonging to the order *Vibrionales*, followed by subsequent shifts in response of different bacterial groups ([Borchardt et al., 2019](https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.11291)).
+To date, our understanding of *Vibrio* and other microbial response to dust input has focused on oligotrophic settings, where dust input is considered to be a critical source of limiting nutrients.
+However, the composition of desert dust can be complex and may have the potential to deliver critical resources that could be exploited by opportunistic microbes (like *Vibrio*) and elicit a growth response despite higher baseline nutrient levels.
+Given that human exposure is likely in these coastal waters, it is important to elucidate the relationship between dust input and microbial blooms in these higher nutrient environments.
+
+This study aims to (1) quantify *Vibrio* population dynamics and composition, and (2) characterize potential microbial blooms and community shifts in response to dust input in coastal sites with higher ambient nutrient levels.
+The findings from this study will provide an increased understanding of the conditions that can elicit potentially harmful blooms (i.e., *Vibrio*), highlighting the need for further research to understand the effects of dust deposition in non-oligotrophic waters.
+We predict that (1) *Vibrio* will be among the first to respond to the influx of dust constituents and (2) that the fertilization effects of dust deposition (and possible successional changes) will be dampened at sites with higher baseline nutrient levels, as dust derived nutrients become less critical for growth.
+
+With a heightened risk of *Vibrio* exposure in coastal waters, it is vital to understand microbial dynamics in nearshore environments.
+If dust is proven to elicit a microbial response, these events can be used to serve as an "early warning" of increased exposure risk to better protect public health.
 
@@
 
-## Data acquisition
+## Description of Study Area
 
-Data was collected at 3 sites in Corpus Christi, TX before, during, and after a Saharan dust event (starting on July 7th and ending on July 19th, 2022).
-Sites represent a gradient in background nutrient levels (Blind Oso and Canals are high, the Gulf is low).
-Water samples were collected for inorganic nutrients, dissolved and particulate organic matter, and microbial analysis.
+Sampling took place at three locations in Corpus Christi, TX: Blind Oso Bay, a residential canal system on Padre Island, and the Gulf.
+Sites were chosen based on differing baseline nutrient levels.
+Blind Oso Bay is a shallow tributary that is popular amongst wadefisherman and kayakers.
+The Bay is often impacted by fluctuating salinities and high levels of nutrients and chlorophyl due to its proximity to a wastewater treatment plant ([Wetz, 2014](https://www.cbbep.org/manager/wp-content/uploads/1312-Oso-Bay-Nutrient-Assessment-Final_Report-6-24-14.pdf)).
+Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
+impaired waters list ([Nicolau & Hill, 2014](https://www.tceq.texas.gov/downloads/water-quality/tmdl/oso-bay-oyster-harvesting-assessment-103/103-oso-bay-monitoring-report-2013.pdf), [Texas Commission on Environmental Quality, 2022](https://www.tceq.texas.gov/downloads/water-quality/assessment/integrated-report-2022/2022-303d.pdf)).
+The residential canal system is also impacted by high nutrients and chlorophyl from storm water runoff, and salinity is primarily driven by precipitation.
+The Gulf site is characterized by constant salinities (\~36) and low nutrient levels.
+
+## Sample Collection and Processing
+
+#### Collection
+
+A high resolution (daily) time series was conducted starting on July 7th and ending on July 19th, 2022 capturing before, during, and after a Saharan dust event. <!--# Add protocol -->
+
 Vertical profiles of salinity, temperature, pH, and dissolved oxygen were obtained using a YSI ProPlus sonde.
+Samples were collected for chlorophyl *a* analysis, inorganic nutrients, dissolved and particulate organic matter, and microbial analysis.
 
-Variables will include `Temperature (°C)`, `Salinity`, `Nitrate`, `Phosphate`, `Copies/mL` produced using quantitative PCR, and `Dust Concentration (AOT)` from the naval research laboratory satellite data.
+#### Vibrio Enumeration
+
+Total *Vibrio* were estimated using genus-specific quantitative polymerase chain reaction (qPCR) ([Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
+<!--# Add specific primers, conditions, etc  -->
 
@@
 More details can be found in the supplementary files, containing the code and comments describing what each line does.
-Please refer to the `project_README.md` file located in the NathanGreenslit-MADA-Project folder for details on the repository contents and instructions on reproducibility
+Please refer to the `project_README.md` file located in the NathanGreenslit-MADA-Project folder for details on the repository contents and instructions on reproducibility\
 
-### Cleaning of Total Vibrio Enumerations Data set
+#### Cleaning of Total Vibrio Enumeration dataset
 
@@
 
-### Cleaning of Dust Concentration (AOT) Data set
+#### Cleaning of Dust Concentration (AOT) Data set
 
@@
 
-### Cleaning of Temperature and Salinity Data set
+#### Cleaning of Environment data set
 
@@
 
-### Dust
+#### Dust
 
@@
 
-### Daily Vibrio
+#### Daily Vibrio
 
@@
 
-### Temperature and Salinity
+#### Temperature & Salinity
 
@@
 
+#### Other Environmental Parameters
+
 <!--# Let's add Nutrient of interest Data -->
@@
 
-<!--# Add distribution statement -->
+Shapiro-Wilk's tests were conducted to assess distribution of the data. Any data that did not have a normal distribution was log-transformed and the test was run again.
 
@@
 
-![](images/00003f.png){width="475"}
+![](images/cca_bo.jpg){width="203"}
 
-![](images/000043.png){width="473"}
+![](images/cca_c2.jpg){width="203"}
 
-![](images/000047.png){width="473"}
+![](images/cca_rd.jpg){width="203"}
 
@@
 
-<!--# Need to update these plots -->
-
-![](images/00000f.png){width="382"}
-
-![](images/00000b.png){width="381"}
-
-![](images/000013.png){width="381"}
+![](images/lm.png){width="329"}
 

### 2023-04-05 products/manuscript/Manuscript.qmd
work cited

@@
     highlight-style: github
+bibliography: MADA.bib
 ---
@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year ([d'Almeida, 1986](https://journals.ametsoc.org/view/journals/apme/25/7/1520-0450_1986_025_0903_amfsdt_2_0_co_2.xml)).
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeida1986].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig.
@@
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
-Dust aerosols can harbor a wide diversity of bacteria ([Kellogg et al., 2004](https://link.springer.com/content/pdf/10.1023/B:AERO.0000032947.88335.bb.pdf)), fungi ([Ramírez Camejo et al., 2022](https://link.springer.com/article/10.1007/s10453-022-09752-9)), virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)).
+Dust aerosols can harbor a wide diversity of bacteria [@kellogg2004], fungi [@ramirez2022], virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)).
 These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron ([Graham & Duce, 1967](https://www.sciencedirect.com/science/article/pii/0004698182901986), [Mills et al., 2004](https://pubmed.ncbi.nlm.nih.gov/15152251/), [Savoie et al., 1989](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/JD094iD04p05069?casa_token=PMRaet9QSYsAAAAA:nb8yBorxdKaEWwGfI5aNILJ0D4h6drNu2YY1y1y8ExbQ03z59EUTQTOHoxinzsxRtR4jzsfSA73_Hig)).
@@
 alginolyticus*, *V.
-vulnificus*) ([Baker-Austin et al., 2018](https://www.nature.com/articles/s41572-018-0005-8)).
+vulnificus*) [@baker2018].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V.
-coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [Ben-Haim et al., 2003](https://pubmed.ncbi.nlm.nih.gov/12656189/)), and mortality in important aquaculture species such as penaeid shrimp (*V.
+coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [@haim2003]), and mortality in important aquaculture species such as penaeid shrimp (*V.
 parahaemolyticus*) and sea breams (*V.

### 2023-04-05 products/manuscript/Manuscript.qmd
work cited

@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeida1986].
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig.
@@
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
-Dust aerosols can harbor a wide diversity of bacteria [@kellogg2004], fungi [@ramirez2022], virus-like particles ([Griffin et al., 2001](https://link.springer.com/article/10.1023/A:1011868218901)), and minerals ([Formenti et al., 2003](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2002JD002648)).
-These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron ([Graham & Duce, 1967](https://www.sciencedirect.com/science/article/pii/0004698182901986), [Mills et al., 2004](https://pubmed.ncbi.nlm.nih.gov/15152251/), [Savoie et al., 1989](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/JD094iD04p05069?casa_token=PMRaet9QSYsAAAAA:nb8yBorxdKaEWwGfI5aNILJ0D4h6drNu2YY1y1y8ExbQ03z59EUTQTOHoxinzsxRtR4jzsfSA73_Hig)).
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575), [Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
+Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004], fungi [@ramirez-camejoFungalDiversitySahara2022], virus-like particles [@griffinAfricanDesertDust], and minerals [@formenti2003].
+These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004 ; @savoieNonseasaltSulfateNitrate1989 ; @graham1982] .
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups [@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
 
 Previous work conducted in the West Florida Shelf has shown that increased iron availability can stimulate the growth of the nitrogen-fixing cyanobacterium *Trichodesmium*.
-The high amount of nitrogen produced via fixation can be adequate enough to stimulate toxic red tides, caused by the dinoflagellate *Karenia brevis* ([Lenes et al., 2008](https://www.sciencedirect.com/science/article/pii/S0278434308000575)).
+The high amount of nitrogen produced via fixation can be adequate enough to stimulate toxic red tides, caused by the dinoflagellate *Karenia brevis* [@lenesSaharanDustPhosphatic2008].
 
@@
 Among those that respond are bacterium belonging to the genus *Vibrio*, a group of ubiquitous marine opportunistic heterotrophs.
-While considered conditionally rare (comprising of \<1% of the microbial community), under the proper conditions, this genus can rapidly bloom to make up a large percentage of the community over a short period of time ([Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113), [Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
-*Vibrio* species are found in brackish to marine waters with salinities ranging from 0-35 and will grow in waters with temperatures ranging from 5°C-40°C ([Sampaio et al., 2022](https://www.mdpi.com/1424-2818/14/2/97)).
+While considered conditionally rare (comprising of \<1% of the microbial community), under the proper conditions, this genus can rapidly bloom to make up a large percentage of the community over a short period of time [@westrichSaharanDustNutrients2016a ; @westrichVibrioPopulationDynamics2018a].
+*Vibrio* species are found in brackish to marine waters with salinities ranging from 0-35 and will grow in waters with temperatures ranging from 5°C-40°C [@sampaioVibrioSppLife2022].
 Each year, plumes of Saharan dust travel across the Atlantic via easterly trade winds.
@@
 *Vibrio* species are excellent opportunists, having multiple advantages that allow them to quickly respond to any newly introduced substrate or nutrients.
-These advantages include multiple copies of rRNA genes, a rapid doubling time, efficient chemotactic motility, large genomic repertoire (consisting of two circular chromosomes), and multiple Fe-siderophore complexes that allow for rapid uptake of iron into the cell ([Jensen et al., 2009](https://academic.oup.com/femsle/article/294/2/207/512983), [Ringgaard et al., 2018](https://journals.asm.org/doi/10.1128/JB.00793-17), [Eagon, 1962](https://journals.asm.org/doi/abs/10.1128/jb.83.4.736-737.1962?casa_token=WdtRNnQWyhkAAAAA:Zre66xvwgEu3wXQykenhnSoy3OA08h1u-4Zl31k9cxxxfV6jq46dzL18Y02OmwPHdY7p7tgH0wVN-4I), [Okada et al., 2005](https://journals.asm.org/doi/10.1128/JB.187.2.752-757.2005), [Payne et al., 2015](https://journals.asm.org/doi/full/10.1128/MMBR.00046-15)).
+These advantages include multiple copies of rRNA genes, a rapid doubling time, efficient chemotactic motility, large genomic repertoire (consisting of two circular chromosomes), and multiple Fe-siderophore complexes that allow for rapid uptake of iron into the cell [@jensenNonrandomMicroheterogeneity16S2009; @ringgaardChemotaxisArraysVibrio2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962 ; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
 
@@
 alginolyticus*, *V.
-vulnificus*) [@baker2018].
+vulnificus*) [@baker-austinVibrioSppInfections2018].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V.
-coralliilyticus*) ([Richards et al., 2015](https://pubmed.ncbi.nlm.nih.gov/25344234/), [Rosenberg & Falkovitz, 2004](https://www.annualreviews.org/doi/10.1146/annurev.micro.58.030603.123610), [@haim2003]), and mortality in important aquaculture species such as penaeid shrimp (*V.
+coralliilyticus*) [@richardsMortalitiesEasternPacific2015 ; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003], and mortality in important aquaculture species such as penaeid shrimp (*V.
 parahaemolyticus*) and sea breams (*V.
-harveyi)* ([Tran et al., 2013](https://www.int-res.com/abstracts/dao/v105/n1/p45-55/), [Haldar et al., 2010](https://www.sciencedirect.com/science/article/pii/S0944501309001177)), among others.
+harveyi)* [@tranDeterminationInfectiousNature2013 ; @haldarIdentificationVibrioHarveyi2010] among others.
 
@@
 2).
-*Vibrio* composition within the larger microbial community also shifted following dust deposition, with initial levels of \<1.4% to a peak of 19.8% of the bacterial community ([Westrich et al., 2016](https://www.pnas.org/doi/abs/10.1073/pnas.1518080113)).
-A similar phenomenon was also observed in the surface waters of the tropical and open ocean mid-Atlantic with *Vibrio* populations increasing 1.5-fold in the mid-Atlantic following deposition ([Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
-In a follow up study in the Florida Keys, episodic dust events during a daily time series promoted a succession of bacterial responses, with declines in *Prochlorococcus* coinciding with initial increases in bacteria belonging to the order *Vibrionales*, followed by subsequent shifts in response of different bacterial groups ([Borchardt et al., 2019](https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.11291)).
+*Vibrio* composition within the larger microbial community also shifted following dust deposition, with initial levels of \<1.4% to a peak of 19.8% of the bacterial community [@westrichSaharanDustNutrients2016a].
+A similar phenomenon was also observed in the surface waters of the tropical and open ocean mid-Atlantic with *Vibrio* populations increasing 1.5-fold in the mid-Atlantic following deposition [@westrichVibrioPopulationDynamics2018a] In a follow up study in the Florida Keys, episodic dust events during a daily time series promoted a succession of bacterial responses, with declines in *Prochlorococcus* coinciding with initial increases in bacteria belonging to the order *Vibrionales*, followed by subsequent shifts in response of different bacterial groups [@borchardtSaharanDustDeposition2020b].
 To date, our understanding of *Vibrio* and other microbial response to dust input has focused on oligotrophic settings, where dust input is considered to be a critical source of limiting nutrients.
@@
 Blind Oso Bay is a shallow tributary that is popular amongst wadefisherman and kayakers.
-The Bay is often impacted by fluctuating salinities and high levels of nutrients and chlorophyl due to its proximity to a wastewater treatment plant ([Wetz, 2014](https://www.cbbep.org/manager/wp-content/uploads/1312-Oso-Bay-Nutrient-Assessment-Final_Report-6-24-14.pdf)).
-Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
+The Bay is often impacted by fluctuating salinities and high levels of nutrients and chlorophyl due to its proximity to a wastewater treatment plant [@wetz] Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
 impaired waters list ([Nicolau & Hill, 2014](https://www.tceq.texas.gov/downloads/water-quality/tmdl/oso-bay-oyster-harvesting-assessment-103/103-oso-bay-monitoring-report-2013.pdf), [Texas Commission on Environmental Quality, 2022](https://www.tceq.texas.gov/downloads/water-quality/assessment/integrated-report-2022/2022-303d.pdf)).
@@
 
-Total *Vibrio* were estimated using genus-specific quantitative polymerase chain reaction (qPCR) ([Westrich et al., 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00012/full)).
-<!--# Add specific primers, conditions, etc  -->
+Total *Vibrio* were estimated using genus-specific quantitative polymerase chain reaction (qPCR) [@westrichSaharanDustNutrients2016a]. <!--# Add specific primers, conditions, etc  -->
 

### 2023-04-05 code/analysis_code/stats/3_uni_variate.qmd
edits

@@
 library(dotwhisker) #Fpr visualization
+library(ggpubr) #For plot stacking
 ```
@@
 
+## Null Model:
+
+#### 5-Fold Cross Validation
+
+```{r}
+fold_bo <- vfold_cv(bo2, v = 5, repeats = 5) #Data was too small to stratify (<20)
+fold_c2 <- vfold_cv(c22, v = 5, repeats = 5) #Data was too small to stratify (<20)
+fold_rd <- vfold_cv(rd2, v = 5, repeats = 5) #Data was too small to stratify (<20)
+```
+
+#### Create Null Model
+
+```{r}
+#Create Null Recipe for Growth at all three sites
+null_bo<- recipe(copies_mL ~ 1, data = bo2) 
+null_c2<- recipe(copies_mL ~ 1, data = c22)
+null_rd<- recipe(copies_mL ~ 1, data = rd2)
+
+#Set up linear model
+lm_mod<- linear_reg() %>%
+  set_engine("lm") %>%
+  set_mode("regression")
+
+#Workflow that adds recipe to model for three sites
+null_wflow_bo<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(null_bo)
+
+null_wflow_c2<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(null_c2)
+
+null_wflow_rd<- 
+  workflow() %>%
+  add_model(lm_mod) %>%
+  add_recipe(null_rd)
+
+#Use workflow to fit model to  data set
+null_fit_bo<-
+  fit_resamples(null_wflow_bo, resamples = fold_bo)
+
+null_fit_c2<-
+  fit_resamples(null_wflow_c2, resamples = fold_c2)
+
+null_fit_rd<-
+  fit_resamples(null_wflow_rd, resamples = fold_rd)
+```
+
+#### Get Metrics for Null
+
+##### RMSE
+
+```{r}
+null_met_bo <- 
+  collect_metrics(null_fit_bo)
+
+null_met_c2 <- 
+  collect_metrics(null_fit_c2)
+
+null_met_rd <- 
+  collect_metrics(null_fit_rd)
+
+null_met_bo
+null_met_c2
+null_met_rd
+```
+
+##### AIC
+
+```{r}
+bo_null_mod<- 
+  lm(copies_mL ~ 1, data = bo2)
+AIC(bo_null_mod)
+
+c2_null_mod<- 
+  lm(copies_mL ~ 1, data = c22)
+AIC(c2_null_mod)
+
+rd_null_mod<- 
+  lm(copies_mL ~ 1, data = rd2)
+AIC(rd_null_mod)
+```
+
+We will come back to these metrics once we run the models on our actual data.
+
+## Actual Data for Model
+
 #### Create Linear Model
@@
 
-### RMSE and R2 to test model performance
+#### Metrics
+
+##### RMSE and R2
 
@@
 
-### AIC to assess Performance
+##### AIC
 
@@
 
-### Plot Residuals
+#### Plot Residuals
 
@@
 ```
+
+## Conclusion
+
+| Site | Model Type | RMSE     | AIC      |
+|------|------------|----------|----------|
+| BO   | Null       | 20472.82 | 275.9331 |
+| BO   | Actual     | 14422.91 | 269.8923 |
+| C2   | Null       | 8472.898 | 253/1876 |
+| C2   | Actual     | 5191.22  | 245.3679 |
+| RD   | Null       | 50386.56 | 296.9821 |
+| RD   | Actual     | 36132.11 | 291.933  |
+
+AIC and RMSE are lower (better performance) than the null.
+
+<!--# Why are these so high? -->
 

### 2023-04-05 products/manuscript/Manuscript.qmd
edits

@@
 
+        -   Why are my RMSE and AIC so high?
+
     -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into.
@@
 
-3.  Looking at my statistics/models, is there anything that you think is strikingly missing?
+3.
+Ultimately, I feel as though I am missing some "oomph"when it comes to my statistical analysis.
+The linear models seem tadequate, but I know that there are multiple variables at play at these sites (and we have the nutrient data for this).
+I am open to any suggestions that may diversify and expand my analyses!
 
@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986].
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig.
@@
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
-Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004], fungi [@ramirez-camejoFungalDiversitySahara2022], virus-like particles [@griffinAfricanDesertDust], and minerals [@formenti2003].
-These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004 ; @savoieNonseasaltSulfateNitrate1989 ; @graham1982] .
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups [@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
+Dust aerosols can harbor a wide diversity of bacteria \[@kelloggCharacterizationAerosolizedBacteria2004\], fungi \[@ramirez-camejoFungalDiversitySahara2022\], virus-like particles \[@griffinAfricanDesertDust\], and minerals \[@formenti2003\].
+These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] .
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
 
@@
 alginolyticus*, *V.
-vulnificus*) [@baker-austinVibrioSppInfections2018].
+vulnificus*) [@baker-austinVibrioSppInfections2018\].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V.
-coralliilyticus*) [@richardsMortalitiesEasternPacific2015 ; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003], and mortality in important aquaculture species such as penaeid shrimp (*V.
+coralliilyticus*) \[@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003\], and mortality in important aquaculture species such as penaeid shrimp (*V.
 parahaemolyticus*) and sea breams (*V.
-harveyi)* [@tranDeterminationInfectiousNature2013 ; @haldarIdentificationVibrioHarveyi2010] among others.
+harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
 

### 2023-04-06 code/analysis_code/env/env_par.qmd
Addedto env data viz

@@
   filter(date %in% c(as.Date("2022-07-07"):as.Date("2022-07-19"))) %>%
+  filter(!date %in% as.Date('2022-07-10')) %>%
   select(!...1)
@@
 
-#### Plotting Daily Temperature
+#### Temperature
 
@@
     y = "Temperature (°C)",
-    title = "Temperature across Daily Time Series by Site",
+  #  title = "Temperature across Daily Time Series by Site",
     color = "Site")+
@@
     axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 15),
-    legend.position = c(0.1,0.2),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
     legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
@@
                                 "Gulf" = "4c848c"))
-
-temp
-
 ```
 
-#### Plotting Daily Salinity
+#### Salinity
 
@@
     y = "Salinity",
-    title = "Salinity across Daily Time Series by Site",
+  #  title = "Temperature across Daily Time Series by Site",
     color = "Site")+
@@
     plot.title= element_text(hjust = 0.5, size = 20),
-    axis.title.x = element_text(size = 15),
-    axis.title.y = element_text(size = 15),
-    axis.text.x = element_text(size = 10),
-    axis.text.y = element_text(size = 10),
-    legend.position = "none")+
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
     scale_color_manual(values = c("Blind Oso" = "#809C52",
@@
                                 "Gulf" = "4c848c"))
-
-sal
 ```
 
-#### Combine Temperature and Salinity Plots
+#### Dissolved Oxygen
 
-```{r, fig.width= 4, fig.height=4}
-fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")
+```{r}
+do<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = do_mgl,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DO (mg/L)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Nitrite:Nitrate
+
+```{r}
+nn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = nn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Nitrate-Nitrite (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Orthophosphate
+
+```{r}
+ortho<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = orthop,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Orthophosphate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DON
+
+```{r}
+don<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = don,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DON (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DOC
+
+```{r}
+doc<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = doc,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DOC (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### pH
+
+```{r}
+pH<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = pH,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "pH",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_text(size = 10),
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Ammonium
+
+```{r}
+amm<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = amm,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Ammonium (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Total Nitrogen
+
+```{r}
+tn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = tn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Total Nitrogen (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Silicate
+
+```{r}
+sil<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = sil,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Silicate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Copies/mL
+
+```{r}
+cop<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = copies_mL,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Vibrio (copies per mL)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 7),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = c(0.2,0.7),
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
+    legend.text = element_text(size = 7)) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Combine Environment Plots
+
+```{r, fig.width= 10, fig.height= 5}
+fig<- ggarrange(cop,temp,sal,sil,do,amm,nn,tn,ortho,doc, pH,don + font("x.text", size = 10), ncol = 4, nrow = 3,align = "v")
 
@@
 
-#### Save as PNG
+# Save as PNG
 

### 2023-04-06 products/manuscript/Manuscript.qmd
Addedto env data viz

@@
 Ultimately, I feel as though I am missing some "oomph"when it comes to my statistical analysis.
-The linear models seem tadequate, but I know that there are multiple variables at play at these sites (and we have the nutrient data for this).
+The linear models seem adequate, but I know that there are multiple variables at play at these sites (and we have the nutrient data for this).
 I am open to any suggestions that may diversify and expand my analyses!
@@
 
-## General Background Information
-
 The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
-Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig.
-1).
+Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico.
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
@@
 
-Previous work has characterized *Vibrio* blooms in response to Saharan dust input in the oligotrophic setting of the Caribbean and subtropical Atlantic (Barbados and Florida Keys, respectively), with surface water concentrations of total *Vibrio* increasing by five to thirty times that found during non-dust conditions and returning to baseline levels within 24-48 hours (Fig.
-2).
+Previous work has characterized *Vibrio* blooms in response to Saharan dust input in the oligotrophic setting of the Caribbean and subtropical Atlantic (Barbados and Florida Keys, respectively), with surface water concentrations of total *Vibrio* increasing by five to thirty times that found during non-dust conditions and returning to baseline levels within 24-48 hours.
 *Vibrio* composition within the larger microbial community also shifted following dust deposition, with initial levels of \<1.4% to a peak of 19.8% of the bacterial community [@westrichSaharanDustNutrients2016a].
@@
 
-{{< pagebreak >}}
-
 # Methods
@@
 
-A high resolution (daily) time series was conducted starting on July 7th and ending on July 19th, 2022 capturing before, during, and after a Saharan dust event. <!--# Add protocol -->
+A high resolution (daily) time series was conducted starting on July 7th and ending on July 19th, 2022 capturing before, during, and after a Saharan dust event.
+Samples were collected in 1L autoclaved Polypropylene Bottles and immediately placed in a cooler on ice.
+Following transport to the laboratory, the samples were concentrated onto 0.2µm pore size, hydrophobic polycarbonate membranes (25mm in diameter) and stored at -80°C until DNA Extraction using a ZymoBIOMICS DNA Miniprep Kit (Cat#:4300).
 
@@
 
-Total *Vibrio* were estimated using genus-specific quantitative polymerase chain reaction (qPCR) [@westrichSaharanDustNutrients2016a]. <!--# Add specific primers, conditions, etc  -->
+Total *Vibrio* concentrations were quantified using a SYBR Green quantitative PCR (qPCR) method.
+Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) PCR product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3.
+Master mix and PCR conditions were derived from [@westrichSaharanDustNutrients2016a].
+Briefly, 5µL of 2X SYBR PowerUp Master Mix (Applied Biosystems, Foster City, CA) were added to primers with a final concentration of 0.16µM and PCR water for a final volume of 10µL.
+Reactions were run in triplicate on a CFX96 Touch Real-Time PCR Detection System (Bio-Rad Laboratories, Hercules, CA) with the following cycling conditions: 2 minutes at 50◦C for UDG activation and 95◦C for 2 min to activate AmpliTaq polymerase and UP, followed by 40 cycles of 95◦C for 3 seconds for denaturation and 60◦C for 30 seconds for annealing and extension.
+Each run was followed by a dissociation step (60C to 95C by 0.5C increments) to determine a melt curve for analysis of specificity.
+Each test was run with positive and negative controls in triplicate.
+Cycle threshold values for each qPCR test was compared to a standard curve representing 10\^1 to 10\^6 gene copies per reaction volume.
+<!--# Add Std Curve specs -->
 
@@
 
-#### Dust
-
-![](images/paste-0FE55FAA.png)
-
-This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series.
-During the time series, two small periods of higher dust AOT occurred on the 9th and 13th respectively.
-A much larger spike in AOT was observed on the 16th.
-
-#### Daily Vibrio
+#### Dust AOD and Vibrio Enumeration
 
@@
 There is a noticeable shift in copies in the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
-The bottom portion of the figure depicts the dust concentration (AOT) across the daily time series.
+The bottom portion of the figure depicts the dust concentration (AOD) across the daily time series.
+This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series.
+During the time series, two small periods of higher dust AOD occurred on the 9th and 13th respectively.
+A much larger spike in AOD was observed on the 16th.
 Initial shift in copies per mL at the Gulf and Blind Oso occur 24-48hr following initial dust introduction on the 13th, and experience another shift around 24-48hr of higher dust input on the 16th.
 
-#### Temperature & Salinity
+#### Environmental Parameters
 
-![](images/temp_sal.png)
+![](images/000031.png)
 
@@
 
-![](images/00000b-01.png){width="406"}
+![](images/000005-02.png){width="342"}
 
-![](images/00000f-01.png){width="404"}
+![](images/000009.png){width="341"}
 
-![](images/000009%20(1)-01.png){width="404"}
+![](images/00000d.png)
 

### 2023-04-06 code/analysis_code/stats/3_uni_variate.qmd
Addedto env data viz
Cor Plots

@@
 | BO   | Actual     | 14422.91 | 269.8923 |
-| C2   | Null       | 8472.898 | 253/1876 |
+| C2   | Null       | 8472.898 | 253.1876 |
 | C2   | Actual     | 5191.22  | 245.3679 |

### 2023-04-06 code/analysis_code/stats/cor.qmd
Cor Plots

---
title: "cor"
format:
  html:
    theme: default
---

**This code creates a correlation matrix to look at relationships between variables.**

[**THIS CODE IS STILL IN THE WORKS**]{.ul}

Link: http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(Hmisc)#Cor Tables
library(corrplot) #Cor Plots
```

#### Data

```{r, include = FALSE}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL and nutrients 
```

# Cleaning

### Clean and Trim

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>% #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
  select(!c(date,hr_cst,...1,Type, t1,t7,t13,t19,tsum,tavg,SUM,t7sum,t7avg))

```

#### Site DF

```{r}
bo<- 
  master0 %>%
  filter(site %in% "Blind Oso") %>%
  select(!site)

c2<- 
  master0 %>%
  filter(site %in% "Canals") %>%
  select(!site)

rd<- 
  master0 %>%
  filter(site %in% "Gulf") %>%
  select(!site)
```

# Correlation Matrix

#### Compute Correlation Matrix

```{r}
b<- 
  cor(bo)
c<- 
  cor(c2)
r<- cor(rd)
```

#### Make Matrix

```{r}
bo2<- rcorr(as.matrix(bo))
c22<-

```

#### Function to Format

```{r}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

#### Look at Matrix again

```{r}
b<- rcorr(as.matrix(bo))
flattenCorrMatrix(b$r, b$P)
```

#### Visualize

```{r}
corrplot(b, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(c, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

corrplot(r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

### 2023-04-06 products/manuscript/Manuscript.qmd
Cor Plots

@@
 
-    -   Script Flow for Dust: Plot Dust and Copies --> Create Model --> Assess Performance (AIC/RMSE/R2) --> Plot Residuals.
-
     -   Why are my RMSE and AIC so high?
@@
 Ultimately, I feel as though I am missing some "oomph"when it comes to my statistical analysis.
-The linear models seem adequate, but I know that there are multiple variables at play at these sites (and we have the nutrient data for this).
+The linear models seem adequate, but I know that there are multiple variables at play at these sites (and we have data to explore some of this).
 I am open to any suggestions that may diversify and expand my analyses!
@@
 The Gulf site is characterized by constant salinities (\~36) and low nutrient levels.
+<!--# update these citations -->
 
@@
 Blind Oso Bay experiences some shifts in temperatures and drastic fluctuations in salinity.
-The Gulf site had a drop in temperature (possibly due to upwelling) and relatively consistent ocean salinities
+The Gulf site had a drop in temperature (possibly due to upwelling) and relatively consistent ocean salinities.
 
-#### Other Environmental Parameters
-
-<!--# Let's add Nutrient of interest Data -->
+<!--# Add detail on other env variables -->
 
@@
 As mentioned above, the lags identified may not correspond to the most significant relationship between the two variables.
-While a lag of -1 for Blind Oso and -2 for the Canals explains the data well, a lag of -2 for the Gulf site was more fitting.
-This makes sense, as taking a look at the copies_mL plot above, we do see a \~2 day lag in response, where we have a spike in dust, and 2 days later, a spike in copies_mL.
+
+| Site | Model Type | RMSE     | AIC      |
+|------|------------|----------|----------|
+| BO   | Null       | 20472.82 | 275.9331 |
+| BO   | Actual     | 14422.91 | 269.8923 |
+| C2   | Null       | 8472.898 | 253.1876 |
+| C2   | Actual     | 5191.22  | 245.3679 |
+| RD   | Null       | 50386.56 | 296.9821 |
+| RD   | Actual     | 36132.11 | 291.933  |
 
@@
 actual will give an R2 of 1.00.
-I am continuing to find a good validation method.
-
-    -   After reviewing module 9, it looks like Information Criteria may be a proper alternative (REF: MOD9/Model Performance and Overfitting).
+I am using AIC as a metric of model performance and comparing data produced models to null models.
 
@@
 
-        -   AIC:
-
-            -   Blind Oso = 271.25
-
-            -   Canals = 243.97
-
-            -   Gulf = 292.13
-
--   I am also planning on running a multi-variate model to look at multiple predictors of interests (temp, salinity, nutrients), but am currently running into an issue with my model output, where rather than getting values, I receive NaN.
-This code can be found in my `dust_copies_stats.qmd` code.
+-   I am also planning on running a multi-variate model to look at multiple predictors of interests (temp, salinity, nutrients). This code can be found in my `dust_copies_stats.qmd` code.
 
@@
 
-![](images/000005-02.png){width="342"}
+![](images/000005-02.png){width="299"}
 
-![](images/000009.png){width="341"}
+![](images/000009.png){width="300"}
 
-![](images/00000d.png)
+![](images/00000d.png){width="296"}
 

### 2023-04-07 code/analysis_code/analysis_README.Rmd
Worked with R2 pearson and wrapped up Correlation plots

--- 2023-03-16 code/analysis_code/analysis_README.Rmd
@@
 (3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
-(4) The stats folder contains `dust_copies_stats.qmd` that runs models on variables of interest. \
+(4) The stats folder contains:
+- `1_distributiom.qmd`: Checks distribution of variables of interest
+- `2_cca.qmd`: Performs cross-correlation analysis to test for lags in time series (between dust input and Vibrio growth)
+- `3_correlation.qmd`: Runs Correlation matrices on all variables
+- `4_uni_variate.qmd` : runs univariate models on variable and predictors of interest.
+- `5_multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)

### 2023-04-07 code/analysis_code/stats/3_correlation.qmd
Worked with R2 pearson and wrapped up Correlation plots

---
title: "cor"
format:
  html:
    theme: default
---

**This code creates a correlation matrix to look at relationships between variables. This helps identify specific relationships of interest**

[**THIS CODE IS STILL IN THE WORKS**]{.ul}

[Link](http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software):

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(Hmisc)#Cor Tables and Lag
library(corrplot) #Cor Plots
library(lubridate) #For dates
```

#### Data

```{r, include = FALSE}
dust0<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust permutations
data0<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data containing copies_mL and nutrients 
```

# Cleaning

### Trim data sets from 7/4 to 7/19

```{r}
dust<- dust0 %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normally distributed)

data<- data0 %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

### Combine into one big data set

```{r}
master0<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") %>% #While this is technically not for hour 1, this gets ride of all   duplicates for the sake of analysis.
  select(!c(hr_cst,...1,Type, t1,t7,t13,t19,tsum,tavg,SUM,dust,t7avg)) %>%
  rename("dust" = "t7sum") #Now dust variable is the t7sum
```

#### Make Lag data sets by Site

This cleans out non-numeric values, creates site lag from CCA, and producea a final daily plot

```{r}
bo<- 
  master0 %>%
  filter(site %in% "Blind Oso"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 1)) %>% #Lag from cca.qmd
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

c2<- 
  master0 %>%
  filter(site %in% "Canals"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 2)) %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

rd<- 
  master0 %>%
  filter(site %in% "Gulf"| date %in% c( as.Date("2022-07-05"), as.Date("2022-07-06"))) %>%
  select(!site) %>% 
  mutate(dust =Lag(dust, shift  = 1)) %>%
  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
  select(!date)

```

# Correlation Matrix

There are different methods for **correlation analysis** : **Pearson parametric correlation test**, **Spearman** and **Kendall** rank-based **correlation analysis**. The default is [pearson correlation coefficient](http://www.sthda.com/english/wiki/correlation-test-between-two-variables) which measures the linear **dependence** between two variables. [kendall and spearman](http://www.sthda.com/english/wiki/correlation-test-between-two-variables) correlation methods are non-parametric **rank-based correlation test**.

#### Compute Correlation Matrix

If your data contain missing values, use the following R code to handle missing values by case-wise deletion.

    cor(x, method = "pearson", use = "complete.obs")

```{r, include = FALSE}
b<- cor(bo)
round(b,2) #Round correlation coeff to "x" digits

c<- cor(c2)
round(c,2)

r<- cor(rd)
round(r,2)
```

#### Make Function to Format

Simple function for formatting a **correlation matrix** into a table with 4 columns containing : Column 1 : row names (variable 1 for the correlation test), Column 2 : column names (variable 2 for the correlation test), Column 3 : the **correlation coefficients**, Column 4 : the **p-values** of the correlations

```{r}
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut])}
```

#### Make Matrix

The function **rcorr()** \[in **Hmisc** package\] can be used to compute the **significance levels** for **pearson** and **spearman correlations**. It returns both the correlation coefficients and the p-value of the correlation for all possible pairs of columns in the data table.

```{r}
bclean<- rcorr(as.matrix(bo))
flattenCorrMatrix(bclean$r, bclean$P)

cclean<- rcorr(as.matrix(c2))
flattenCorrMatrix(cclean$r, cclean$P)

rclean<- rcorr(as.matrix(rd))
flattenCorrMatrix(rclean$r, rclean$P)
```

#### Visualize with `corrplot()` function

The function **corrplot()** takes the **correlation matrix** as the first argument. The second argument (type="upper") is used to display only the upper triangular of the **correlation matrix**.

```{r}
corrplot(b, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Blind Oso",  mar=c(0,0,1,0)) #Adds and lowers title

corrplot(c, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Canals", mar=c(0,0,1,0))

corrplot(r, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45,
         title = "Gulf", mar=c(0,0,1,0))
```

It's also possible to **combine correlogram with the significance test**. We'll use the result *res.cor2* generated in the previous section with **rcorr**() function \[in **Hmisc** package\]:

<!--# Get this to work -->

```{r}
# Insignificant correlation are crossed
#corrplot(bclean$r, type="upper", order="hclust", 
        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
# Insignificant correlations are leaved blank
#corrplot(bclean$r, type="upper", order="hclust", 
        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
```

### 2023-04-07 code/analysis_code/stats/4_uni_variate.qmd
Worked with R2 pearson and wrapped up Correlation plots

--- 2023-04-06 code/analysis_code/stats/3_uni_variate.qmd
@@
 ```{r}
-bo2<- bo%>% mutate(
+bo2<- bo %>% mutate(
   lag =Lag(bo$t7sum, shift  = 1)) %>% #Make lag from ccf
@@
   stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
+  stat_correlation(method = "pearson")+
   labs(title = "LM DustxCopies (BO+1)",
@@
   stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
+  stat_correlation(method = "pearson")+
   labs(title = "LM DustxCopies (C2+2)",
@@
   stat_smooth(method = "lm")+
-  stat_poly_eq(aes(x = lag, y = copies_mL),
-               rr.digits = 4) +
+  stat_correlation(method = "pearson")+
   labs(title = "LM DustxCopies (RD+2)",
@@
 
-#### Stack and Save
+#### Stac
 
 ```{r}
-png(file = here("results", "plots", "stats", "lm.png"))
 ggarrange(fig1, fig2, fig3 + font("x.text", size = 10), ncol = 1, nrow = 3)
-dev.off()
 ```
@@
 <!--# Is there a lag for Nutrient data too? -->
+
+<!--# Need to change R2 input here too -->
 

### 2023-04-07 code/analysis_code/stats/5_multi_variate.qmd
Worked with R2 pearson and wrapped up Correlation plots

--- 2023-04-03 code/analysis_code/stats/4_multi_variate.qmd
@@
 
-### Libraries
+# Getting Started
+
+#### Libraries
 
@@
 
-### Clean and Trim
+# Cleaning
+
+#### Clean and Trim
 
@@
 
-### Combine into one big data set
+#### Make into one data set
 
@@
 mutate(
-  lag =Lag(rd_a$t7sum, shift  = 2)) %>% #Make 2 day lag
+  lag =Lag(rd_a$t7sum, shift  = 1)) %>% #Make 2 day lag
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19')))%>%

### 2023-04-07 products/manuscript/Manuscript.qmd
Worked with R2 pearson and wrapped up Correlation plots

@@
 
-    -   Why are my RMSE and AIC so high?
+    -   Why are my RMSE and AIC so high (see table under *Conclusion* section)?
 
@@
 I am open to any suggestions that may diversify and expand my analyses!
+
+4.
+In my `code/analysis_code/stats/3_uni_variate.qmd` code I was playing around with R2 for my linear models.
+Previously I have been plotting R2 onto the graphs using `stat_poly_equation()` but then I came across the `stat_correlation()` command.
+When specifying "pearson" as the method, I get a MUCH higher R2 value for the models (\~0.48 to \~0.70).
+I have been trying to figure out what `stat_poly_equation` uses as its method and why there is such a difference.
+At least with dust and copies_mL, I expect Pearson's to be a good method.
+But is there some serious error that I am making here?
 
@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year \[@dalmeidaModelSaharanDust1986\].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico.
@@
 These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] .
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a\].
 
@@
 alginolyticus*, *V.
-vulnificus*) [@baker-austinVibrioSppInfections2018\].
+vulnificus*) \[@baker-austinVibrioSppInfections2018\].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 parahaemolyticus*) and sea breams (*V.
-harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
+harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010\] among others.
 
@@
 
-![](images/lm.png){width="329"}
+![](images/Screen%20Shot%202023-04-06%20at%2010.23.28%20PM.png)
 

### 2023-04-07 code/analysis_code/stats/4_uni_variate.qmd
edited questions

@@
   stat_correlation(method = "pearson")+
-  labs(title = "LM DustxCopies (RD+2)",
+  labs(title = "LM DustxCopies (RD+1)",
        x = "log(dust)")
@@
 rd_metrics
+
 ```

### 2023-04-07 products/manuscript/Manuscript.qmd
edited questions

@@
 
-1.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
+1.
+In my `code/analysis_code/stats/3_uni_variate.qmd` code I was playing around with R2 for my linear models.
+Previously I have been plotting R2 onto the graphs using `stat_poly_equation()` and have also checked through `tidymodeling`.
+But then I came across the `stat_correlation()` command.
+When specifying "pearson" as the method, I get a **MUCH** higher R2 value for the models (\~0.48 to \~0.70).
+I have been trying to figure out what `stat_poly_equation` uses as its method and why there is such a difference.
+Additionally, my actual model produces the same lower R2 values (using `rsq()`) as the `stat_poly_equation()`.
+Is there some serious error that I am making here?
+
+2.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
 
@@
 
-    -   If this looks good, I plan to do the same workflow for all of the environmental parameters as well. For now, I just have plots to see what may be worth diving into.
+    -   If this looks good, I plan to do the same workflow for all of the environmental parameters (that seem to have a relationship) as well.
+For now, I just have plots to see what may be worth diving into.
 
-2.  Looking at `code/analysis_code/stats/4_multi_variate.qmd` , I am a little stuck on how to best create a multivariate model.
+3.  Looking at `code/analysis_code/stats/4_multi_variate.qmd` , I am a little stuck on how to best create a multivariate model.
 
@@
 
-3.
-Ultimately, I feel as though I am missing some "oomph"when it comes to my statistical analysis.
+4.
+Ultimately, I feel as though I am missing some "oomph" when it comes to my statistical analysis.
 The linear models seem adequate, but I know that there are multiple variables at play at these sites (and we have data to explore some of this).
 I am open to any suggestions that may diversify and expand my analyses!
-
-4.
-In my `code/analysis_code/stats/3_uni_variate.qmd` code I was playing around with R2 for my linear models.
-Previously I have been plotting R2 onto the graphs using `stat_poly_equation()` but then I came across the `stat_correlation()` command.
-When specifying "pearson" as the method, I get a MUCH higher R2 value for the models (\~0.48 to \~0.70).
-I have been trying to figure out what `stat_poly_equation` uses as its method and why there is such a difference.
-At least with dust and copies_mL, I expect Pearson's to be a good method.
-But is there some serious error that I am making here?
+I am thinking some sort of PCA that depicts which predictor variables have the greatest influence on the outcome.
 

### 2023-04-07 products/manuscript/Manuscript.qmd
Project Part 4

@@
 Is there some serious error that I am making here?
+At least for dust and copies_mL, these are normally distributed, so this is why I am assuming Pearson's is an accepted method.
 
@@
 
-    -   Why are my RMSE and AIC so high (see table under *Conclusion* section)?
+    -   Why is my RMSE so high (see table under *Conclusion* section)?
 
@@
 I am thinking some sort of PCA that depicts which predictor variables have the greatest influence on the outcome.
+
+Thank you!
 
@@
 
+\*need to add details on other environmental variables
+
 <!--# Add detail on other env variables -->
@@
 
-Shapiro-Wilk's tests were conducted to assess distribution of the data. Any data that did not have a normal distribution was log-transformed and the test was run again.
+Shapiro-Wilk's tests were conducted to assess distribution of the data.
+Any data that did not have a normal distribution was log-transformed and the test was run again.
+Variables that continue to fail normality will have non-parametric tests run on them.
 
@@
 
-![](images/cca_bo.jpg){width="203"}
-
-![](images/cca_c2.jpg){width="203"}
-
-![](images/cca_rd.jpg){width="203"}
-
-According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1.
+CCF plots showed that BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1.
 

### 2023-04-10 code/analysis_code/stats/3_correlation.qmd
distributuion edits

@@
   filter(!is.na(date),
-         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
-  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normally distributed)
+         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
 
@@
 
-#### Make Lag data sets by Site
+### Make Lag data sets by Site
 
-This cleans out non-numeric values, creates site lag from CCA, and producea a final daily plot
+This cleans out non-numeric values, creates site lag from CCA, and produces a final daily plot
 
@@
 
+### Log10 Transform non-normally distributed data and Remove Non-Parametric Variables
+
+Distributions can be found in `1_distribution.qmd`
+
+```{r}
+#Blind Oso
+bo<-
+  bo %>%
+  mutate(amm = log10(amm), #Log-transforms data
+         din_dip = log10(din_dip),
+         doc = log10(doc),
+         dust = log10(dust)) %>%
+  select(!c(secchi, tn, tdn, don)) #Removes non-parametric data
+
+#Canals
+c2<- 
+  c2 %>%
+  mutate(orthop = log10(orthop),
+         dust = log10(dust)) %>%
+  select(!c(secchi, nn, sil, doc,toc, tn))
+
+#Gulf
+rd<- 
+  rd %>%
+  mutate(doc = log10(doc),
+         toc = log10(toc),
+         dust = log10(dust)) %>%
+  select(!c(sal, do_mgl, secchi, nn, tn))
+```
+
 # Correlation Matrix
@@
 ```{r, include = FALSE}
-b<- cor(bo)
+b<- cor(bo) #Default method = pearson
 round(b,2) #Round correlation coeff to "x" digits

### 2023-04-10 code/analysis_code/stats/4_uni_variate.qmd
distributuion edits

@@
        x = "log(dust)")
+fig1
+fig2
+fig3
 ```

### 2023-04-10 code/analysis_code/stats/3_correlation.qmd
edits

@@
 
-```{r}
+```{r, fig.width=3, fig.height=3}
 corrplot(b, type = "upper", order = "hclust", 

### 2023-04-11 code/analysis_code/stats/3_correlation.qmd
added chl to enc data

@@
          doc = log10(doc),
-         dust = log10(dust)) %>%
+         dust = log10(dust),
+         chl = log10(chl)) %>%
   select(!c(secchi, tn, tdn, don)) #Removes non-parametric data

### 2023-04-11 code/analysis_code/stats/3_correlation.qmd
swirtched corr plots to spearmans

--- 2023-04-10 code/analysis_code/stats/3_correlation.qmd
@@
 
+<!--# If we are using a nonparametric correlation (i.e. Spearman) we do not log-transform data -->
+
 Distributions can be found in `1_distribution.qmd`
@@
 #Blind Oso
-bo<-
-  bo %>%
-  mutate(amm = log10(amm), #Log-transforms data
-         din_dip = log10(din_dip),
-         doc = log10(doc),
-         dust = log10(dust)) %>%
-  select(!c(secchi, tn, tdn, don)) #Removes non-parametric data
+#bo<-
+  #bo %>%
+ # mutate(amm = log10(amm), #Log-transforms data
+        # din_dip = log10(din_dip),
+        # doc = log10(doc),
+        # dust = log10(dust),
+         #chl = log10(chl)) %>%
+ # select(!c(secchi, tn, tdn, don)) #Removes non-parametric data
 
 #Canals
-c2<- 
-  c2 %>%
-  mutate(orthop = log10(orthop),
-         dust = log10(dust)) %>%
-  select(!c(secchi, nn, sil, doc,toc, tn))
+#c2<- 
+  #c2 %>%
+  #mutate(orthop = log10(orthop),
+         #dust = log10(dust)) %>%
+  #select(!c(secchi, nn, sil, doc,toc, tn))
 
 #Gulf
-rd<- 
-  rd %>%
-  mutate(doc = log10(doc),
-         toc = log10(toc),
-         dust = log10(dust)) %>%
-  select(!c(sal, do_mgl, secchi, nn, tn))
+#rd<- 
+ # rd %>%
+ # mutate(doc = log10(doc),
+      #   toc = log10(toc),
+        # dust = log10(dust)) %>%
+  #select(!c(sal, do_mgl, secchi, nn, tn))
 ```
@@
 ```{r, include = FALSE}
-b<- cor(bo) #Default method = pearson
+b<- cor(bo, method = "spearman") #Default method = pearson
 round(b,2) #Round correlation coeff to "x" digits
 
-c<- cor(c2)
+c<- cor(c2, method = "spearman")
 round(c,2)
 
-r<- cor(rd)
+r<- cor(rd, method = "spearman")
 round(r,2)
@@
 
+<!--# This is using Spearman's with raw data -->
+
 ```{r}
-corrplot(b, type = "upper", order = "hclust", 
+corrplot(b, type = "upper", 
          tl.col = "black", tl.srt = 45,
@@
 
-corrplot(c, type = "upper", order = "hclust", 
+corrplot(c, type = "upper", 
          tl.col = "black", tl.srt = 45,
@@
 
-corrplot(r, type = "upper", order = "hclust", 
-         tl.col = "black", tl.srt = 45,
-         title = "Gulf", mar=c(0,0,1,0))
+corrplot(r, type = "upper", 
+         tl.col = "black", tl.srt = 45, mar=c(0,0,1,0))
+         title(main = "Gulf" )
 ```
@@
 #corrplot(bclean$r, type="upper", order="hclust", 
-        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
-# Insignificant correlations are leaved blank
-#corrplot(bclean$r, type="upper", order="hclust", 
-        # p.mat = bclean$P, sig.level = 0.01, insig = "blank")
+        # p.mat = bclean$P, sig.level = 0.05, insig = "blank")
+
+#corrplot(cclean$r, type="upper", order="hclust", 
+        # p.mat = cclean$P, sig.level = 0.05, insig = "blank")
+
+#corrplot(rclean$r, type="upper", order="hclust", 
+         #p.mat = rclean$P, sig.level = 0.05, insig = "blank")
+
 ```

### 2023-04-11 code/analysis_code/stats/4_uni_variate.qmd
swirtched corr plots to spearmans

--- 2023-04-07 code/analysis_code/stats/4_uni_variate.qmd
@@
   filter(!is.na(date),
-         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) %>%
-  mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)
+         date %in% as.Date("2022-07-04"):as.Date("2022-07-19")) #%>%
+  #mutate(t7sum = log10(t7sum)) #Log transform Dust conc (as it is not normal)
 
@@
   filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.
-master<- master0[c(1,8,14,32)] #Select columns of interest
+master<- master0[c(1,8,14,34)] #Select columns of interest
 
@@
 ```{r}
-fig1<- bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
+bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_correlation(method = "spearman")+
   labs(title = "LM DustxCopies (BO+1)",
        x = "log(dust)")
-
-fig2<- c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
+ c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_correlation(method = "spearman")+
   labs(title = "LM DustxCopies (C2+2)",
        x = "log(dust)")
-
-fig3<- rd2 %>% ggplot(aes(lag, copies_mL)) +
+ rd2 %>% ggplot(aes(lag, copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_correlation(method = "spearman")+
   labs(title = "LM DustxCopies (RD+1)",
        x = "log(dust)")
+
 ```

### 2023-04-17 code/analysis_code/env/env_par.qmd
edits

--- 2023-04-04 code/analysis_code/env/env_par.qmd
@@
     axis.text.x = element_blank(),
-    axis.title.y = element_text(size = 15),
-    legend.position = c(0.1,0.2),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
     legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
@@
     plot.title= element_text(hjust = 0.5, size = 20),
-    axis.title.x = element_text(size = 15),
-    axis.title.y = element_text(size = 15),
-    axis.text.x = element_text(size = 10),
-    axis.text.y = element_text(size = 10),
-    legend.position = "none")+
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
     scale_color_manual(values = c("Blind Oso" = "#809C52",
@@
                                 "Gulf" = "4c848c"))
-
-sal
 ```
 
-#### Combine Temperature and Salinity Plots
+#### Dissolved Oxygen
 
-```{r, fig.width= 4, fig.height=4}
-fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")
+```{r}
+do<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = do_mgl,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DO (mg/L)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Nitrite:Nitrate
+
+```{r}
+nn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = nn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Nitrate-Nitrite (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = c(0.3,0.7),
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
+    legend.text = element_text(size = 7)) + #Moves legend onto plot  + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Orthophosphate
+
+```{r}
+ortho<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = orthop,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Orthophosphate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DON
+
+```{r}
+don<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = don,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DON (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+   axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DOC
+
+```{r}
+doc<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = doc,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DOC (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### pH
+
+```{r}
+pH<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = pH,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "pH",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Ammonium
+
+```{r}
+amm<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = amm,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Ammonium (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Total Nitrogen
+
+```{r}
+tn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = tn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Total Nitrogen (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Silicate
+
+```{r}
+sil<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = sil,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Silicate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Copies/mL
+
+```{r}
+cop<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = copies_mL,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Vibrio (copies per mL)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = c(0.2,0.7),
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
+    legend.text = element_text(size = 7)) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### ![](images/paste-21541164.png)Combine Environment Plots
+
+```{r}
+fig<- ggarrange(temp,
+                sal,
+                #do,
+                nn,
+                ortho,
+                don,
+                doc,
+                #pH,
+                #amm,
+               # tn,
+               # sil,
+                #cop,
+                ncol = 3, nrow = 2,align = "v")
 

### 2023-04-17 code/analysis_code/stats/4_uni_variate.qmd
edits

--- 2023-04-07 code/analysis_code/stats/4_uni_variate.qmd
@@
 ```{r}
-fig1<- bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
+fig1<-
+  bo2 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4)+
   labs(title = "LM DustxCopies (BO+1)",
        x = "log(dust)")
-
-fig2<- c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
+ fig2<-
+   c22 %>% ggplot(aes(x = lag, y = copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4)+
   labs(title = "LM DustxCopies (C2+2)",
        x = "log(dust)")
-
-fig3<- rd2 %>% ggplot(aes(lag, copies_mL)) +
+fig3<-
+  rd2 %>% ggplot(aes(lag, copies_mL)) +
   geom_point() +
   stat_smooth(method = "lm")+
-  stat_correlation(method = "pearson")+
+  stat_poly_eq(aes(x = lag, y = copies_mL),
+               rr.digits = 4)+
   labs(title = "LM DustxCopies (RD+1)",
        x = "log(dust)")
+
 ```

### 2023-04-17 code/analysis_code/stats/5_multi_variate_models/bo_multi.qmd
edits

---
title: "multi-variate"
format:
  pdf:
    theme: default
---

# Getting Started

#### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(lubridate)#Dates
library(kableExtra) #Tables
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

# Cleaning

#### Clean and Trim

```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

#### Make into one data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(site, sal, temp, copies_mL,t7sum, date, orthop,nn, don,doc) %>%
  mutate(t7sum = log10(t7sum))
```

#### Site-specific Data sets and Clean df

```{r}
b<- master %>%
  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
b<- b %>%
mutate(
  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")
```

# Models

## 0. Null Model:

### Create 5 Fold

```{r}
fold_bo <- vfold_cv(b, v = 5, repeats = 5) #Data was too small to stratify (<20)
```

### Make Models

```{r, include= FALSE}
#Create Null Recipe for Growth at all three sites
null_bo<- recipe(copies_mL ~ 1, data = b)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model for three sites
null_wflow_bo<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_bo)

#Use workflow to fit model to  data set
null_fit_bo<-
  fit_resamples(null_wflow_bo, resamples = fold_bo)

```

### Get Metrics for Null

#### RMSE

```{r}
null_met_bo <- 
  collect_metrics(null_fit_bo)
null_met_bo
rm0<- 20985.12
rs0<- "NA"
```

#### AIC

```{r}
bo_null_mod<- 
  lm(copies_mL ~ 1, data = b)
aic0<-AIC(bo_null_mod) %>% round(digits = 2)
```

## 1. Dust Model

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe1<- recipe(copies_mL ~ log10dust, data = b)

#Workflow that adds recipe to model
Growth_wflow1<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe1)

#Use workflow to fit model to  data set
growth_fit1<- Growth_wflow1 %>%
  fit(data = b)

#View as Tibble 
growth_fit1 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test1 <- augment(growth_fit1, b)
rmse1 <- aug_test1 %>% rmse(truth = copies_mL, .pred)
rsq1 <- aug_test1 %>% rsq(truth = copies_mL, .pred)

mod1<- lm(copies_mL ~ log10dust, b)
aic1<-AIC(mod1) %>% round(digits = 2)

metrics1<- full_join(rmse1, rsq1)
metrics1
```

## 2. Water Chemistry Model

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe2<- recipe(copies_mL ~ sal+temp, data = b)

#Workflow that adds recipe to model
Growth_wflow2<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe2)

#Use workflow to fit model to  data set
growth_fit2<- Growth_wflow2 %>%
  fit(data = b)

#View as Tibble 
growth_fit2 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test2 <- augment(growth_fit2, b)
rmse2 <- aug_test2 %>% rmse(truth = copies_mL, .pred)
rsq2 <- aug_test2 %>% rsq(truth = copies_mL, .pred)

mod2<- lm(copies_mL ~ sal+temp, b)
aic2<- AIC(mod2)%>% round(digits = 2)

metrics2<- full_join(rmse2, rsq2)
metrics2
```

## 3. Chemistry + Dust

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe3<- recipe(copies_mL ~ sal+temp + log10dust, data = b)

#Workflow that adds recipe to model
Growth_wflow3<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe3)

#Use workflow to fit model to  data set
growth_fit3<- Growth_wflow3 %>%
  fit(data = b)

#View as Tibble 
growth_fit3 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test3 <- augment(growth_fit3, b)
rmse3 <- aug_test3 %>% rmse(truth = copies_mL, .pred)
rsq3 <- aug_test3 %>% rsq(truth = copies_mL, .pred)

mod3<- lm(copies_mL ~ sal+temp + log10dust, b)
aic3<- AIC(mod3)%>% round(digits = 2)

metrics3<- full_join(rmse3, rsq3)
metrics3
```

## 4. Model 3 + Nutrients

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe4<- recipe(copies_mL ~ sal+temp+orthop+log10dust+nn, data = b)

#Workflow that adds recipe to model
Growth_wflow4<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe4)

#Use workflow to fit model to  data set
growth_fit4<- Growth_wflow4 %>%
  fit(data = b)

#View as Tibble 
growth_fit4 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test4 <- augment(growth_fit4, b)
rmse4 <- aug_test4 %>% rmse(truth = copies_mL, .pred)
rsq4 <- aug_test4 %>% rsq(truth = copies_mL, .pred)

mod4<- lm(copies_mL ~ sal+temp+orthop+log10dust+nn, b)
aic4<- AIC(mod4)%>% round(digits = 2)

metrics4<- full_join(rmse4, rsq4)
metrics4
```

## 5. Module 4 + DOM

#### Model

```{r}
#Create Recipe for Growth 
growth_recipe5<- recipe(copies_mL ~ sal+temp+log10dust+orthop+nn+don+doc, data = b)

#Workflow that adds recipe to model
Growth_wflow5<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(growth_recipe5)

#Use workflow to fit model to  data set
growth_fit5<- Growth_wflow5 %>%
  fit(data = b)

#View as Tibble 
growth_fit5 %>%
  extract_fit_parsnip() %>%
  tidy()
```

#### Metrics

```{r}
aug_test5 <- augment(growth_fit5, b)
rmse5 <- aug_test5 %>% rmse(truth = copies_mL, .pred)
rsq5 <- aug_test5 %>% rsq(truth = copies_mL, .pred)

mod5<- lm(copies_mL ~ sal+temp+log10dust+orthop+nn+don+doc, b)
aic5<-AIC(mod5)%>% round(digits = 2)

metrics5<- full_join(rmse5, rsq5)
metrics5
```

# Tables

#### Label Metrics

```{r}
#R2
rs1<- rsq1$.estimate %>% round(digits = 2)
rs2<- rsq2$.estimate %>% round(digits = 2)
rs3<- rsq3$.estimate %>% round(digits = 2)
rs4<- rsq4$.estimate %>% round(digits = 2)
rs5<- rsq5$.estimate %>% round(digits = 2)

#RMSE
rm1<- rmse1$.estimate 
rm2<- rmse2$.estimate
rm3<- rmse3$.estimate
rm4<- rmse4$.estimate
rm5<- rmse5$.estimate
```

#### Make Table

```{r}
kable(data.frame(Model = c("Null Model", "Model 1: Dust", "Model 2: Water Chemistry", "Model 3: Chemistry + Dust", "Model 4: Model 3 + Nutrients", "Model 5: Model 4 + DOM"),
             R2 = c(rs0,rs1,rs2,rs3,rs4,rs5),
             RMSE = c(rm0,rm1,rm2,rm3,rm4,rm5),
             AIC = c(aic0, aic1,aic2,aic3,aic4,aic5)),
      caption = "Blind Oso: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.") 
```

####

### 2023-04-17 code/analysis_code/stats/5_multi_variate_models/c2_multi.qmd
edits

--- 2023-04-17 code/analysis_code/stats/5_multi_variate_models/bo_multi.qmd
@@
 b<- master %>%
-  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
+  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
 b<- b %>%
 mutate(
-  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
+  lag =Lag(b$t7sum, shift  = 2)) %>% #Create 2 day lag
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
@@
 #Create Null Recipe for Growth at all three sites
-null_bo<- recipe(copies_mL ~ 1, data = b)
+null<- recipe(copies_mL ~ 1, data = b)
 
@@
 #Workflow that adds recipe to model for three sites
-null_wflow_bo<- 
+null_wflow<- 
   workflow() %>%
   add_model(lm_mod) %>%
-  add_recipe(null_bo)
+  add_recipe(null)
 
 #Use workflow to fit model to  data set
-null_fit_bo<-
-  fit_resamples(null_wflow_bo, resamples = fold_bo)
+null_fit<-
+  fit_resamples(null_wflow, resamples = fold_bo)
 
@@
 ```{r}
-null_met_bo <- 
-  collect_metrics(null_fit_bo)
-null_met_bo
-rm0<- 20985.12
+null_met <- 
+  collect_metrics(null_fit)
+null_met
+rm0<- 8082.466
 rs0<- "NA"
@@
              AIC = c(aic0, aic1,aic2,aic3,aic4,aic5)),
-      caption = "Blind Oso: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.") 
+      caption = "Canals: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.") 
 ```

### 2023-04-17 code/analysis_code/stats/5_multi_variate_models/rd_multi.qmd
edits

--- 2023-04-17 code/analysis_code/stats/5_multi_variate_models/c2_multi.qmd
@@
 b<- master %>%
-  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
+  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
 b<- b %>%
 mutate(
-  lag =Lag(b$t7sum, shift  = 2)) %>% #Create 2 day lag
+  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
 filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
@@
 ```{r}
-fold_bo <- vfold_cv(b, v = 5, repeats = 5) #Data was too small to stratify (<20)
+fold <- vfold_cv(b, v = 5, repeats = 5) #Data was too small to stratify (<20)
 ```
@@
 null_fit<-
-  fit_resamples(null_wflow, resamples = fold_bo)
+  fit_resamples(null_wflow, resamples = fold)
 
@@
 null_met
-rm0<- 8082.466
+rm0<- 50771.79
 rs0<- "NA"
@@
 ```{r}
-bo_null_mod<- 
+null_mod<- 
   lm(copies_mL ~ 1, data = b)
-aic0<-AIC(bo_null_mod) %>% round(digits = 2)
+aic0<-AIC(null_mod) %>% round(digits = 2)
 ```
@@
              AIC = c(aic0, aic1,aic2,aic3,aic4,aic5)),
-      caption = "Canals: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.") 
+      caption = "Gulf: Comparative linear regression analysis of Vibrio copies per mL between 6 operational models.")
 ```

### 2023-04-17 products/manuscript/Manuscript.qmd
edits

@@
     toc: false
-    number-sections: true
+    number-sections: false
     highlight-style: github
@@
 
-# Questions for Dr. Handel:
+# Edits
 
 1.
-In my `code/analysis_code/stats/3_uni_variate.qmd` code I was playing around with R2 for my linear models.
-Previously I have been plotting R2 onto the graphs using `stat_poly_equation()` and have also checked through `tidymodeling`.
-But then I came across the `stat_correlation()` command.
-When specifying "pearson" as the method, I get a **MUCH** higher R2 value for the models (\~0.48 to \~0.70).
-I have been trying to figure out what `stat_poly_equation` uses as its method and why there is such a difference.
-Additionally, my actual model produces the same lower R2 values (using `rsq()`) as the `stat_poly_equation()`.
-Is there some serious error that I am making here?
-At least for dust and copies_mL, these are normally distributed, so this is why I am assuming Pearson's is an accepted method.
-
-2.  Does my overall workflow in `code/analysis_code/stats/3_uni_variate.qmd` make sense? Is there anything that I am missing, or be a better option for analysis?
-
-    -   Why is my RMSE so high (see table under *Conclusion* section)?
-
-    -   If this looks good, I plan to do the same workflow for all of the environmental parameters (that seem to have a relationship) as well.
-For now, I just have plots to see what may be worth diving into.
-
-3.  Looking at `code/analysis_code/stats/4_multi_variate.qmd` , I am a little stuck on how to best create a multivariate model.
-
-    -   Example: We can see this Salinity at Blind Oso drop at the same time that dust is coming in. This will most likely have an impact of Vibrio growth, but Salinity is probably non-linear.
-
-    -   Taking a look at the section titled Blind Oso: Growth \~Salinity + Dust, does my overall workflow look correct?
-
-4.
-Ultimately, I feel as though I am missing some "oomph" when it comes to my statistical analysis.
-The linear models seem adequate, but I know that there are multiple variables at play at these sites (and we have data to explore some of this).
-I am open to any suggestions that may diversify and expand my analyses!
-I am thinking some sort of PCA that depicts which predictor variables have the greatest influence on the outcome.
-
-Thank you!
+I would do a lot of visual inspections to see how the different models fit and what they predict.
+You can try a linear model and a more complex ML one and compare to see if maybe something beyond a linear relationship improves results.
+2.
+Given that you don't have a lot of data, I'm not sure a PCA makes sense.
+I think in-depth analysis of the relation between the variables and the outcome and then in addition to the linear models trying a few more complex ones will probably be the better approach.
+3.  The figures in the manuscript look good, but some are too small to read. Needs breaking up and/or increasing font size
 
@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year \[@dalmeidaModelSaharanDust1986\].
+## Saharan Dust
+
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico.
@@
 These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] .
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a\].
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
+
+## Microbial Blooms
 
@@
 The high amount of nitrogen produced via fixation can be adequate enough to stimulate toxic red tides, caused by the dinoflagellate *Karenia brevis* [@lenesSaharanDustPhosphatic2008].
+
+### Vibrio
 
@@
 alginolyticus*, *V.
-vulnificus*) \[@baker-austinVibrioSppInfections2018\].
+vulnificus*) [@baker-austinVibrioSppInfections2018\].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 parahaemolyticus*) and sea breams (*V.
-harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010\] among others.
+harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
+
+### Previous Work
 
@@
 Given that human exposure is likely in these coastal waters, it is important to elucidate the relationship between dust input and microbial blooms in these higher nutrient environments.
+
+## Present Research
 
@@
 
-#### Cleaning of Total Vibrio Enumeration dataset
+#### Cleaning of Total Vibrio Enumeration data set
 
@@
 
+<!--# Have paragraph describing and then have figures. -->
+
 ## Exploratory/Descriptive analysis
@@
 
-![](images/paste-C33B6E43.png)
+![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png)
 
@@
 
-![](images/000031.png)
+![Figure 2: Environmental parameters across daily time series](images/00000f%20(1).png)
 
@@
 
-## Basic statistical analysis
+## Statistical analysis
 
@@
 
-[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022.
-Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
-Therefore, these CCA were made with only this time series.
-This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
-But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days.
-Below we will run linear models that DO contain these prior dust days.
-And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
-But this ultiamtely means that the lag identified above.
-may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
-
 ### Linear Regression Models
 
-![](images/Screen%20Shot%202023-04-06%20at%2010.23.28%20PM.png)
+![](images/Screen%20Shot%202023-04-17%20at%2011.21.32%20AM.png)
 
@@
 As mentioned above, the lags identified may not correspond to the most significant relationship between the two variables.
+
+<!--# Update table -->
 
@@
 
--   I am also planning on running a multi-variate model to look at multiple predictors of interests (temp, salinity, nutrients). This code can be found in my `dust_copies_stats.qmd` code.
+### Multi-variate Linear Regressions
 
-### Residuals:
+![](images/Screen%20Shot%202023-04-17%20at%203.16.17%20PM.png)
 
-Instead of (or in addition to) plotting observed versus predicted for continuous outcomes, you can plot the difference between the two.
-These differences are called the residuals.
-What you are looking for is a cloud of points with no discernible pattern.
-If there is a pattern (e.g., an overall skew, or more points above the 0 y-axes than below), it again suggests that there is still some pattern/signal in the data that the model didn't capture.
+![](images/Screen%20Shot%202023-04-17%20at%203.17.06%20PM.png)
 
-![](images/000005-02.png){width="299"}
+![](images/Screen%20Shot%202023-04-17%20at%203.17.43%20PM.png)
 
-![](images/000009.png){width="300"}
-
-![](images/00000d.png){width="296"}
+<!--# Fix Table 1 titles -->
 
@@
 
+[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022.
+Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
+Therefore, these CCA were made with only this time series.
+This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
+But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days.
+Below we will run linear models that DO contain these prior dust days.
+And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
+But this ultimately means that the lag identified above.
+may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
+
 ## Conclusions

### 2023-04-18 code/analysis_code/env/env_par.qmd
edits

--- 2023-04-04 code/analysis_code/env/env_par.qmd
@@
     axis.title.y = element_text(size = 15),
-    legend.position = c(0.1,0.2),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
     legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
@@
     plot.title= element_text(hjust = 0.5, size = 20),
-    axis.title.x = element_text(size = 15),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
     axis.title.y = element_text(size = 15),
-    axis.text.x = element_text(size = 10),
-    axis.text.y = element_text(size = 10),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Dissolved Oxygen
+
+```{r}
+do<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = do_mgl,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DO (mg/L)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Nitrite:Nitrate
+
+```{r}
+nn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = nn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Nitrate-Nitrite (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+   # legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+   # legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
+   # legend.text = element_text(size = 7)) + #Moves legend onto plot  + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Orthophosphate
+
+```{r}
+ortho<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = orthop,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Orthophosphate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DON (TDN - NN - Amm)
+
+```{r}
+don<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = don,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DON (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+   axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DOC
+
+```{r}
+doc<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = doc,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DOC (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    axis.title.x = element_blank(),
+   # axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### pH
+
+```{r}
+pH<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = pH,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "pH",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+  #  axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Ammonium
+
+```{r}
+amm<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = amm,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Ammonium (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+  #  axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Total Nitrogen
+
+```{r}
+tn<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = tn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Total Nitrogen (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Silicate
+
+```{r}
+sil<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = sil,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Silicate (µM)",
+  #  title = "Temperature across Daily Time Series by Site",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.position = "none",
+    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### TDN
+
+```{r}
+tdn<-
+   daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = tdn,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "TDN (µM)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+   # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
     legend.position = "none")+
@@
                                 "Gulf" = "4c848c"))
-
-sal
 ```
 
-#### Combine Temperature and Salinity Plots
+#### DIN:DIP ((Amm + NN / Ortho))
 
-```{r, fig.width= 4, fig.height=4}
-fig<- ggarrange(temp,sal + font("x.text", size = 10), ncol = 1, nrow = 2,align = "v")
+```{r}
+din_dip<- 
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = din_dip,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DIN:DIP (µM)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    #axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### DIN (Amm-NN)
+
+```{r}
+din<- 
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = din,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "DIN (µM)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+   # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Secchi
+
+```{r}
+secchi<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = secchi,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Secchi Depth (m)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+   # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Chlorophyll
+
+```{r}
+chl<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = chl,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Chlorophyl (µg/L)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+   # axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+    legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### Copies/mL
+
+```{r}
+cop<-
+  daily %>%
+  ggplot() + geom_line(
+  aes(
+    x = date,
+    y = copies_mL,
+    color = site),
+  size = 1)+
+  theme_bw()+
+  labs(
+    x = "Date",
+    y = "Vibrio (copies per mL)",
+    color = "Site")+
+  theme(
+    plot.title= element_text(hjust = 0.5, size = 20),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+  #  axis.text.x = element_blank(),
+    axis.title.y = element_text(size = 15),
+    axis.text.y = element_text(size = 7),
+   # legend.key.size = unit(0.3,'cm'),
+    legend.position = "none")+
+  #  legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'),
+  #  legend.text = element_text(size = 7)) + #Moves legend onto plot 
+    scale_color_manual(values = c("Blind Oso" = "#809C52",
+                                "Canals" = "#ba4716",
+                                "Gulf" = "4c848c"))
+```
+
+#### ALL
+
+```{r, fig.width= 8, fig.height=4}
+fig<- ggarrange(#temp,
+                #sal,
+               # do,
+                #nn,
+                #ortho,
+                #don,
+                #doc,
+                #pH,
+                #amm,
+                tn,
+                sil,
+                cop,
+                chl,
+                secchi,
+                din,
+                din_dip,
+                tdn,
+                ncol = 3, 
+                nrow = 3,
+                align = "v")
 

### 2023-04-18 products/manuscript/Manuscript.qmd
edits

@@
 
+### Correlation Matirices
+
+<!--# Add these in -->
+
 ### Linear Regression Models
 
-![](images/Screen%20Shot%202023-04-17%20at%2011.21.32%20AM.png)
+![Figure 3: Linear regression models examining the relationship between dust and vibrio growth](images/Screen%20Shot%202023-04-17%20at%2011.21.32%20AM.png)
 
@@
 
+This study aimed to examine microbial response to Saharan dust deposition in high nutrient coastal waters.
+Specifically focusing on bacteria belong to the genus *Vibrio*, we found that growth response following dust input varied by site.
+This could be due to differing site dynamics such as salinity, temperature, and nutrient levels.
+
 ## Summary and Interpretation
+
+### Methods
+
+For this analysis, we examined *Vibrio* growth dynamics in response to Saharan Dust depsoition in three sites in Corpus Christi, TX that had differing baseline nutrient levels.
+The Gulf, with relatively constant salinities and low nutrient levels, Blind Oso Bay with high nutrients and fluctuating salinities, and a residential canal system with consistent temperature and salinities and nutrient levels.
+Samples were collected before, during, and after a dust event and *Vibrio* were enumerated using qPCR.
+Data on nutrient levels and other environmental parameters (e.g.
+salinity and temperature) were also collected during this period.
+The three data sets (Vibrio Enumeration, Environmental parameters, and Dust Concentration) were then processed and cleaned for analysis.
+
+### Findings/Explanations
+
+#### Dust
+
+The daily time series captured two dust events.
+One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57.
+Dust AOD returned to lower concentrations following the time series.
+
+<!--# What are AOD units? -->
+
+#### Environmental Parameters
+
+Environmental parameters were site-specific.
+Temperature and salinities were consistent and high at the Canals ( <!--# Add ranges -->).
+While salinity remained relatively constant at the Gulf , temperature was observed to drop starting on the 12th of July, possibly due to a coastal upwelling event.
+Blind Oso exhibited fluctuations in both temperature and salinity within the daily time series with a drop in salinity from 42 to 20 within an eight day period.
+
+Nutrients (orthophosphate and nitrate+nitrite) were observed to increase following dust deposition at both Blind Oso and the Canals site, with the former having more dramatic peaks.
+Nutrient levels at the Gulf remained relatively constant throughout the time series.
+DON and DOC had distinct peaks at Blind Oso on days following dust deposition, however the other two sites were relatively stable.
+
+#### Vibrio Enumeration
+
+Peaks in *Vibrio* growth were observed at the Gulf 1-2 days following dust deposition.
+Blind Oso exhibited a similar response, but was slightly dampened in comparison.
+This could be due to the site-specific nutrient dynamics.
+While the Gulf exhibited a sharper increase during the second, heavy dust event, Blind Oso exhibited a smaller increase as compared to the light dust event. \
+\
+If the Gulf has presumably lower nutrient levels, the constituents from the influx of dust may be used up rapidly, so we still see a sharp secondary growth response at the Gulf during the heavy dust event.
+Blind Oso Bay has presumably higher nutrient levels, so while we see an initial spike in growth, the secondary spike is not as dramatic.
+So in this case, it may be that the dust has reached a maximum to how much it can support growth, possibly due to the dust constituents not being as essential for growth at this inland site.
+
+#### Statistical Analysis
 
@@
 
+<!--# Sample size  -->
+
 ## Conclusions

### 2023-04-18 code/analysis_code/stats/3_correlation.qmd
edits
Made spearman and pearson Cor plots

@@
 **This code creates a correlation matrix to look at relationships between variables. This helps identify specific relationships of interest**
-
-[**THIS CODE IS STILL IN THE WORKS**]{.ul}
 
@@
 
-### Log10 Transform non-normally distributed data and Remove Non-Parametric Variables
+#### Make Datasets for normal and non-normal variables
 
-<!--# If we are using a nonparametric correlation (i.e. Spearman) we do not log-transform data -->
+Variables that are normally distributed across all three sites will be in one df. All others will be in another.
+
+```{r}
+b_norm<- 
+  bo %>%
+  select(!c(sal, secchi, nn, tdn, don, doc, toc, tn))
+b_para<- 
+  bo %>%
+  select(sal, secchi, nn, tdn, don, doc, toc, tn)
+
+c_norm<- 
+  c2 %>%
+  select(!c(sal, secchi, nn, tdn, don, doc, toc, tn))
+c_para<- 
+  c2 %>%
+  select(sal, secchi, nn, tdn, don, doc, toc, tn)
+
+r_norm<- 
+  rd %>%
+  select(!c(sal, secchi, nn, tdn, don, doc, toc, tn))
+r_para<- 
+  rd %>%
+  select(sal, secchi, nn, tdn, don, doc, toc, tn)
+```
+
+### Log10 Transform non-normally distributed data for Norm df
 
@@
 #Blind Oso
-#bo<-
-  #bo %>%
- # mutate(amm = log10(amm), #Log-transforms data
-        # din_dip = log10(din_dip),
-        # doc = log10(doc),
-        # dust = log10(dust),
-         #chl = log10(chl)) %>%
- # select(!c(secchi, tn, tdn, don)) #Removes non-parametric data
+b_norm<-
+  b_norm %>%
+  mutate(amm = log10(amm), #Log-transforms data
+         din_dip = log10(din_dip),
+         dust = log10(dust),
+         chl = log10(chl))
 
 #Canals
-#c2<- 
-  #c2 %>%
-  #mutate(orthop = log10(orthop),
-         #dust = log10(dust)) %>%
-  #select(!c(secchi, nn, sil, doc,toc, tn))
+c_norm<- 
+  c_norm %>%
+  mutate(orthop = log10(orthop),
+         dust = log10(dust))
 
 #Gulf
-#rd<- 
- # rd %>%
- # mutate(doc = log10(doc),
-      #   toc = log10(toc),
-        # dust = log10(dust)) %>%
-  #select(!c(sal, do_mgl, secchi, nn, tn))
+r_norm<- 
+  r_norm %>%
+  mutate(dust = log10(dust))
 ```
@@
 ```{r, include = FALSE}
-b<- cor(bo, method = "spearman") #Default method = pearson
+#Normal Data = Pearson's
+b<- cor(b_norm, method = "pearson") #Default method = pearson
 round(b,2) #Round correlation coeff to "x" digits
 
-c<- cor(c2, method = "spearman")
+c<- cor(c_norm, method = "pearson")
 round(c,2)
 
-r<- cor(rd, method = "spearman")
+r<- cor(r_norm, method = "pearson")
+round(r,2)
+
+#Non-parametric Data  = Spearman's
+b1<- cor(b_para, method = "spearman") 
+round(b,2) #Round correlation coeff to "x" digits
+
+c1<- cor(c_para, method = "spearman")
+round(c,2)
+
+r1<- cor(r_para, method = "spearman")
 round(r,2)
@@
 ```{r}
-bclean<- rcorr(as.matrix(bo))
+bclean<- rcorr(as.matrix(b_norm))
 flattenCorrMatrix(bclean$r, bclean$P)
 
-cclean<- rcorr(as.matrix(c2))
+cclean<- rcorr(as.matrix(c_norm))
 flattenCorrMatrix(cclean$r, cclean$P)
 
-rclean<- rcorr(as.matrix(rd))
+rclean<- rcorr(as.matrix(r_norm))
 flattenCorrMatrix(rclean$r, rclean$P)
@@
 
-<!--# This is using Spearman's with raw data -->
+##### Normal
 
@@
          tl.col = "black", tl.srt = 45,
-         title = "Blind Oso",  mar=c(0,0,1,0)) #Adds and lowers title
+         title = "Blind Oso_Pearson",  mar=c(0,0,1,0)) #Adds and lowers title
 
@@
          tl.col = "black", tl.srt = 45,
-         title = "Canals", mar=c(0,0,1,0))
+         title = "Canals_Pearson", mar=c(0,0,1,0))
 
@@
          tl.col = "black", tl.srt = 45, mar=c(0,0,1,0))
-         title(main = "Gulf" )
+         title(main = "Gulf_Pearson" )
+```
+
+##### Non-Para
+
+```{r}
+corrplot(b1, type = "upper", 
+         tl.col = "black", tl.srt = 45,
+         title = "Blind Oso_Spearman",  mar=c(0,0,1,0)) #Adds and lowers title
+
+corrplot(c1, type = "upper", 
+         tl.col = "black", tl.srt = 45,
+         title = "Canals_Spearman", mar=c(0,0,1,0))
+
+corrplot(r1, type = "upper", 
+         tl.col = "black", tl.srt = 45, mar=c(0,0,1,0))
+         title(main = "Gulf_Spearman" )
 ```

### 2023-04-18 products/manuscript/Manuscript.qmd
Made spearman and pearson Cor plots

@@
 You can try a linear model and a more complex ML one and compare to see if maybe something beyond a linear relationship improves results.
+
 2.
@@
 I think in-depth analysis of the relation between the variables and the outcome and then in addition to the linear models trying a few more complex ones will probably be the better approach.
+
+    <!--# Try a model that is copies_mL ~. -->
+
 3.  The figures in the manuscript look good, but some are too small to read. Needs breaking up and/or increasing font size
@@
 Dust concentration (AOT) was collected at time points 0 hour, 6 hour, 12 hour, and 18 hour.
-In this script, I made different data sets based on time points as well as a data set containing the summed dust concentration (of all time points) per day.
-For downstream statistical analysis, it will be helpful to have these different data sets to compare to time of sample collection.
+In this script, I converted timpoints from UTC to CST/CDT and created different dust permutations (individual time points, sums by day, and average by day).
+Because samples were collected between 0700 and 1100, a 24h composite permutation was made.
+This permutation is what was used for all downstream analyses.
 
@@
 
+This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series.
+The daily time series captured two dust events.
+One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57.
+Dust AOD returned to lower concentrations following the time series.
+The top portion of the figure depicts total enumerated Vibrio as copies per mL, with color by site.
+There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
+An initial shift in copies per mL at the Gulf and Blind Oso occur 24-48hr following initial dust introduction on the 13th, and experience another shift around 24-48hr of higher dust input on the 16th
+
 ![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png)
-
-The top portion of the figure depicts total enumerated Vibrio as copies per mL, with color by site.
-There is a noticeable shift in copies in the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
-The bottom portion of the figure depicts the dust concentration (AOD) across the daily time series.
-This figure depicts the summed dust concentration 24hr prior to sample collection (time points 13hr--> 7hr the next day) across the daily time series.
-During the time series, two small periods of higher dust AOD occurred on the 9th and 13th respectively.
-A much larger spike in AOD was observed on the 16th.
-Initial shift in copies per mL at the Gulf and Blind Oso occur 24-48hr following initial dust introduction on the 13th, and experience another shift around 24-48hr of higher dust input on the 16th.
 
@@
 
-### Correlation Matirices
+### Correlation Matrices
 
@@
 Specifically focusing on bacteria belong to the genus *Vibrio*, we found that growth response following dust input varied by site.
-This could be due to differing site dynamics such as salinity, temperature, and nutrient levels.
 
@@
 
-For this analysis, we examined *Vibrio* growth dynamics in response to Saharan Dust depsoition in three sites in Corpus Christi, TX that had differing baseline nutrient levels.
+For this analysis, we examined *Vibrio* growth dynamics in response to Saharan Dust deposition in three sites in Corpus Christi, TX that had differing baseline nutrient levels.
 The Gulf, with relatively constant salinities and low nutrient levels, Blind Oso Bay with high nutrients and fluctuating salinities, and a residential canal system with consistent temperature and salinities and nutrient levels.
@@
 ### Findings/Explanations
-
-#### Dust
-
-The daily time series captured two dust events.
-One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57.
-Dust AOD returned to lower concentrations following the time series.
 
@@
 
-Peaks in *Vibrio* growth were observed at the Gulf 1-2 days following dust deposition.
-Blind Oso exhibited a similar response, but was slightly dampened in comparison.
-This could be due to the site-specific nutrient dynamics.
-While the Gulf exhibited a sharper increase during the second, heavy dust event, Blind Oso exhibited a smaller increase as compared to the light dust event. \
-\
-If the Gulf has presumably lower nutrient levels, the constituents from the influx of dust may be used up rapidly, so we still see a sharp secondary growth response at the Gulf during the heavy dust event.
-Blind Oso Bay has presumably higher nutrient levels, so while we see an initial spike in growth, the secondary spike is not as dramatic.
-So in this case, it may be that the dust has reached a maximum to how much it can support growth, possibly due to the dust constituents not being as essential for growth at this inland site.
+Overall, *Vibrio* growth was observed following dust events at both the Gulf and Blind Oso.
+Peaks in growth were observed 24-48 hours following dust deposition.
+Linear regression analysis confirmed a correlation between growth and dust input (R2 = 0.49, 0.56, 0.44 for Blind Oso, Canals, and Gulf respectively).
+However, the environmental dynamics of each site most likely contributed the degree of response per site.
+Multivariate models were tested using specific environmental parameters: dust, DOM (DON, DOC), temperature, salinity, orthophosphate, and nitrate+nitrite.
+Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (R2 = 0.73, 0.7, and 0.76 for Blind Oso, Canals, and Gulf respectively), suggesting that while dust may contribute to a growth response in *Vibrio* following dust input, the site specific dynamics aid in dictating the degree of response observed.
 
-#### Statistical Analysis
+The Canals may have a heavily dampened growth response due to a high and constant salinity of 40-42.
+In general, optimal growth occurs at salinities of \~25-35, and less tolerant species will drop as salinities climb <!--# CITE --> .
+
+Additionally, the Gulf exhibited a sharper increase during the second, heavier dust event, while Blind Oso exhibited a smaller increase during the heavy dust event.
+This difference in response may be attributed to the nutrient dynamics of the sites.
+In marine waters, nutrients are typically limiting due to tight microbial recycling <!--# CITE -->.
+Therefore the introduction on nutrients at such a site may result in rapid uptake and utilization.
+At Blind Oso, these dust constituents may not be as essential for growth, and the dust may have reached a maximum in how much it can support growth.
 
 ## Strengths and Limitations
+
+At Blind Oso for example, the salinity is observed to drop from 42-20 within an eight day period that coincided with dust input.
+It is difficult to parse these two variables out as either or both could have contributed to growth at this site.
 

### 2023-04-19 code/analysis_code/stats/5_multi_variate_models/rf_models.qmd
+ RF model

---
title: "lasso"
format:
  html:
    theme: default
---

# Getting Started

#### Libraries

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(lubridate)#Dates
library(kableExtra) #Tables
library(glmnet)
```

#### Data

```{r}
dust<- read_csv(here("data", "processed_data","dust_master.csv")) #Different Dust attributes
data<- read_csv(here("data", "processed_data", "tx_master.csv")) #Master Data
```

# Cleaning

#### Clean and Trim

```{r}
dust<- dust %>%
  filter(!is.na(date),
         date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))

data<- data %>%
  filter(date %in% as.Date("2022-07-04"):as.Date("2022-07-19"))
```

#### Make into one data set

```{r}
master<- 
  dust %>% full_join(data) %>%
  filter(hr_cst %in% "01") #While this is technically not for hour 1, this gets ride of all duplicates for the sake of analysis.

master<- master %>%
  select(site, sal, temp, copies_mL,t7sum, date, orthop,nn, don,doc) %>%
  mutate(t7sum = log10(t7sum))
```

#### Site-specific Data sets and Clean df

```{r}
b<- master %>%
  filter(site %in% "Blind Oso" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
b<- b %>%
mutate(
  lag =Lag(b$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")

c<- master %>%
  filter(site %in% "Canals" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
c<- c %>%
mutate(
  lag =Lag(c$t7sum, shift  = 2)) %>% #Create 2 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")

r<- master %>%
  filter(site %in% "Gulf" | date %in% as.Date('2022-07-05') : as.Date('2022-07-06'))
r<- r %>%
mutate(
  lag =Lag(r$t7sum, shift  = 1)) %>% #Create 1 day lag
filter(between(date, as.Date('2022-07-07'), as.Date('2022-07-19'))) %>%
select(!c(date,t7sum,site)) %>%
  rename("log10dust" = "lag")
```

# Models

## 0. Null Model:

### Create 5 Fold

```{r}
fold_bo <- vfold_cv(b, v = 5, repeats = 5) 
fold_c2 <- vfold_cv(c, v = 5, repeats = 5) 
fold_rd <- vfold_cv(r, v = 5, repeats = 5) 
```

### Make Models

```{r, include= FALSE}
#Create Null Recipe for Growth at all three sites
null_bo<- recipe(copies_mL ~ 1, data = b) 
null_c2<- recipe(copies_mL ~ 1, data = c) 
null_rd<- recipe(copies_mL ~ 1, data = r)

#Set up linear model
lm_mod<- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

#Workflow that adds recipe to model for three sites
null_wflow_bo<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_bo)

null_wflow_c2<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_c2)

null_wflow_rd<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_rd)

#Use workflow to fit model to  data set
null_fit_bo<-
  fit_resamples(null_wflow_bo, resamples = fold_bo)

null_fit_c2<-
  fit_resamples(null_wflow_c2, resamples = fold_c2)

null_fit_rd<-
  fit_resamples(null_wflow_rd, resamples = fold_rd)

```

### Get Metrics for Null

#### RMSE

```{r}
null_met_bo <- 
  collect_metrics(null_fit_bo)
null_met_bo

null_met_c2 <- 
  collect_metrics(null_fit_c2)
null_met_c2

null_met_rd <- 
  collect_metrics(null_fit_rd)
null_met_rd
```

#### AIC

```{r}
bo_null_mod<- 
  lm(copies_mL ~ 1, data = b)
AIC(bo_null_mod) %>% round(digits = 2)

c2_null_mod<- 
  lm(copies_mL ~ 1, data = c)
AIC(c2_null_mod) %>% round(digits = 2)

rd_null_mod<- 
  lm(copies_mL ~ 1, data = r)
AIC(rd_null_mod) %>% round(digits = 2)
```

## Blind Oso Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = b)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_bo,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(b) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(b) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```

## Canals Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = c)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_c2,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(c) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(c) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```

## Gulf Random Forest

#### Recipe

```{r}
rec<- recipe(copies_mL ~ ., data = r)
```

```{r}
cores <- parallel::detectCores()
cores
```

#### Specify Model

```{r}
rf_mod <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_engine("ranger", num.threads = cores) %>% 
  set_mode("regression")
```

#### Creating Workflow

```{r}
rf_wf <- workflow() %>%
  add_model(rf_mod) %>%
  add_recipe(rec)
```

#### Create Tuning Grid

```{r}
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6),
                        min_n = c(40,50,60), 
                        trees = c(500,1000)  )
```

#### Cross-validation

```{r}
rf_resample <- 
  rf_wf %>% 
  tune_grid(fold_rd,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(rmse))
```

```{r}
rf_resample %>%
  collect_metrics()
```

##### Plot Model Performance

```{r}
#Plot of actual train data
rf_resample %>%
  autoplot()
```

##### Showing and Selecting Best Performing Models

```{r}
#Showing best performing tree models
rf_resample %>%
  show_best(n=1)

#Selects best performing model
best_rf <- rf_resample %>%
  select_best(method = "rmse")
```

##### Create Final Fit

```{r}
rf_final_wf <- 
  rf_wf %>% 
  finalize_workflow(best_rf)

#Create workflow for fitting model to train_data2 predictions
rf_final_fit <- 
  rf_final_wf %>%
  fit(r) 
```

##### Calculate Residuals

```{r}
rf_residuals <- rf_final_fit %>%
  augment(r) %>% #use augment() to make predictions from train data
  select(c(.pred, copies_mL)) %>%
  mutate(.resid = copies_mL - .pred) #calculate residuals and make new row.

rf_residuals
```

##### Model Predictions from Tuned Model vs Actual Outcomes

```{r}
rf_pred_plot <- ggplot(rf_residuals, 
                          aes(x = copies_mL, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Random Forest", 
       x = "Copies/mL Actual", 
       y = "Copies/mL Prediction")
rf_pred_plot

rf_residual_plot <- ggplot(rf_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Random Forest", 
       x = "Copies/mL Prediction", 
       y = "Residuals")
plot(rf_residual_plot) #view plot
```

### 2023-04-19 code/analysis_code/stats/4_uni_variate.qmd
edits

@@
 format:
-  html:
+  pdf:
     theme: default
@@
 library(ggpubr) #For plot stacking
+library(kableExtra) #Tables
 ```
@@
                rr.digits = 4)+
-  labs(title = "LM DustxCopies (BO+1)",
-       x = "log(dust)")
+  stat_correlation(method = "pearson",
+                   label.x = 0.05,
+                   label.y = 0.7,
+                   small.r = TRUE)+
+  labs(title = "Blind Oso",
+       x = "log10(dust)",
+       y = "Copies per mL")+
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank())
  fig2<-
@@
                rr.digits = 4)+
-  labs(title = "LM DustxCopies (C2+2)",
-       x = "log(dust)")
+  stat_correlation(method = "pearson",
+                   label.x = 0.05,
+                   label.y = 0.7,
+                   small.r = TRUE)+
+  labs(title = "Canals",
+       x = "log10(dust)",
+       y = "Copies per mL") +
+   theme_bw()+
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.ticks.x = element_blank())
+
 fig3<-
@@
                rr.digits = 4)+
-  labs(title = "LM DustxCopies (RD+1)",
-       x = "log(dust)")
+  stat_correlation(method = "pearson",
+                   label.x = 0.05,
+                   label.y = 0.7,
+                   small.r = TRUE)+
+  labs(title = "Gulf",
+       x = "log10(dust)",
+       y = "Copies per mL")+
+   theme_bw()+
+  theme(axis.text.x = element_blank(),
+        #axis.title.x = element_blank(),
+        axis.ticks.x = element_blank())
 
@@
 
-#### Stac
+#### Stack
 
-```{r}
+```{r, fig.height=4}
 ggarrange(fig1, fig2, fig3 + font("x.text", size = 10), ncol = 1, nrow = 3)
@@
 null_met_rd
+
+rmnb<- 20927.53
+rmnc<- 7934.59
+rmnr<- 49831.58
+
 ```
@@
 AIC(rd_null_mod)
+
+aicnb<- 275.93
+aicnc<- 253.19
+aicnr<- 296.98
 ```
@@
 
+rmb<- 14422.91
+rmc<- 5191.22
+rmr<- 36132.11
+
 ```
@@
 AIC(rd_mod)
+
+aicb<- 269.89
+aicc<- 245.37
+aicr<- 291.93
 ```
@@
 
-<!--# Why are these so high? -->
+#### Make Table
+
+```{r}
+kable(data.frame(Model = c("Null RMSE", "Actual RMSE", "Null AIC ", "Actual AIC"),
+                 Blind_Oso = c(rmnb,rmb,aicnb, aicb),
+                 Canals = c(rmnc, rmc, aicnc, aicc),
+                 Gulf = c(rmnr, rmr, aicnr, aicr)),
+      caption = "Linear Regression Model Metrics") 
+```
 

### 2023-04-19 code/analysis_code/stats/5_multi_variate_models/rf_models.qmd
edits

@@
 null_met_rd
+
+rmnb<- 20178.47
+rmnc<- 8091.394
+rmnr<- 52065.59
 ```
@@
   select_best(method = "rmse")
+rmb<- 15645.98
 ```
@@
   select_best(method = "rmse")
+rmc<- 7038.47
 ```
@@
   select_best(method = "rmse")
+rmr<- 33826.74
 ```
@@
 ```
+
+# Conclusion
+
+#### Table
+
+```{r}
+kable(data.frame(Model = c("Null RMSE", "Actual RMSE"),
+                 Blind_Oso = c(rmnb,rmb),
+                 Canals = c(rmnc, rmc),
+                 Gulf = c(rmnr, rmr)),
+      caption = "Random Forest Regression Model Metrics")
+
+```

### 2023-04-19 products/manuscript/Manuscript.qmd
+ RF model
edits

@@
 There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
-An initial shift in copies per mL at the Gulf and Blind Oso occur 24-48hr following initial dust introduction on the 13th, and experience another shift around 24-48hr of higher dust input on the 16th
 
@@
 
+Environmental parameters were site-specific.
+Temperature and salinities were consistent and high at the Canals ( <!--# Add ranges -->).
+While salinity remained relatively constant at the Gulf , temperature was observed to drop starting on the 12th of July, possibly due to a coastal upwelling event.
+Blind Oso exhibited fluctuations in both temperature and salinity within the daily time series with a drop in salinity from 42 to 20 within an eight day period.
+
+Nutrients (orthophosphate and nitrate+nitrite) were observed to increase following dust deposition at both Blind Oso and the Canals site, with the former having more dramatic peaks.
+Nutrient levels at the Gulf remained relatively constant throughout the time series.
+DON and DOC had distinct peaks at Blind Oso on days following dust deposition, where DON exhibited smaller peaks at the Canals.
+
 ![Figure 2: Environmental parameters across daily time series](images/00000f%20(1).png)
-
-Temperature and salinity at the Canal site is relatively high and consistent as compared to the other sites.
-Blind Oso Bay experiences some shifts in temperatures and drastic fluctuations in salinity.
-The Gulf site had a drop in temperature (possibly due to upwelling) and relatively consistent ocean salinities.
-
-\*need to add details on other environmental variables
 
@@
 
-<!--# Add these in -->
+Data that is normally distributed was run with Pearson's Correlation Coefficient. Data that is not normally distributed was run with Spearman's Correlation Coefficient.
+
+![Figure 3](images/000081.png)
 
@@
 
-![Figure 3: Linear regression models examining the relationship between dust and vibrio growth](images/Screen%20Shot%202023-04-17%20at%2011.21.32%20AM.png)
+![Figure 4](images/000055.png){width="481"}
 
-Above are linear regressions that examine the relationship between dust concentration and copies_mL.
-As mentioned above, the lags identified may not correspond to the most significant relationship between the two variables.
+![Table 1: Univariate Linear Regression Model Metrics](images/Screen%20Shot%202023-04-19%20at%201.33.37%20PM.png){width="367"}
 
-<!--# Update table -->
+### Multi-variate Models
 
-| Site | Model Type | RMSE     | AIC      |
-|------|------------|----------|----------|
-| BO   | Null       | 20472.82 | 275.9331 |
-| BO   | Actual     | 14422.91 | 269.8923 |
-| C2   | Null       | 8472.898 | 253.1876 |
-| C2   | Actual     | 5191.22  | 245.3679 |
-| RD   | Null       | 50386.56 | 296.9821 |
-| RD   | Actual     | 36132.11 | 291.933  |
+#### Random Forest
 
-[A few things to note:]{.ul}
+![Table 2: Random Forest Model Metrics](images/Screen%20Shot%202023-04-19%20at%202.08.25%20PM.png){width="367"}
 
--   My stats code contains some chunks that create predicted values based off of the current data set.
-Since my data set is small, I am not able to portion of \~25% to use to test my model as it could impact the strength of my model.
-Therefore, plotting the predicted vs.
-actual will give an R2 of 1.00.
-I am using AIC as a metric of model performance and comparing data produced models to null models.
+#### Linear Regressions
 
-    -   Will need to interpret findings as exploratory and hypothesis generating, and will need to be careful about trying to draw generalizable conclusions.
+![Table 3: Blind Oso: Comparative linear regression analysis of Vibrio growth between six operational models.](images/Screen%20Shot%202023-04-19%20at%201.36.29%20PM.png){width="367"}
 
-### Multi-variate Linear Regressions
+![Table 4: Canals: Comparative linear regression analysis of Vibrio growth between six operational models.](images/Screen%20Shot%202023-04-19%20at%201.37.02%20PM.png){width="367"}
 
-![](images/Screen%20Shot%202023-04-17%20at%203.16.17%20PM.png)
+![Table 5: Gulf: Comparative linear regression analysis of Vibrio growth between six operational models.](images/Screen%20Shot%202023-04-19%20at%201.37.21%20PM.png){width="367"}
 
-![](images/Screen%20Shot%202023-04-17%20at%203.17.06%20PM.png)
-
-![](images/Screen%20Shot%202023-04-17%20at%203.17.43%20PM.png)
-
-<!--# Fix Table 1 titles -->
+Multi-variate models that included dust, water chemistry, nutrients, and DOM performed the best, with mean RMSE values much lower than the Null Model.
 
@@
 
-#### Environmental Parameters
-
-Environmental parameters were site-specific.
-Temperature and salinities were consistent and high at the Canals ( <!--# Add ranges -->).
-While salinity remained relatively constant at the Gulf , temperature was observed to drop starting on the 12th of July, possibly due to a coastal upwelling event.
-Blind Oso exhibited fluctuations in both temperature and salinity within the daily time series with a drop in salinity from 42 to 20 within an eight day period.
-
-Nutrients (orthophosphate and nitrate+nitrite) were observed to increase following dust deposition at both Blind Oso and the Canals site, with the former having more dramatic peaks.
-Nutrient levels at the Gulf remained relatively constant throughout the time series.
-DON and DOC had distinct peaks at Blind Oso on days following dust deposition, however the other two sites were relatively stable.
-
 #### Vibrio Enumeration
@@
 Overall, *Vibrio* growth was observed following dust events at both the Gulf and Blind Oso.
-Peaks in growth were observed 24-48 hours following dust deposition.
+Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
 Linear regression analysis confirmed a correlation between growth and dust input (R2 = 0.49, 0.56, 0.44 for Blind Oso, Canals, and Gulf respectively).
+
+#### Site-Specific Dynamics
+
 However, the environmental dynamics of each site most likely contributed the degree of response per site.
@@
 
-## Conclusions
+## Conclusions/Broader Impacts
 

### 2023-04-20 code/analysis_code/env/env_par.qmd
mansuscript edits

@@
   theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-   axis.ticks.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+   #axis.ticks.x = element_blank(),
     axis.title.y = element_text(size = 15),
@@
   theme(
-    axis.title.x = element_blank(),
-   # axis.text.x = element_blank(),
-    axis.ticks.x = element_blank(),
+    #axis.title.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    #axis.ticks.x = element_blank(),
     axis.title.y = element_text(size = 15),
@@
     plot.title= element_text(hjust = 0.5, size = 20),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-   # axis.text.x = element_blank(),
+   # axis.title.x = element_blank(),
+    #axis.ticks.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title.y = element_text(size = 15),
@@
     plot.title= element_text(hjust = 0.5, size = 20),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-   # axis.text.x = element_blank(),
+    #axis.title.x = element_blank(),
+   # axis.ticks.x = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1),
     axis.title.y = element_text(size = 15),
@@
 ```{r, fig.width= 8, fig.height=4}
-fig<- ggarrange(#temp,
-                #sal,
-               # do,
-                #nn,
-                #ortho,
-                #don,
-                #doc,
+fig<- ggarrange(temp,
+                sal,
+                #do,
+                nn,
+                ortho,
+                don,
+                doc,
                 #pH,
                 #amm,
-                tn,
-                sil,
-                cop,
+                #tn,
+                #sil,
+                #cop,
+                tdn,
+                #secchi,
+                #din,
+                #din_dip,
                 chl,
-                secchi,
-                din,
-                din_dip,
-                tdn,
-                ncol = 3, 
-                nrow = 3,
+                ncol = 4, 
+                nrow = 2,
                 align = "v")

### 2023-04-20 products/manuscript/Manuscript.qmd
mansuscript edits

@@
 
-![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png)
+![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png){width="1491"}
 
@@
 
-![Figure 2: Environmental parameters across daily time series](images/00000f%20(1).png)
-
-<!--# Add detail on other env variables -->
+![Figure 2:](images/000019.png)
 
@@
 
-![Figure 4](images/000055.png){width="481"}
+Linear regression models between dust AOD and copies per mL of Vibrio.
+R2 values indicates the strength of the linear model, and the Pearson Correlation Coefficient indicates the strength of the linear relationship between the two variables.
 
-![Table 1: Univariate Linear Regression Model Metrics](images/Screen%20Shot%202023-04-19%20at%201.33.37%20PM.png){width="367"}
+![Figure 4](images/000055.png){width="1491"}
+
+![Table 1: Uni-variate Linear Regression Model Metrics](images/Screen%20Shot%202023-04-19%20at%201.33.37%20PM.png){width="367"}
 
@@
 
-#### Vibrio Enumeration
+#### Vibrio and Dust
 
-Overall, *Vibrio* growth was observed following dust events at both the Gulf and Blind Oso.
-Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
-Linear regression analysis confirmed a correlation between growth and dust input (R2 = 0.49, 0.56, 0.44 for Blind Oso, Canals, and Gulf respectively).
+*Vibrio* growth was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02).
+Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition, with both the Gulf and Blind Oso having a peak response 24hr following peak dust deposition, and the Canals 48hr after peak deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
+<!--# Add xfold increases -->.
+
+The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for growth.
+*Vibrio* species have one of the fastest doubling times among bacteria \[@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to rapidly exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes \[@ringgaardChemotaxisArraysVibrio2018 ; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016\].
+These advantageous traits have resulted in *Vibrio* being classified as 'oppurtunitrophs'.
+The following declines in *Vibrio* gene copies/mL may be attributed to topdown control, such as viral lysis <!--# Cite -->, or bottom-up controls due to an exhaustion of the resources provided by the dust.
 
@@
 
-However, the environmental dynamics of each site most likely contributed the degree of response per site.
-Multivariate models were tested using specific environmental parameters: dust, DOM (DON, DOC), temperature, salinity, orthophosphate, and nitrate+nitrite.
-Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (R2 = 0.73, 0.7, and 0.76 for Blind Oso, Canals, and Gulf respectively), suggesting that while dust may contribute to a growth response in *Vibrio* following dust input, the site specific dynamics aid in dictating the degree of response observed.
+Multivariate models were tested using specific environmental parameters that have been associated with *Vibrio* population dynamics [@bullington2022].
+Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (Table 3, Table 4, Table5), suggesting that while dust may contribute to a growth response in *Vibrio*, the site specific nutrient and chemistry dynamics play an important role in dictating the degree of response observed.
+
+<!--# Add monthly average of nutrients -->
+
+Similar to previous work done in coastal and offshore sites, the Gulf experienced
+
+Blind Oso Bay is an enclosed shallow tributary that frequently expereince high levels of nutrients (total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP).
+The Bay experienced a dramatic drop in salinity from July 11th to July 19th.
+This decrease in salinity could be attributed to discharge from the OSP.
 

### 2023-04-20 code/analysis_code/stats/3_correlation.qmd
edits

--- 2023-04-11 code/analysis_code/stats/3_correlation.qmd
@@
 library(lubridate) #For dates
+library(ggpubr) #Plot stacking
 ```
@@
 
-### Log10 Transform non-normally distributed data and Remove Non-Parametric Variables
-
-<!--# If we are using a nonparametric correlation (i.e. Spearman) we do not log-transform data -->
-
-Distributions can be found in `1_distribution.qmd`
+#### Make  datasets for MADA analysis
 
 ```{r}
-#Blind Oso
-#bo<-
-  #bo %>%
- # mutate(amm = log10(amm), #Log-transforms data
-        # din_dip = log10(din_dip),
-        # doc = log10(doc),
-        # dust = log10(dust),
-         #chl = log10(chl)) %>%
- # select(!c(secchi, tn, tdn, don)) #Removes non-parametric data
+b1<- 
+  bo %>%
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
 
-#Canals
-#c2<- 
-  #c2 %>%
-  #mutate(orthop = log10(orthop),
-         #dust = log10(dust)) %>%
-  #select(!c(secchi, nn, sil, doc,toc, tn))
+c1<- 
+  c2 %>%
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
 
-#Gulf
-#rd<- 
- # rd %>%
- # mutate(doc = log10(doc),
-      #   toc = log10(toc),
-        # dust = log10(dust)) %>%
-  #select(!c(sal, do_mgl, secchi, nn, tn))
+r1<- 
+  rd %>%
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
 ```
@@
 ```{r, include = FALSE}
-b<- cor(bo, method = "spearman") #Default method = pearson
-round(b,2) #Round correlation coeff to "x" digits
+b2<- cor(b1, method = "spearman") 
+round(b1,2) #Round correlation coeff to "x" digits
 
-c<- cor(c2, method = "spearman")
-round(c,2)
+c2<- cor(c1, method = "spearman")
+round(c1,2)
 
-r<- cor(rd, method = "spearman")
-round(r,2)
-```
-
-#### Make Function to Format
-
-Simple function for formatting a **correlation matrix** into a table with 4 columns containing : Column 1 : row names (variable 1 for the correlation test), Column 2 : column names (variable 2 for the correlation test), Column 3 : the **correlation coefficients**, Column 4 : the **p-values** of the correlations
-
-```{r}
-flattenCorrMatrix <- function(cormat, pmat) {
-  ut <- upper.tri(cormat)
-  data.frame(
-    row = rownames(cormat)[row(cormat)[ut]],
-    column = rownames(cormat)[col(cormat)[ut]],
-    cor  =(cormat)[ut],
-    p = pmat[ut])}
-```
-
-#### Make Matrix
-
-The function **rcorr()** \[in **Hmisc** package\] can be used to compute the **significance levels** for **pearson** and **spearman correlations**.
-It returns both the correlation coefficients and the p-value of the correlation for all possible pairs of columns in the data table.
-
-```{r}
-bclean<- rcorr(as.matrix(bo))
-flattenCorrMatrix(bclean$r, bclean$P)
-
-cclean<- rcorr(as.matrix(c2))
-flattenCorrMatrix(cclean$r, cclean$P)
-
-rclean<- rcorr(as.matrix(rd))
-flattenCorrMatrix(rclean$r, rclean$P)
+r2<- cor(r1, method = "spearman")
+round(r1,2)
 ```
@@
 
-<!--# This is using Spearman's with raw data -->
+```{r}
+par(mfrow=c(1,3))
 
-```{r}
-corrplot(b, type = "upper", 
-         tl.col = "black", tl.srt = 45,
+#Non-Parametric Data
+ corrplot(b2, type = "upper", 
+         tl.col = "black", tl.srt = 90,
          title = "Blind Oso",  mar=c(0,0,1,0)) #Adds and lowers title
 
-corrplot(c, type = "upper", 
-         tl.col = "black", tl.srt = 45,
+corrplot(c2, type = "upper", 
+         tl.col = "black", tl.srt = 90,
          title = "Canals", mar=c(0,0,1,0))
 
-corrplot(r, type = "upper", 
-         tl.col = "black", tl.srt = 45, mar=c(0,0,1,0))
+corrplot(r2, type = "upper", 
+         tl.col = "black", tl.srt = 90, mar=c(0,0,1,0))
          title(main = "Gulf" )
 ```
+
+#####
 

### 2023-04-20 products/manuscript/Manuscript.qmd
edits

@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year \[@dalmeidaModelSaharanDust1986\].
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico.
@@
 These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] .
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a\].
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
 
@@
 *Vibrio* species are found in brackish to marine waters with salinities ranging from 0-35 and will grow in waters with temperatures ranging from 5°C-40°C [@sampaioVibrioSppLife2022].
-Each year, plumes of Saharan dust travel across the Atlantic via easterly trade winds.
-These plumes are then deposited in the surface waters of the mid-Atlantic, Caribbean, and Gulf of Mexico.
-These dust aerosols harbor a wide range of fungi, bacteria, virus-like-particles, minerals, and nutrients (NO3, Fe, PO4).
-Addition of these nutrients to otherwise oligotrophic settings can result in large and rapid blooms of potentially harmful microbes, presenting a danger to both marine and human health.
 
@@
 parahaemolyticus*) and sea breams (*V.
-harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010\] among others.
+harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
 
@@
 
-<!--# Have paragraph describing and then have figures. -->
-
 ## Exploratory/Descriptive analysis
 
-#### Dust AOD and Vibrio Enumeration
+#### Dust AOD and Vibrio Enumeration <!--# What are AOD units? -->
 
@@
 
-Data that is normally distributed was run with Pearson's Correlation Coefficient. Data that is not normally distributed was run with Spearman's Correlation Coefficient.
+Correlation matrices were run on each variable (in order of left to right: temperature, salinity, nitrate+nitrite, orthophosphate, dissolved organic nitrogen, dissolved organic carbon, total dissolved nitrogen, chlorophyll, *Vibrio* gene copies/mL, and dust).
+A majority of the environmental variables had a non-normal distribution, thus Spearman's Correlation (indicated by the color gradient) was used to examine the relationship between the variables.
 
-![Figure 3](images/000081.png)
+![](images/000175.png)
 
@@
 
+<!--# Add RF graphs here -->
+
 {{< pagebreak >}}
@@
 
-This study aimed to examine microbial response to Saharan dust deposition in high nutrient coastal waters.
-Specifically focusing on bacteria belong to the genus *Vibrio*, we found that growth response following dust input varied by site.
-
 ## Summary and Interpretation
 
-### Methods
+This study aimed to examine microbial response to Saharan dust deposition in high nutrient coastal waters, specifically focusing on bacteria belong to the genus *Vibrio*, a bacterium that is important for marine processes such as carbon cycling, as well as human and marine health.
+
+## Methods
 
@@
 
-### Findings/Explanations
-
-<!--# What are AOD units? -->
-
-#### Vibrio and Dust
+## Findings/Explanations
 
 *Vibrio* growth was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02).
-Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition, with both the Gulf and Blind Oso having a peak response 24hr following peak dust deposition, and the Canals 48hr after peak deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
-<!--# Add xfold increases -->.
-
 The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for growth.
-*Vibrio* species have one of the fastest doubling times among bacteria \[@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to rapidly exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes \[@ringgaardChemotaxisArraysVibrio2018 ; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016\].
+*Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to rapidly exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes \[@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
 These advantageous traits have resulted in *Vibrio* being classified as 'oppurtunitrophs'.
@@
 
-#### Site-Specific Dynamics
+Similar to previous findings, peaks in growth were observed 24-48 hours following dust deposition, with both the Gulf and Blind Oso having a peak response 24hr following peak dust deposition, and the Canals 48hr after peak deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
+<!--# Add xfold increases -->.
 
@@
 
-<!--# Add monthly average of nutrients -->
+*Vibrio* gene copies/mL at the Gulf experienced the greatest change in growth <!--# Add xfold --> as compared to the other two sites.
+This response in a lower nutrient marine environment lines up with previous work that was conducted in the Florida Keys and North Atlantic, where a 30-fold and 1.6-fold population increase was observed within a 24hr period following dust deposition.
 
-Similar to previous work done in coastal and offshore sites, the Gulf experienced
+While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 1).
+Field studies have suggested that salinities for optimal growth lie between 5-25, depending on the species [@sullivanEffectsTemperatureSalinity2018].
+The higher salinities at this site (40-42) may have attributed to this observed weaker growth response.
 
-Blind Oso Bay is an enclosed shallow tributary that frequently expereince high levels of nutrients (total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP).
-The Bay experienced a dramatic drop in salinity from July 11th to July 19th.
-This decrease in salinity could be attributed to discharge from the OSP.
-
-The Canals may have a heavily dampened growth response due to a high and constant salinity of 40-42.
-In general, optimal growth occurs at salinities of \~25-35, and less tolerant species will drop as salinities climb <!--# CITE --> .
+*Vibrio* growth was observed at Blind Oso Bay, but slightly dampened in comparison to the Gulf.
+This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) [@wetz].
+The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP.
+While growth was significantly correlated with dust deposition, there are many environmental factors at play at this site, making it difficult to identify the specific mechanisms that are eliciting the growth response.
 
@@
 Therefore the introduction on nutrients at such a site may result in rapid uptake and utilization.
-At Blind Oso, these dust constituents may not be as essential for growth, and the dust may have reached a maximum in how much it can support growth.
+At Blind Oso, a priming effect may be occurring, where these dust constituents may not be as essential for growth following an initial deposition, reaching a maximum in how much it can support growth.
+
+<!--# Add monthly average of nutrients -->
 
@@
 
-[**One important thing to note**]{.ul}: Our data set contains dust data for all of 2022.
-Our copies_mL data only contains values from 7/7/22 --> 7/19/22.
-Therefore, these CCA were made with only this time series.
-This limits our analysis, as we have dust data prior that can explain copies_mL on the earlier dates 7/7 or 7/8.
-But we cannot incorporate this into the CCA as we have NA for copies_mL for those early days.
-Below we will run linear models that DO contain these prior dust days.
-And since we have identified a lag, we can shift the dust data to line up with the copies_mL that it corresponds with directly - and we no longer have NAs.
-But this ultimately means that the lag identified above.
-may not be the best fit for our model below, so some additional work is needed that tests other lags in the code.
+<!--# Difficult to parse out sites  -->
 
@@
 
-## Conclusions/Broader Impacts
+## Conclusions
+
+Ultimately, *Vibrio* growth was observed following dust deposition in each of these sites.
+However, the degree of response varied by site.
+This may be explained by each site having a specific and complex environment with varying water chemistry and nutrient levels.
+
+Sites like the Gulf, that have lower nutrient levels, may have a more responsive microbial population, as they are temporarily relieved form nutrient limitation.
+Inland sites similar to the Canals or Blind Oso may exhibit dampened *Vibrio* growth following dust input, as the dust derived nutrients may be less critical for growth, as there is already a supporting nutrient foundation, or the environmental conditions may not be favorable for growth initially.
+The growth response at Blind Oso, despite having higher nutrients, suggests that there may be other constiuents in the dust that are not present in the site that elicit this response.
+More work will be needed to fully understand the dynamics, as there were many other attributes that may have also contributed to growth in the Bay.
+
+To date, our understanding of *Vibrio* and other microbial response to dust input has focused on oligotrophic settings, where dust input is considered to be a critical source of limiting nutrients, and iron in particular.
+This study aided in filling the gap in how comparable previous findings in oligotrophic settings are to coastal areas with higher ambient nutrient levels.
+The composition of desert dust can be complex and may have the potential to deliver critical resources that could be more readily exploited by opportunistic bacteria (like *Vibrio*) and elicit a population bloom despite higher baseline nutrient levels.
+During these bloom periods, human exposure through recreation or consumption of raw seafood may be more likely, presenting a risk to the health of the public.
 

### 2023-04-20 code/analysis_code/dust/dust_and_copies.qmd
Addedto env data viz
edits

@@
     x = "Date",
-    y = "Dust Aerosol Optical Density")+
+    y = "Dust Aerosol Optical Density",
+    title = "Dust AOD over Daily Time Series")+
   theme(
@@
     axis.title.x = element_text(size = 15),
-    axis.text.x = element_text(size = 10))
+    axis.text.x = element_text(size = 10),
+    plot.title = element_text(hjust = 0.5, size = 20))
 
@@
     plot.title = element_text(hjust = 0.5, size = 20),  #center title
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
+  # axis.text.x = element_blank(),
+   # axis.title.x = element_blank(),
+   # axis.ticks.x = element_blank(),
     axis.title.y = element_text(size = 10),

### 2023-04-20 products/manuscript/Manuscript.qmd
edits

@@
 
-*What I am looking for* I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels.
+I expect to see a more dramatic growth response in the Gulf (low nutrient), whereas the higher nutrient sites may exhibit a dampened growth response since they already have high baseline nutrient levels.
 If we still see a growth response despite already having supportive background nutrient levels at these sites, this may suggest that there are other constituents in the dust that can elicit a growth response.
 Nutrients, dust input, salinity, and temperature will be the primary factors of interest.
-
-*How I will analyze it* - Look at Vibrio growth over time series (estimated from qPCR) - Examine relationships between dust input and Vibrio growth - Examine influence of site-specific environmental parameters (Temperature, Salinity, Nutrients) on growth response - Run linear models on factors like: dust x growth and nutrients x growth - NMDS to see which parameters have the strongest influence on growth
 
@@
 Following transport to the laboratory, the samples were concentrated onto 0.2µm pore size, hydrophobic polycarbonate membranes (25mm in diameter) and stored at -80°C until DNA Extraction using a ZymoBIOMICS DNA Miniprep Kit (Cat#:4300).
+To evalulate contamination in the lab, negative controls (DI and MilliQ Water) were introduced at sample collection, filtration, DNA Extraction, and downstream quantitative polymerase chain reaction (qPCR).
 
@@
 
-Total *Vibrio* concentrations were quantified using a SYBR Green quantitative PCR (qPCR) method.
+Total *Vibrio* concentrations were quantified using a SYBR Green qPCR method.
 Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) PCR product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3.
@@
 Cycle threshold values for each qPCR test was compared to a standard curve representing 10\^1 to 10\^6 gene copies per reaction volume.
-<!--# Add Std Curve specs -->
 
@@
 
-## Exploratory/Descriptive analysis
+## Exploratory Analysis
 
-#### Dust AOD and Vibrio Enumeration <!--# What are AOD units? -->
+### Visualization
+
+#### Dust AOD <!--# What are AOD units? -->
 
@@
 Dust AOD returned to lower concentrations following the time series.
-The top portion of the figure depicts total enumerated Vibrio as copies per mL, with color by site.
+
+![Figure 1:](images/000009-01.png)
+
+#### Vibrio Enumeration
+
+This figure depicts total enumerated Vibrio as copies per mL, with color by site.
 There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
 
-![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png){width="1491"}
+![Figure 2:](images/000005-03.png)
 
@@
 
-![Figure 2:](images/000019.png)
+![Figure 3:](images/000019.png)
 
-## Statistical analysis
+### Statistical analysis
 
-### Check Distributions:
+#### Check Distributions:
 
@@
 
-### Cross-correlation Analysis:
+#### Cross-correlation Analysis:
 
@@
 
-### Correlation Matrices
+#### Correlation Matrices
 
@@
 
-![](images/000175.png)
+![Figure 4:](images/000175.png)
+
+## Full analysis
 
@@
 
-![Figure 4](images/000055.png){width="1491"}
+![Figure 5:](images/000055.png){width="1491"}
 
@@
 ![Table 2: Random Forest Model Metrics](images/Screen%20Shot%202023-04-19%20at%202.08.25%20PM.png){width="367"}
+
+![Figure 6:](images/000017-02.png)
+
+![Figure 7:](images/000007-02.png)
 
@@
 
-## Full analysis
+##
 
@@
 
-While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 1).
+While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 2).
 Field studies have suggested that salinities for optimal growth lie between 5-25, depending on the species [@sullivanEffectsTemperatureSalinity2018].
@@
 
-## Strengths and Limitations
+## Limitations
 
-At Blind Oso for example, the salinity is observed to drop from 42-20 within an eight day period that coincided with dust input.
-It is difficult to parse these two variables out as either or both could have contributed to growth at this site.
+One primary limitation of this project was the small samples size (n = 12 per site).
+One sample per day limits the interpretation of the results of this project and changes the scope from an explanatory report to exploratory.
 
-<!--# Difficult to parse out sites  -->
+Additionally, while qPCR can reliably detect the concentration of DNA of the target, the results do not differentiate between living and dead cells, and therefore provides an estimate of quantification.
 
-<!--# Sample size  -->
+While it can be proposed that the degree of microbial response is contingent on the environmental settings of the site, further work is needed to look at each specific environmental parameter and how it may impact *Vibrio* abundance.
+At Blind Oso for example, the drastic drop in salinity during an eight day period coincided with dust input.
+It then becomes difficult to parse these two variables out as either may have contributed to growth at this site.
 

### 2023-04-21 code/analysis_code/stats/3_correlation.qmd
Edits for Manuscript Review

@@
   select(!site) %>% 
-  mutate(dust =Lag(dust, shift  = 1)) %>% #Lag from cca.qmd
+  mutate(dust_lag =Lag(dust, shift  = 1)) %>% #Lag from cca.qmd
   filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
@@
   select(!site) %>% 
-  mutate(dust =Lag(dust, shift  = 2)) %>%
+  mutate(dust_lag =Lag(dust, shift  = 2)) %>%
   filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
@@
   select(!site) %>% 
-  mutate(dust =Lag(dust, shift  = 1)) %>%
+  mutate(dust_lag =Lag(dust, shift  = 1)) %>%
   filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19')) %>%
@@
   bo %>%
-  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust, dust_lag)
 
@@
   c2 %>%
-  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust, dust_lag)
 
@@
   rd %>%
-  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust)
+  select(temp, sal, nn, orthop, don, doc, tdn, chl, copies_mL, dust, dust_lag)
 ```

### 2023-04-21 products/manuscript/Manuscript.qmd
Edits for Manuscript Review

@@
 ---
-
-# Edits
-
-1.
-I would do a lot of visual inspections to see how the different models fit and what they predict.
-You can try a linear model and a more complex ML one and compare to see if maybe something beyond a linear relationship improves results.
-
-2.
-Given that you don't have a lot of data, I'm not sure a PCA makes sense.
-I think in-depth analysis of the relation between the variables and the outcome and then in addition to the linear models trying a few more complex ones will probably be the better approach.
-
-    <!--# Try a model that is copies_mL ~. -->
-
-3.  The figures in the manuscript look good, but some are too small to read. Needs breaking up and/or increasing font size
 
@@
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
-Dust aerosols can harbor a wide diversity of bacteria \[@kelloggCharacterizationAerosolizedBacteria2004\], fungi \[@ramirez-camejoFungalDiversitySahara2022\], virus-like particles \[@griffinAfricanDesertDust\], and minerals \[@formenti2003\].
-These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] .
-This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
+Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004\], fungi [@ramirez-camejoFungalDiversitySahara2022\], virus-like particles [@griffinAfricanDesertDust\], and minerals [@formentiChemicalCompositionMineral2003\].
+These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989; @grahamAtmosphericTransportPhosphorus1982\] .
+This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups [@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
 
@@
 alginolyticus*, *V.
-vulnificus*) [@baker-austinVibrioSppInfections2018\].
+vulnificus*) [@baker-austinVibrioSppInfections2018].
 Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V.
@@
 coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V.
-coralliilyticus*) \[@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003\], and mortality in important aquaculture species such as penaeid shrimp (*V.
+coralliilyticus*) [@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003], and mortality in important aquaculture species such as penaeid shrimp (*V.
 parahaemolyticus*) and sea breams (*V.
-harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
+harveyi)* [@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
 
@@
 
-How do Saharan dust events influence Vibrio populations in high nutrient coastal waters?
+**How do Saharan dust events influence Vibrio populations in high nutrient coastal waters?**
 
@@
 Sampling took place at three locations in Corpus Christi, TX: Blind Oso Bay, a residential canal system on Padre Island, and the Gulf.
-Sites were chosen based on differing baseline nutrient levels.
+Sites were chosen based on differing baseline nutrient levels (defined here as the 2022 average from monthly observations).
 Blind Oso Bay is a shallow tributary that is popular amongst wadefisherman and kayakers.
-The Bay is often impacted by fluctuating salinities and high levels of nutrients and chlorophyl due to its proximity to a wastewater treatment plant [@wetz] Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
-impaired waters list ([Nicolau & Hill, 2014](https://www.tceq.texas.gov/downloads/water-quality/tmdl/oso-bay-oyster-harvesting-assessment-103/103-oso-bay-monitoring-report-2013.pdf), [Texas Commission on Environmental Quality, 2022](https://www.tceq.texas.gov/downloads/water-quality/assessment/integrated-report-2022/2022-303d.pdf)).
-The residential canal system is also impacted by high nutrients and chlorophyl from storm water runoff, and salinity is primarily driven by precipitation.
-The Gulf site is characterized by constant salinities (\~36) and low nutrient levels.
-<!--# update these citations -->
+The Bay is often impacted by fluctuating salinities and high levels of nutrients (baseline:13.03 µM nitrate+nitrite, 4.44 µM orthophosphate) and chlorophyll due to its proximity to wastewater treatment discharge (Wetz, 2014).­ Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
+impaired waters list (Nicolau & Hill, 2014, Texas Commission on Environmental Quality, 2022).
+The residential canal system (baseline: 0.69µM nitrate+nitrite, 0.20 µM orthophosphate), and can also be impacted by high nutrients and chlorophyll from storm water runoff; salinity is primarily driven by precipitation.
+The Gulf site is characterized by constant salinities (\~36) and relatively low baseline nutrient levels (baseline: 0.19µM nitrate+nitrite and 0.36µM orthophosphate).
 
@@
 Total *Vibrio* concentrations were quantified using a SYBR Green qPCR method.
-Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) PCR product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3.
+Estimates were made using genus-specific quantitative polymerase chain reaction (qPCR) product using Vibrio group specific primers targeting a variable region of the 16S rRNA gene, 567F, 5′GGCGTAAAGCGCATGCAGGT3′ and 680R, 5′-GAAATTCTACCCCCCTCTACAG-3.
 Master mix and PCR conditions were derived from [@westrichSaharanDustNutrients2016a].
@@
 This project works with three primary data sets.
-(1) Enumeration of Vibrio bacteria as copies/mL from quantitative PCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi_2.csv).
+(1) Enumeration of Vibrio bacteria as copies/mL from qPCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi_2.csv).
 Below is a brief summary of how each data set was cleaned.
 More details can be found in the supplementary files, containing the code and comments describing what each line does.
-Please refer to the `project_README.md` file located in the NathanGreenslit-MADA-Project folder for details on the repository contents and instructions on reproducibility\
+Please refer to the `project_README.md` file located in the NathanGreenslit-MADA-Project folder for details on the repository contents and instructions on reproducibility.\
 
@@
 Dust concentration (AOT) was collected at time points 0 hour, 6 hour, 12 hour, and 18 hour.
-In this script, I converted timpoints from UTC to CST/CDT and created different dust permutations (individual time points, sums by day, and average by day).
+In this script, I converted timepoints from UTC to CST/CDT and created different dust permutations (individual time points, sums by day, and average by day).
 Because samples were collected between 0700 and 1100, a 24h composite permutation was made.
+As an example, Dust AOD for the morning of July 12th would be calculated using the last two time points of the 11th, and the first two time points of the 12th.
 This permutation is what was used for all downstream analyses.
@@
 
-{{< pagebreak >}}
-
 # Results
@@
 
-### Visualization
-
-#### Dust AOD <!--# What are AOD units? -->
+### Dust AOD
 
@@
 
-![Figure 1:](images/000009-01.png)
+![Figure 1: Dust Aerosol Optical Density (AOD) over daily time series.](images/000009-01.png)
 
-#### Vibrio Enumeration
+### Vibrio Enumeration
 
 This figure depicts total enumerated Vibrio as copies per mL, with color by site.
-There is a noticeable shift in *Vibrio* growth at the Gulf site, a less noticeable but still present shift in Blind Oso Bay, and practically no changes in the Canal site.
+There is a noticeable shift in *Vibrio* growth at the Gulf, a less noticeable but still present shift in Blind Oso Bay, and little fluctuation at the Canals.
 
-![Figure 2:](images/000005-03.png)
+![Figure 2: Vibrio gene copies/mL over daily time series across site.](images/000005-03.png)
 
-#### Environmental Parameters
+### Environmental Parameters
 
 Environmental parameters were site-specific.
-Temperature and salinities were consistent and high at the Canals ( <!--# Add ranges -->).
-While salinity remained relatively constant at the Gulf , temperature was observed to drop starting on the 12th of July, possibly due to a coastal upwelling event.
+Temperature and salinities were consistent and high at the Canals.
+While salinity remained relatively constant at the Gulf, temperature was observed to drop starting on the 12th of July, possibly due to a coastal upwelling event.
 Blind Oso exhibited fluctuations in both temperature and salinity within the daily time series with a drop in salinity from 42 to 20 within an eight day period.
@@
 Nutrient levels at the Gulf remained relatively constant throughout the time series.
-DON and DOC had distinct peaks at Blind Oso on days following dust deposition, where DON exhibited smaller peaks at the Canals.
+Dissolved organic matter (DOC, DON), TDN, and Chlorophyll exhibited distinct peaks in the latter half of the time series at Blind Oso, while less clear trends were seen in the other two sites.
 
-![Figure 3:](images/000019.png)
+![Figure 3: Envonmental parameters across daily time series with color by site.](images/000019.png)
 
-### Statistical analysis
+## Statistical Analysis
 
-#### Check Distributions:
+### Distributions
 
-Shapiro-Wilk's tests were conducted to assess distribution of the data.
+Density and Q-Q plots as well as Shapiro-Wilk's tests were conducted to assess distribution of the data.
 Any data that did not have a normal distribution was log-transformed and the test was run again.
 Variables that continue to fail normality will have non-parametric tests run on them.
+<!--# Add more details about visualization -->
 
-#### Cross-correlation Analysis:
+### Cross-correlation Analysis
 
@@
 
-#### Correlation Matrices
+### Correlation Matrices
 
-Correlation matrices were run on each variable (in order of left to right: temperature, salinity, nitrate+nitrite, orthophosphate, dissolved organic nitrogen, dissolved organic carbon, total dissolved nitrogen, chlorophyll, *Vibrio* gene copies/mL, and dust).
+Correlation matrices were run on each variable (in order of left to right: temperature, salinity, nitrate+nitrite, orthophosphate, dissolved organic nitrogen, dissolved organic carbon, total dissolved nitrogen, chlorophyll, *Vibrio* gene copies/mL, dust, and lagged dust for growth delay analysis).
 A majority of the environmental variables had a non-normal distribution, thus Spearman's Correlation (indicated by the color gradient) was used to examine the relationship between the variables.
 
-![Figure 4:](images/000175.png)
+![Figure 4: Correlation matrices of environmental parameteres by site.](images/000009-02.png)
 
@@
 
-![Figure 5:](images/000055.png){width="1491"}
+![Figure 5: Linear Regression Output for Dust AOD (log-transformed) and Vibrio gene copies/mL](images/000055.png){width="1491"}
 
@@
 
-![Figure 6:](images/000017-02.png)
+![Figure 6: Vibrio gene copies/mL at Blind Oso predicted by Random Forest](images/000017-02.png)
 
-![Figure 7:](images/000007-02.png)
+![Figure 7: Vibrio gene copies/mL at Canals predicted by Random Forest](images/000007-02.png)
+
+![Figure 8: Vibrio gene copies/mL at the Gulf predicted by Random Forest](images/000007-03.png)
 
@@
 
-##
-
-<!--# Add RF graphs here -->
-
-{{< pagebreak >}}
-
 # Discussion
 
-## Summary and Interpretation
-
-This study aimed to examine microbial response to Saharan dust deposition in high nutrient coastal waters, specifically focusing on bacteria belong to the genus *Vibrio*, a bacterium that is important for marine processes such as carbon cycling, as well as human and marine health.
-
-## Methods
+This study aimed to examine microbial response to Saharan dust deposition in high nutrient coastal waters, specifically focusing on bacteria belong to the genus *Vibrio*, a bacterium that is important for marine processes such as carbon cycling, and has implications for both marine and human health.
 
 For this analysis, we examined *Vibrio* growth dynamics in response to Saharan Dust deposition in three sites in Corpus Christi, TX that had differing baseline nutrient levels.
-The Gulf, with relatively constant salinities and low nutrient levels, Blind Oso Bay with high nutrients and fluctuating salinities, and a residential canal system with consistent temperature and salinities and nutrient levels.
-Samples were collected before, during, and after a dust event and *Vibrio* were enumerated using qPCR.
-Data on nutrient levels and other environmental parameters (e.g.
-salinity and temperature) were also collected during this period.
+To examine the impact of dust deposition on *Vibrio* growth, samples were collected before, during, and after a dust event and *Vibrio* were enumerated using qPCR.
+Data on nutrient levels and other environmental parameters including salinity and temperature, were also collected during this period.
 The three data sets (Vibrio Enumeration, Environmental parameters, and Dust Concentration) were then processed and cleaned for analysis.
@@
 
-*Vibrio* growth was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02).
-The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for growth.
-*Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to rapidly exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes \[@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
-These advantageous traits have resulted in *Vibrio* being classified as 'oppurtunitrophs'.
-The following declines in *Vibrio* gene copies/mL may be attributed to topdown control, such as viral lysis <!--# Cite -->, or bottom-up controls due to an exhaustion of the resources provided by the dust.
+*Vibrio* gene copies/mL was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02).
+The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for bloom formation.
+*Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes [@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
+Initially comprising a small proportion of the marine microbial communities (\<1%), *Vibrio* can come to quickly dominate, and thus have been classified as 'oppurtunitrophs'.
+The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quiroz2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psenner1992].
 
-Similar to previous findings, peaks in growth were observed 24-48 hours following dust deposition, with both the Gulf and Blind Oso having a peak response 24hr following peak dust deposition, and the Canals 48hr after peak deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
-<!--# Add xfold increases -->.
+Similar to previous findings set in the Florida Keys and North Atlantic [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a], peaks in growth were observed 24-48 hours following dust deposition.
+Comparing peak *Vibrio* gene copies/mL to initial levels on the 7th of July, the Gulf experienced a 338-fold increase in *Vibrio* gene copies/mL, Blind Oso a 3.4-fold increase, and the Canals a 1.3-fold increase.
 
-Multivariate models were tested using specific environmental parameters that have been associated with *Vibrio* population dynamics [@bullington2022].
+Multivariate models were tested using specific environmental parameters that have been associated with *Vibrio* population dynamics.
 Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (Table 3, Table 4, Table5), suggesting that while dust may contribute to a growth response in *Vibrio*, the site specific nutrient and chemistry dynamics play an important role in dictating the degree of response observed.
 
-*Vibrio* gene copies/mL at the Gulf experienced the greatest change in growth <!--# Add xfold --> as compared to the other two sites.
-This response in a lower nutrient marine environment lines up with previous work that was conducted in the Florida Keys and North Atlantic, where a 30-fold and 1.6-fold population increase was observed within a 24hr period following dust deposition.
+*Vibrio* gene copies/mL at the Gulf experienced the greatest degree in growth as compared to the other two sites.
+This response in a lower nutrient marine environment lines up with previous work that was conducted in the Florida Keys and North Atlantic, where a 30-fold and 1.6-fold population increase was observed within a 24hr period following dust deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a].
 
-While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 2).
+While *Vibrio* growth exhibited a strong correlation at the Canals, the degree of growth was lesser in comparison to the other sites (Figure 3).
 Field studies have suggested that salinities for optimal growth lie between 5-25, depending on the species [@sullivanEffectsTemperatureSalinity2018].
@@
 This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) [@wetz].
-The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP.
-While growth was significantly correlated with dust deposition, there are many environmental factors at play at this site, making it difficult to identify the specific mechanisms that are eliciting the growth response.
+The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP (Figure 3).
+While growth was significantly correlated with dust deposition, there are many environmental factors at play at this site, limiting the identification of the specific mechanisms that are eliciting this response.
 
-Additionally, the Gulf exhibited a sharper increase during the second, heavier dust event, while Blind Oso exhibited a smaller increase during the heavy dust event.
+The Gulf exhibited a sharper increase during the second, heavier dust event, while Blind Oso exhibited a smaller increase during the heavy dust event (Figure 2).
 This difference in response may be attributed to the nutrient dynamics of the sites.
-In marine waters, nutrients are typically limiting due to tight microbial recycling <!--# CITE -->.
+In marine waters, nutrients are typically limiting due to tight microbial recycling.
 Therefore the introduction on nutrients at such a site may result in rapid uptake and utilization.
-At Blind Oso, a priming effect may be occurring, where these dust constituents may not be as essential for growth following an initial deposition, reaching a maximum in how much it can support growth.
-
-<!--# Add monthly average of nutrients -->
+At Blind Oso, a priming effect may be occurring, where dust constituents resulting from a secondary flux may not be as essential for growth following an initial deposition event, where the constituents are reaching a maximum in how much they can support growth.
 
@@
 
+The annual dust data set indicates a period of slightly elevated AOD (0.16) on the 5th of July, prior to sampling.
+This light dust period could have the potential to impact initial *Vibrio* levels at the beginning of the time series, which may serve as an explanation for initial higher *Vibrio* gene copies/mL at the Canals, and why the correlation between dust and *Vibrio* gene copies/mL was so high at this site.
+Future work should expand upon the temporal resolution, allowing for an extended period of low dust AOD prior to sampling.
+
 ## Conclusions
@@
 
-Sites like the Gulf, that have lower nutrient levels, may have a more responsive microbial population, as they are temporarily relieved form nutrient limitation.
-Inland sites similar to the Canals or Blind Oso may exhibit dampened *Vibrio* growth following dust input, as the dust derived nutrients may be less critical for growth, as there is already a supporting nutrient foundation, or the environmental conditions may not be favorable for growth initially.
-The growth response at Blind Oso, despite having higher nutrients, suggests that there may be other constiuents in the dust that are not present in the site that elicit this response.
-More work will be needed to fully understand the dynamics, as there were many other attributes that may have also contributed to growth in the Bay.
+Sites like the Gulf, that have lower nutrient levels, may have a more responsive microbial population, as they are temporarily relieved from nutrient limitation.
+Inland sites, similar to the Canals or Blind Oso, may experience higher nutrient levels as a result of nutrient runoff and proximity to residential areas, and thus exhibit dampened *Vibrio* growth following dust input, as the dust derived nutrients may be less critical for growth if there is already a supporting nutrient foundation.
+Alternatively, and potentially in the case at the Canals, the environmental conditions may not be favorable for growth initially.
+The apparent growth response at Blind Oso despite having higher nutrients, suggests that there may be other constituents in the dust that are not present in the site that elicit this response.
+More work will be needed to fully understand the dynamics, as there were many other attributes that may have also contributed to growth at the Bay.
 
@@
 During these bloom periods, human exposure through recreation or consumption of raw seafood may be more likely, presenting a risk to the health of the public.
+Understanding the relationships that exist between dust deposition and potential harmful bacterial blooms in these regions can aid in modelling predictions of increased exposure risk to better protect public health.
 

### 2023-04-24 code/analysis_code/stats/5_multi_variate_models/random_forest_models.qmd
addressing Raqi's comments

--- 2023-04-19 code/analysis_code/stats/5_multi_variate_models/rf_models.qmd
@@
   geom_point() + 
-  labs(title = "Predictions vs Actual: Random Forest", 
+  labs(title = "Blind Oso: Predictions vs Actual: Random Forest", 
        x = "Copies/mL Actual", 
-       y = "Copies/mL Prediction")
+       y = "Copies/mL Prediction") +
+  theme_bw()
 rf_pred_plot
@@
   geom_point() + 
-  labs(title = "Predictions vs Residuals: Random Forest", 
+  labs(title = "Blind Oso: Predictions vs Residuals: Random Forest", 
        x = "Copies/mL Prediction", 
        y = "Residuals")
-plot(rf_residual_plot) #view plot
+plot(rf_residual_plot) + #view plot
+  theme_bw()
 ```
@@
   geom_point() + 
-  labs(title = "Predictions vs Actual: Random Forest", 
+  labs(title = "Canals: Predictions vs Actual: Random Forest", 
        x = "Copies/mL Actual", 
-       y = "Copies/mL Prediction")
+       y = "Copies/mL Prediction") +
+  theme_bw()
 rf_pred_plot
@@
   geom_point() + 
-  labs(title = "Predictions vs Actual: Random Forest", 
+  labs(title = "Gulf: Predictions vs Actual: Random Forest", 
        x = "Copies/mL Actual", 
-       y = "Copies/mL Prediction")
+       y = "Copies/mL Prediction") + 
+  theme_bw()
 rf_pred_plot

### 2023-04-27 reviews/FranciscoReview.qmd

---
title: Project Review Template 
author: RAQUEL FRANCISCO
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
editor: 
  markdown: 
    wrap: 72
---

# Overview

Title of project: Manuscript for a Data Analysis Project

Name of project author(s): Nathan Greenslit

Name of project reviewer: Raquel Francisco

# Specific project content evaluation

## Background, Context and Motivation

### Feedback and Comments

Very interesting topic. Strong introduction and justification. Simple and elegant use of regression model. Great use of models and exploratory graphs with a small data set that explores both qualitative and spatial-temporal data.

No errors seen in the introduction.

### Summary assessment

-   strong contextualization and motivation

## Question description

### Feedback and Comments

Strong hypothesis backed up by previous research. Excellent introduction that lead well into the knowledge gap and it's importance to answer.

### Summary assessment

-   question/hypotheses fully clear

## Data description

### Feedback and Comments

Reproducible! Data source provided and easily accessed. Data cleaning was simple and logical. Data visualization was great. Only qualm was that there were many folders for raw data and data cleaning that were a bit difficult to navigate.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

### Feedback and Comments

•	Data explore (i.e., 1. Dust, etc)

  o I really enjoyed all your raw data. I would consider in the future merging these data exploration files for ease similar to your stats folder. For example just having one data exploration files with all the qmds. I don’t see the benefit in so many folders.

•	Distribution

  o Loved this section. Very unique.

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

### Feedback and Comments

•	2_cca

  o Error line 79:

     Warning: Unknown or uninitialised column: copies_mL.Error in ts(x) : 'ts' object must have one or more observations

• Same error for 85, 91, 113, 120, 127

Maybe I’m missing a package? If so add to library. (PS I made sure to run this one second as instructed.)

•	Univariate

  o For having a small data set, your models look like they are working with you 😊

•	Multivariate

  o In your read me you state: 5_multi_variate.qmd: runs multivariate models on variable of interest (copies_mL) BUT it seem like you have a multivariate folder with multiple qmds, please correct statement

     Rf_models.qmd error line 221-226 (so rest of script wont run.)

     Error: All inputs to metric_set() must be functions. These inputs are not: (1).

### Summary assessment

-   strong and reasonable analysis

## Presentation

### Feedback and Comments

•	Manuscript

  o Line 42, I don’t think you italicize families (but double check)

### Summary assessment

-   results are very well presented

## Discussion/Conclusions

### Feedback and Comments

Strong conclusions that were re-enforced by both the raw data and the modeled data (both univariate and multivariate models).

### Summary assessment

- strong, complete and clear discussion

## Further comments

I think this was a well explored dataset with interesting results.

# Overall project content evaluation

## Structure

### Feedback and Comments

Readme files were excellent! Very easy to follow and allowed me to run
the majority of your code with little to no problems at all. It really
made everything look very intentional.

### Summary assessment

-   well structured

## Documentation

### Feedback and Comments

The code was clear and straightforward. The only issues I had with the organization was the large amount of subfolders, however, this may just be a personal preference.

### Summary assessment

-   fully and well documented

## Reproducibility

### Feedback and Comments

Very minor error messages came up when reproducing code (found above). Only one area was unable to render.

### Summary assessment

-   fully reproducible without issues

## Thoroughness

### Feedback and Comments

I found this manuscript incredibly thorough with lots of data exploration and an ample amount of analysis. I look forward to when this is published! I am sure it will be great!

### Summary assessment

-   strong level of thorougness

## Further comments

I believe this project was well done and implemented many of the
strategies review in the course. I believe it should receive full marks
(3/3).

### 2023-04-29 code/analysis_code/analysis_README.Rmd
Edits for Manuscript Review
made edsits to analaysis readme

--- 2023-03-22 code/analysis_code/analysis_README.Rmd
@@
-This folder contains code for analyzing cleaned data sets and aids in visualization of parameters of interest.\
-(1) The dust subfolder contains `daily_dust.qmd`, which looks at dust concentration over the daily time series. It also contains `dust_and_copies.qmd` which combine Vibrio enumeration with dust data.\
-(2) The env sub-folder contains `env_par.qmd` that examines temperature and salinity over the time series.\
-(3) The qPCR sub-folder contains `daily_qmd` that has data on Vibrio enumeration. This is also seen in the `dust_and_copies.qmd` in the dust subfolder \
-(4) The stats folder contains:
-- `distributiom.qmd`: Checks distribution of variables of interest
-- `multi_variate.qmd`: runs multivariate models on variable of interest (copies_mL)
-- `uni_variate.qmd` : runs univariate models on variable and predictors of interest.
+---
+editor_options: 
+  markdown: 
+    wrap: 72
+---
+
+This folder contains code for analyzing cleaned data sets and aids in
+visualization of parameters of interest.\
+(1) `dust` folder
+
+-    `daily_dust.qmd`: looks at dust concentration over the daily time
+    series. It also contains
+
+-   `dust_and_copies.qmd:` combines Vibrio enumeration with dust data.
+
+\
+(2) The `environment` folder
+
+-    `env_par.qmd`: examines environmental parameters of interest.
+
+\
+(3) The `qPCR` folder
+
+-    `daily_qmd`: has data on Vibrio enumeration. This is also seen in
+    the `dust_and_copies.qmd` in the dust subfolder
+
+\
+(4) The `stats` folder:
+
+-    `1_distributiom.qmd`: Checks distribution of variables of interest
+
+-   `2_cca.qmd`: Performs cross-correlation analysis to test for lags in
+    time series (between dust input and Vibrio growth)
+
+-    `3_correlation.qmd`: Runs Correlation matrices on all variables
+
+-    `4_uni_variate.qmd` : runs univariate models on variable and
+    predictors of interest.
+
+-    `5_multi_variate_models folder`: Contains scripts that runs
+    multivariate models on variable of interest (copies_mL) and random
+    forest models.

### 2023-05-01 reviews/LariscyReview.qmd
+Leah's Review

---
title: Project Review Template 
author: YOUR NAME
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Nathan Greenslit MADA Project

Name of project author(s): Nathan Greenslit

Name of project reviewer: Leah Lariscy

# Instructions

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

# Specific project content evaluation

Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Background was thoroughly explained. From your writing I can see your work is very relevant to public health near coastal waters and I have a good idea of what work has already been done and what you hope to accomplish. I can't say I would change much here. Good work!

### Summary assessment

-   strong contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

It is clear what your goal is in the analysis and is mostly clear how you plan to use your data to explore your hypothesis.

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

I personally would at least want to see a list of variables for each set, even if it just ends up in the supplementary file (especially for environmental factors since this wasn't discussed in detail). However, I though you did a fantastic job with the data source description as it was very detailed.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Data cleaning description was detailed, all steps made sense and were explained well. Exploratory analysis was clear and figures were appropriate (and looked really good).

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Your analysis methods seemed completely appropriate, and you did a great job explaining why you chose to do what you did.

### Summary assessment

-   strong and reasonable analysis

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Figures look professional and publishable.

### Summary assessment

-   results are very well presented

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Good enough to be published to be honest.. really great work!

### Summary assessment

-   strong, complete and clear discussion

## Further comments

*Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project.*

I would have liked to see more data description, but the project overall is solid and well done!

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Files were organized well and the README file was thorough

### Summary assessment

-   well structured

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The code was well-documented for the most part. There were comments on a large portion of the code explaining what the code was doing, and use of visual mode is especially helpful because sections were well laid out.

### Summary assessment

-   fully and well documented

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I did not run into any issues with reproducing scripts, there were detailed instructions left in the README file. However, results in the manuscript are not reproducible so that should be addressed.

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

This study seemed very thorough. Alternative analysis methods, limitations to the data, and hypotheses were fully discussed.

### Summary assessment

-   strong level of thoroughness

## Further comments

*Add any other comments regarding the overall project here. Write anything you think can help your classmate improve their project.*

Great job Nathan :-) this is seriously awesome work! I would make sure your images are reproducible in the manuscript and I would also change the title to include information about the data.

\`\`\`

### 2023-05-01 code/wrangling_code/wrangling_README.Rmd
Updated README and citations

--- 2023-02-24 code/wrangling_code/wrangling_README.Rmd
@@
-The folder contains various cleaning codes for data sets on environmental parameters (dust, temperature, salinity), and total Vibrio counts in sea surface waters.
+The folder contains various cleaning codes:
+
+\
+(1) `dust.qmd`: This code takes dust concentration data from NRL and cleans it.
+Steps include, switching to the correct timezone, creating concentration sums per day, as well as combining the dust data set with the *VIbrio* enumeration data set.
+
+\
+(2) `total_copies_calculator.qmd`: This script is for cleaning qPCR results on Total Vibrio enumeration.
+It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio.
+The data set contains a daily and monthly time series.
+We will be examining daily.
+
+\
+(3) `df_master.qmd`: This script combines the three clean data sets for dust concentration, copies per mL, and environmental parameters

### 2023-05-01 data/processed_data/processed_data_README.Rmd
Updated README and citations

--- 2023-03-14 data/processed_data/processed_data_README.Rmd
@@
----
+------------------------------------------------------------------------
 
 This folder contains processed data that has been cleaned from code in the `wrangling` folder.\
-(1) `total_clean_copies`: Vibrio enumeration (as copies per mL) across time series.\
-(2) `dust_clean.rds`: Cleaned dust data
-(3) `ysi_2.csv`: Environmental parameter data
-(4) `tx_master.csv` which is a compilation of all relevant data. 
-(5)`dust_master.csv`: Manually created dust data examining different time points, sums, avereages (helpful for statistics)
+
+## `total_clean_copies.csv`
+
+Vibrio enumeration as gene copies per mL across time series
+
+#### Variables
+
+date, site, type (monthly or daily time series), and copies_mL.
+
+## `dust_clean.rds`
+
+Cleaned dust data
+
+#### Variables
+
+date, SUM (AOT summed by date), hr_cst (hour of collection in CST), and dust (AOT)
+
+## `ysi_2.csv`
+
+Environmental parameter data-- did not need cleaning
+
+#### Variables
+
+site, date, chlorophyll (chl), salinity (sal), temperature (temp), Dissolved Oxygen in mg/L (do_mgl), Dissolved Oxygen in % (do_per), pH, secchi depth (secchi), ammonium (amm), Nitrite+Nitrate (nn), orthophosphate (orthop), silicate (sil), dissolved inorganic nitrogen (din), dissolved inorganic nitrogen:dissolved inorganic phosphorous (din_dip), total dissolved nitrogen (tdn), dissolved organic nitrogen (don), dissolved organic carbon (doc), total organic carbon (toc), total nitrogen (tn).
+
+## `tx_master.csv`
+
+A compilation of all cleaned variables mentioned above
+
+## `dust_master.csv`
+
+Manually created dust data examining different dust permutations: time points, sums, averages (helpful for statistics)
+
+#### Variables
+
+date, time at collection 1 hour (t1), 7 hour (t7), 13 hour (t13), and 19 hour (t19).
+Sum of all 4 time points (tsum), average of all four time points (tavg), 24 hour hour composite sum (t7sum), 24 hour composite average (t7sum).

### 2023-05-01 data/raw_data/rawdata_README.Rmd
editing for grammer and flow
Updated README and citations

--- 2023-02-24 data/raw_data/rawdata_README.Rmd
@@
-This folder contains dust concentrations in AOT from the Naval Reserach lab, environmental parameters (temperature and salinity) that were collected during the summer at the three sites, and  qPCR results for Total Vibrio.
+This folder contains two raw data sets:
+
+## `dust.csv`
+
+#### Variables:
+
+yyymmddtt: Year_month_day_time of collection
+
+dust: dust aerosol optical density. Concentrations (AOT) were collected at 4 time points (UTC): T0,T6,T12, and T18.
+
+smoke: smoke
+
+seasalt: seasalt
+
+SS Total AOT: Total aerosol optical thickness
+
+date: date
+
+type: denotes daily sampling time period
+
+## `copies_master.csv`
+
+#### Variables:
+
+Type: Denoted daily sampling period
+
+Sample: Site and Day short code. Ex. BO11 = Blind Oso on 7/11/23
+
+Cq: Cycle threshold value from qPCR. Indicated the cycle number that amplification was detected. A lower Cq value indicates earlier detection, and a higher concentration of target,\

### 2023-05-01 products/manuscript/Manuscript.qmd
Updated README and citations

@@
 ---
-title: "Nathan Greenslit Data Analysis Project"
+title: "Dusty Days and a Vibrio Craze: The Influence of Saharan Dust Deposition on Microbial Communities in High Nutrient Coastal Waters "
 subtitle: ""
@@
 Initially comprising a small proportion of the marine microbial communities (\<1%), *Vibrio* can come to quickly dominate, and thus have been classified as 'oppurtunitrophs'.
-The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quiroz2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psenner1992].
+The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quirozInteractionsVibrioPhages2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psennerAreRapidChanges1992].
 
@@
 *Vibrio* growth was observed at Blind Oso Bay, but slightly dampened in comparison to the Gulf.
-This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) [@wetz].
+This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) (Wetz, 2014).
 The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP (Figure 3).

### 2023-05-02 code/analysis_code/stats/5_multi_variate_models/random_forest_models.qmd
Made direct path to images in mansscript

@@
 
+ggsave(here("results", "plots", "rf_bo.png"))
+```
+
+```{r}
 rf_residual_plot <- ggplot(rf_residuals, 
@@
 
+ggsave(here("results", "plots", "rf_c2.png"))
+
+```
+
+```{r}
 rf_residual_plot <- ggplot(rf_residuals, 
@@
 
+ggsave(here("results", "plots", "rf_rd.png"))
+
+```
+
+```{r}
+
 rf_residual_plot <- ggplot(rf_residuals, 

### 2023-05-02 products/manuscript/Manuscript.qmd
Made direct path to images in mansscript

@@
 
-![Figure 1: Dust Aerosol Optical Density (AOD) over daily time series.](images/000009-01.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 1: Dust Aerosol Optical Density (AOD) over daily time series"}
+knitr::include_graphics(here("results", "plots", "dust.png"))
+```
 
@@
 
-![Figure 2: Vibrio gene copies/mL over daily time series across site.](images/000005-03.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 2: Vibrio gene copies/mL over daily time series across site"}
+knitr::include_graphics(here("results", "plots", "copies.png"))
+```
 
@@
 
-![Figure 3: Envonmental parameters across daily time series with color by site.](images/000019.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Envonmental parameters across daily time series with color by site"}
+knitr::include_graphics(here("results", "plots", "env.png"))
+```
 
@@
 
-![Figure 5: Linear Regression Output for Dust AOD (log-transformed) and Vibrio gene copies/mL](images/000055.png){width="1491"}
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 5: Linear Regression Output for Dust AOD (log-transformed) and Vibrio gene copies/mL"}
+knitr::include_graphics(here("results", "plots", "uni.png"))
+```
 
@@
 
-![Figure 6: Vibrio gene copies/mL at Blind Oso predicted by Random Forest](images/000017-02.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 6: Vibrio gene copies/mL at Blind Oso predicted by Random Forest"}
+knitr::include_graphics(here("results", "plots", "rf_bo.png"))
+```
 
-![Figure 7: Vibrio gene copies/mL at Canals predicted by Random Forest](images/000007-02.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 7: Vibrio gene copies/mL at Canals predicted by Random Forest"}
+knitr::include_graphics(here("results", "plots", "rf_c2.png"))
+```
 
-![Figure 8: Vibrio gene copies/mL at the Gulf predicted by Random Forest](images/000007-03.png)
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 8: Vibrio gene copies/mL at the Gulf predicted by Random Forest"}
+knitr::include_graphics(here("results", "plots", "rf_rd.png"))
+```
 

### 2023-05-03 products/manuscript/Manuscript.qmd
Grammer /citations checks for manuscript

@@
 
-The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\].
+The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986].
 Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico.
 These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each.
-Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004\], fungi [@ramirez-camejoFungalDiversitySahara2022\], virus-like particles [@griffinAfricanDesertDust\], and minerals [@formentiChemicalCompositionMineral2003\].
-These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989; @grahamAtmosphericTransportPhosphorus1982\] .
+Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004], fungi [@ramirez-camejoFungalDiversitySahara2022], virus-like particles [@griffinAfricanDesertDust2001], and minerals [@formentiChemicalCompositionMineral2003].
+These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989; @grahamAtmosphericTransportPhosphorus1982].
 This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups [@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].
@@
 To date, our understanding of *Vibrio* and other microbial response to dust input has focused on oligotrophic settings, where dust input is considered to be a critical source of limiting nutrients.
-However, the composition of desert dust can be complex and may have the potential to deliver critical resources that could be exploited by opportunistic microbes (like *Vibrio*) and elicit a growth response despite higher baseline nutrient levels.
+It is unknown how comparable these findings are to coastal areas with higher ambient nutrient levels than these oligotrophic environments.
+The composition of desert dust can be complex and may have the potential to deliver critical resources that could be exploited by opportunistic microbes (like *Vibrio*) and elicit a growth response despite higher baseline nutrient levels.
 Given that human exposure is likely in these coastal waters, it is important to elucidate the relationship between dust input and microbial blooms in these higher nutrient environments.
@@
 Blind Oso Bay is a shallow tributary that is popular amongst wadefisherman and kayakers.
-The Bay is often impacted by fluctuating salinities and high levels of nutrients (baseline:13.03 µM nitrate+nitrite, 4.44 µM orthophosphate) and chlorophyll due to its proximity to wastewater treatment discharge (Wetz, 2014).­ Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
+The Bay is often impacted by fluctuating salinities and high levels of nutrients (average:13.03 µM nitrate+nitrite, 4.44 µM orthophosphate) and chlorophyll due to its proximity to wastewater treatment discharge (Wetz, 2014).­ Additionally, the Bay has persistent issues with high levels of fecal indicator bacteria, placing it on the U.S.
 impaired waters list (Nicolau & Hill, 2014, Texas Commission on Environmental Quality, 2022).
-The residential canal system (baseline: 0.69µM nitrate+nitrite, 0.20 µM orthophosphate), and can also be impacted by high nutrients and chlorophyll from storm water runoff; salinity is primarily driven by precipitation.
-The Gulf site is characterized by constant salinities (\~36) and relatively low baseline nutrient levels (baseline: 0.19µM nitrate+nitrite and 0.36µM orthophosphate).
+The residential canal system (average: 0.69µM nitrate+nitrite, 0.20 µM orthophosphate), and can also be impacted by high nutrients and chlorophyll from storm water runoff; salinity is primarily driven by precipitation.
+The Gulf site is characterized by constant salinities (\~36) and relatively low baseline nutrient levels (average: 0.19µM nitrate+nitrite and 0.36µM orthophosphate).
 
@@
 This project works with three primary data sets.
-(1) Enumeration of Vibrio bacteria as copies/mL from qPCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Temperature and Salinity across the daily time series (ysi_2.csv).
+(1) Enumeration of Vibrio bacteria as copies/mL from qPCR (copies_master.csv), (2) Dust concentrations as aerosol optical thickness (AOT) derived from the Naval Research Lab (nrl_conc.csv), and (3) Environmental Parameters across the daily time series (ysi_2.csv).
 Below is a brief summary of how each data set was cleaned.
@@
 qPCR was conducted in triplicate, so the last step was to take the average of the three replicates to have a final value per sample.
+Samples that were non-detect recieved a value of 0 copies/mL.
 A date column was also added for each sample.
@@
 The daily time series captured two dust events.
-One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 55-57.
+One light event starting on the 13th of July, where AOD reached 0.2, and a much heavier dust event from the 16th to the 18th of July with AOD = 0.55-0.57.
 Dust AOD returned to lower concentrations following the time series.
@@
 
-```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Envonmental parameters across daily time series with color by site"}
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Envonmental parameters across daily time series with color by site(Blind Oso = green, Canals = red, and the Gulf = Blue"}
 knitr::include_graphics(here("results", "plots", "env.png"))
@@
 
-Previous literature has shown that Vibrio respond 12-24 hours following dust deposition events.
+Previous literature has shown that Vibrio respond 12-24 hours following dust deposition events [@westrichSaharanDustNutrients2016a].
 Based on this, we need to look at lags between the two times series for dust data and growth data.
@@
 
-CCF plots showed that BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1.
+CCF plots (found in supplemtary scripts) showed that BO has a significant lag relationship a value of -1.
+In other words, dust occurs, and one day later, we see a response in *Vibrio* gene copies per mL.
+C2 has a lag at -2, and RD at -1.
 
@@
 
-#### Linear Regressions
+#### Multi-varaiate Linear Regressions
 
@@
 
-Multi-variate models that included dust, water chemistry, nutrients, and DOM performed the best, with mean RMSE values much lower than the Null Model.
+Multi-variate models that included dust, water chemistry, nutrients, and DOM (Model 5) resulted in the best performance, with mean RMSE values much lower than the Null Model.
 
@@
 The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for bloom formation.
-*Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes [@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
+*Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962], and are able to exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes [@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016].
 Initially comprising a small proportion of the marine microbial communities (\<1%), *Vibrio* can come to quickly dominate, and thus have been classified as 'oppurtunitrophs'.
-The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quirozInteractionsVibrioPhages2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psennerAreRapidChanges1992].
+The obsereved declines in *Vibrio* gene copies/mL following deposition may be attributed to top down control, such as viral lysis [@molina-quirozInteractionsVibrioPhages2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psennerAreRapidChanges1992].
 
 Similar to previous findings set in the Florida Keys and North Atlantic [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a], peaks in growth were observed 24-48 hours following dust deposition.
-Comparing peak *Vibrio* gene copies/mL to initial levels on the 7th of July, the Gulf experienced a 338-fold increase in *Vibrio* gene copies/mL, Blind Oso a 3.4-fold increase, and the Canals a 1.3-fold increase.
+Comparing peak *Vibrio* gene copies/mL to initial levels on the 9th of July (chosen to ameliorate the possibility of growth following dust deposition prior to sampling period), the Gulf experienced a 33-fold increase in *Vibrio* gene copies/mL, Blind Oso a 2-fold increase, and the Canals a 5.6-fold increase.
 
@@
 
-The Gulf exhibited a sharper increase during the second, heavier dust event, while Blind Oso exhibited a smaller increase during the heavy dust event (Figure 2).
+The Gulf exhibited a sharper increase in growth during the second, heavier dust event, while Blind Oso exhibited a smaller increase during the heavy dust event (Figure 2).
 This difference in response may be attributed to the nutrient dynamics of the sites.
 In marine waters, nutrients are typically limiting due to tight microbial recycling.
-Therefore the introduction on nutrients at such a site may result in rapid uptake and utilization.
+Therefore the introduction of nutrients at such a site may result in rapid uptake and utilization.
 At Blind Oso, a priming effect may be occurring, where dust constituents resulting from a secondary flux may not be as essential for growth following an initial deposition event, where the constituents are reaching a maximum in how much they can support growth.
@@
 
-While it can be proposed that the degree of microbial response is contingent on the environmental settings of the site, further work is needed to look at each specific environmental parameter and how it may impact *Vibrio* abundance.
+While it can be proposed that the degree of microbial response is contingent on the environmental settings of the site, further work is needed to look at each specific environmental parameter and assess how it may impact *Vibrio* abundance.
 At Blind Oso for example, the drastic drop in salinity during an eight day period coincided with dust input.

### 2023-05-04 code/analysis_code/stats/3_correlation.qmd
Final Edits :)

@@
 #####
-
-It's also possible to **combine correlogram with the significance test**. We'll use the result *res.cor2* generated in the previous section with **rcorr**() function \[in **Hmisc** package\]:
-
-<!--# Get this to work -->
-
-```{r}
-# Insignificant correlation are crossed
-#corrplot(bclean$r, type="upper", order="hclust", 
-        # p.mat = bclean$P, sig.level = 0.05, insig = "blank")
-
-#corrplot(cclean$r, type="upper", order="hclust", 
-        # p.mat = cclean$P, sig.level = 0.05, insig = "blank")
-
-#corrplot(rclean$r, type="upper", order="hclust", 
-         #p.mat = rclean$P, sig.level = 0.05, insig = "blank")
-
-```

### 2023-05-04 products/manuscript/Manuscript.qmd
Final Edits :)

@@
 
-```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Envonmental parameters across daily time series with color by site(Blind Oso = green, Canals = red, and the Gulf = Blue"}
+```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Envonmental parameters across daily time series with color by site (Blind Oso = green, Canals = red, and the Gulf = Blue)"}
 knitr::include_graphics(here("results", "plots", "env.png"))
@@
 
-The annual dust data set indicates a period of slightly elevated AOD (0.16) on the 5th of July, prior to sampling.
+The annual dust data set indicates a period of slightly elevated AOD (0.16) on the 5th of July, prior to sampling (SI).
 This light dust period could have the potential to impact initial *Vibrio* levels at the beginning of the time series, which may serve as an explanation for initial higher *Vibrio* gene copies/mL at the Canals, and why the correlation between dust and *Vibrio* gene copies/mL was so high at this site.

### 2023-05-06 code/analysis_code/dust/daily_dusts.qmd
Final Edits
After double checking manuscript, I noticed that figure proportions were off. (small text etc). I went back and readjusted but am wondering why I am just now noticing this. I am working on my laptop (as opposed to my desktop), so i am wondering if R is somehow correcting for screen resolution, and I am seeing different sized plots due to this. Please let me know if my figures look off when you review, thanks!

@@
   dust %>%
-  filter(date %in% as.Date('2022-07-07') : as.Date('2022-07-19'))
+  filter(date %in% as.Date('2022-07-01') : as.Date('2022-07-31'))
 ```
@@
   theme_bw()
+
+ggsave(here("results", "plots", "dust_july.png"))
 ```

### 2023-05-06 code/analysis_code/environment/env_par.qmd
Final Edits
After double checking manuscript, I noticed that figure proportions were off. (small text etc). I went back and readjusted but am wondering why I am just now noticing this. I am working on my laptop (as opposed to my desktop), so i am wondering if R is somehow correcting for screen resolution, and I am seeing different sized plots due to this. Please let me know if my figures look off when you review, thanks!

--- 2023-04-20 code/analysis_code/env/env_par.qmd
@@
 
-```{r, fig.width= 8, fig.height=4}
+```{r}
 fig<- ggarrange(temp,
@@
 fig
+
+ggsave(here("results", "plots", "env.png"))
 ```
 
-# Save as PNG
-
-```{r}
-png(file = here("results", "plots", "temp_sal.png"))
-fig
-dev.off()
-```
+#

### 2023-05-06 products/manuscript/Manuscript.qmd
Final Edits
After double checking manuscript, I noticed that figure proportions were off. (small text etc). I went back and readjusted but am wondering why I am just now noticing this. I am working on my laptop (as opposed to my desktop), so i am wondering if R is somehow correcting for screen resolution, and I am seeing different sized plots due to this. Please let me know if my figures look off when you review, thanks!

@@
 
-Density and Q-Q plots as well as Shapiro-Wilk's tests were conducted to assess distribution of the data.
+Density and Q-Q plots as well as Shapiro-Wilk's tests were conducted to assess distribution of the data (SI).
 Any data that did not have a normal distribution was log-transformed and the test was run again.
@@
 
-CCF plots (found in supplemtary scripts) showed that BO has a significant lag relationship a value of -1.
+CCF plots (found in supplementary file) showed that BO has a significant lag relationship a value of -1.
 In other words, dust occurs, and one day later, we see a response in *Vibrio* gene copies per mL.

### 2023-05-06 products/manuscript/supplement/Supplementary_Material.qmd
Final Edits
After double checking manuscript, I noticed that figure proportions were off. (small text etc). I went back and readjusted but am wondering why I am just now noticing this. I am working on my laptop (as opposed to my desktop), so i am wondering if R is somehow correcting for screen resolution, and I am seeing different sized plots due to this. Please let me know if my figures look off when you review, thanks!

---
title: "Greenslit Project Supplementary Material"

subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Additional Figures

## Dust for July

```{r}
knitr::include_graphics(here("results", "plots", "dust_july.png"))
```

**Dust AOD for month of July. Note slightly elevated dust levels prior to sampling period on the 5th. This may have impacted initial Vibrio levels on the 7th.**

## Distribution Tests (Q-Q Plots)

#### Copies per mL

![](images/paste-2F0DEEC8.png)

![](images/paste-0D0E739C.png)

![](images/paste-879F1B24.png)

#### Dust

![](images/paste-8CCC7BA6.png)

## Cross Correlation Analysis

![](images/paste-3B108E2C.png)

![![](images/paste-62178137.png)](images/paste-022F00A9.png)

**A negative Lag value indicates that x is a predictor of y, or in this case, dust is a predictor of copies_mL- so we will be looking at these negative values. A positive ACF indicates a positive relationship between the two variables, where a negative ACF indicates a negative relationship. According to the ccf plots, BO has a significant lag relationship a value of -1. In other words, dust occurs, and one day later, we see a response in copies_mL. C2 has a lag at -2, and RD at -1**
